### 2024-09-13 paper.qmd
Update paper.qmd
Adding waterfall format, lines 35-41

---
title: "Tournaments"
author: "Greg"
format: html
editor: visual
---

## Tournaments

 - Single elimination (NCAA tournament, NFL playoffs)
  - Seeding issues (Tennis: https://www.liveabout.com/definition-of-seeding-3207821)
   Reseeding after each round.  Straight knockout.  
  - Second faces Third for silver medal.  Guarantees than sivler has 1 loss and bronze has 2 losses.  
  - With a 3rd place match (Badminton: https://www.nbcolympics.com/news/badminton-101-olympic-competition-format)
  - Single elimination with series (NBA, MLB, NHL)
  - Ladder tournaments (Korean baseball does best of 3 with one team starting up 1-0 in the opening round)
  - Seeding issues (NFL re-seeds, NCAA tournament straight bracket)
  - Group/Pool play followed by knockout stages (World Cup, Euro, Overwatch World CUp. )

 - Double elimination (Overwatch Champions League)
  - Repechage (Olympic Wrestling, Tae Kwon Do, Judo) (These also aware 2 bronze medals)
  - True Double elimination
  - Double elimination but a single loss before finals means you can onyl go for third place (College wrestling)

  - Triple Elimination (Curling! https://thegrandslamofcurling.com/bracketology-101-triple-knockout-explained-2021/)

 - Swiss Tournament (https://en.chessbase.com/post/125-years-swiss-system)

 - Round Robin
  - Single and Double.  
  - Baseball is an imbalanced multiple round robin qualifying round.

  - https://www.tandfonline.com/doi/abs/10.1080/01621459.1952.10501178

  - Waterfall format
   - Introduced in the Smash 64 community, [they claim](https://thesmashwriter.wordpress.com/2017/04/24/what-is-a-waterfall-bracket/)
   - Recent example was [Super Smash Con 2023](https://www.start.gg/tournament/super-smash-con-2023/event/64-1v1-singles/overview)
    - Players are seeded $1 \rightarrow n$ and grouped into $k$ divisions, 1 best, $n$ and $k$ worst. 
    - The worst $\frac{n + \epsilon}{k}$ play in [relatively small round robin pools](https://www.start.gg/tournament/super-smash-con-2023/event/64-1v1-singles/brackets/1184919/1833431). Some fraction advance. In this linked example, There are 16 pools in division 4, where each pool is roughly 13 players, with 4 from each division advancing. 
    - The advancing players are then entered into pools with the slightly better-seeded players, and again complete round robins. This iterates over and over again. In this example, to make the numbers work, there are also [wildcard rounds](https://www.start.gg/tournament/super-smash-con-2023/event/64-1v1-singles/brackets/1184927/1833460) to advance an extra few players to further divisions. 
    - In Smash 64, this ultimately leads to a top 16 or 32, which then becomes a double-elimination tournament (all prior losses are ignored). The [top 32](https://www.start.gg/tournament/super-smash-con-2023/event/64-1v1-singles/brackets/1184923/1833456) is seeded by the [top division pool results](https://www.start.gg/tournament/super-smash-con-2023/event/64-1v1-singles/brackets/1184922/1833451).

 - Heat based events (Track 5000m and below, swimming, )

 - Ranking Round followed by single elimination (Olympic Archery https://www.nbcolympics.com/news/archery-101-olympic-competition-format-and-scoring)

 - Elimination racing (Cycling, Bachelor/Bachelorette, Survivor)

 Tournament: a series of contests between a number of competitors, who compete for an overall prize. Oxford Dictionary
  Wiki: "A tournament is a competition involving at least three competitors, all participating in a sport or game"

 (https://www.espn.com/olympics/story/_/id/40536088/2024-paris-summer-olympics-track-field-medaling-format-rules)

Curling and DOE! https://khazna.ku.ac.ae/en/publications/applying-design-of-experiments-tooptimize-the-performance-level-o

Poker: https://arxiv.org/abs/physics/0703122

Graph Theory Stuff: 
Chickens: https://link.springer.com/article/10.1007/BF02476378

Are elections tournaments?

Tournament of Tournaments.

 - aggregation of rankings (golf fedex cup, formula one, tour de france)

https://liquipedia.net/overwatch/Overwatch_Champions_Series/2024/North_America/Stage_2
Swiss, Top Cut Swiss, Group play then a double elimiation bracket!

https://www.vgcguide.com/swiss-and-top-cut

```{r}
results <- data.frame(h = c("i","i","j","i","j","l"), a = c("j","l","l","k","k","k"), hpoints = c(3,3,3,1,1,1), apoints = c(0,0,0,1,1,1))

library(tidyverse)     
h <-results %>% group_by(h) %>% summarize(sum = sum(hpoints))  
a <- results %>% group_by(a) %>% summarize(sum = sum(apoints)) 
names(a) <- c("h","sum")
d <- rbind(h,a)
d %>% group_by(h) %>%  summarize(sum = sum(sum))  %>% arrange(-sum)

#Now reverse the outcomes
results <- data.frame(h = c("i","i","j","i","j","l"), a = c("j","l","l","k","k","k"), hpoints = c(0,0,0,1,1,1), apoints = c(3,3,3,1,1,1))

library(tidyverse)     
h <-results %>% group_by(h) %>% summarize(sum = sum(hpoints))  
a <- results %>% group_by(a) %>% summarize(sum = sum(apoints)) 
names(a) <- c("h","sum")
d <- rbind(h,a)
d %>% group_by(h) %>%  summarize(sum = sum(sum))  %>%arrange(-sum)

```