### 2024-10-13 Tournament_Simulations.qmd
Add files via upload

---
title: "Tournaments"
author: "Zach Culp"
format: html
editor: visual
---

Creating the teams and the strengths

```{r}
# Define the number of teams 
num_teams <- 8
teams <- paste("Team", 1:num_teams)

# Assign random strengths to each team 
team_strengths <- data.frame(
  Team = teams,
  Strength = round(runif(num_teams, min = 0, max = 1),2)
)
#team_strengths
```

Simulating a match using the Bradley Terry model

```{r}
# Function to simulate a match using Bradley-Terry model
simulate_match <- function(team1, team2, strengths) {
  strength1 <- strengths$Strength[strengths$Team == team1]
  strength2 <- strengths$Strength[strengths$Team == team2]

  prob_team1_wins <- strength1 / (strength1 + strength2)  # BT model probability

  winner <- ifelse(runif(1) < prob_team1_wins, team1, team2)  # Randomize outcome
  return(winner)
}
```

# Round Robin Tournaments

Example of a round robin structure (each team plays each other once)

```{r}
simulate_roundrobin <- function(teams, strengths){
  n <- length(teams)
  win_counts <- numeric(n)
  for (i in seq_along(teams)){
    for (j in seq_along(teams)){
      if (i != j){
        match_winner <- simulate_match(teams[i],teams[j],strengths)
        #cat(paste(teams[i], teams[j], match_winner), "\n")
        if (match_winner == teams[i]) {
          win_counts[i] <- win_counts[i] + 1
        } else {
          win_counts[j] <- win_counts[j] + 1
        }
      }
    }
  }
  results <- data.frame(
    Team = teams,
    Wins = win_counts
  )

  # Ranks teams based on number of wins (includes ties)
  results$Rank <- rank(-results$Wins, ties.method = "average")
  results <- results[order(results$Rank),]
  return(results)
}
round_robin <- simulate_roundrobin(teams, team_strengths)
round_robin
```

Simulation of many round robin tournaments.

```{r}
# Define the number of simulations
n <- 10000  # Set the number of simulations you want
total_ranks <- data.frame(matrix(NA, nrow = n, ncol = length(teams)))
colnames(total_ranks) <- teams

# Run multiple simulations
for (i in 1:n) {
  simulation <- simulate_roundrobin(teams, team_strengths)  # Call the tournament simulation function
  ranks <- simulation$Rank
  team_order <- simulation$Team
  total_ranks[i, team_order] <- ranks
}
# Calculate average ranks for each team
average_ranks <- colMeans(total_ranks, na.rm = TRUE)

# Print average ranks for each team
for (team in teams) {
  team_strength <- team_strengths$Strength[team_strengths$Team == team]
  cat(paste(team, "-> Avg. Rank:", average_ranks[team], "Strength:", team_strength), "\n")
}

```

This structure plays out exactly how it would be expected as the ranks follow the order of strengths.

# Single Elimination

Creates an example single elimination tournament with only one simulation.

```{r}
# defining the rank value for each finish
rank_values <- c(
  # Add more ranks as needed
  quarterfinalist = 6.5,
  semifinalist = 3.5,
  finalist = 2,
  champion = 1
)

# Simulate the single-elimination tournament
simulate_tournament <- function(teams, strengths) {
  ranks <- rep(NA, length(teams)) # Creates empty ranks to be filled later
  names(ranks) <- teams  # Assign team names to ranks for easier reference

  round_number <- 1  # To track the current round

  while (length(teams) > 1) {
    #cat("\n--- Round", round_number, "---\n")
    next_round <- c() # Makes sure only winners are included in next round

    for (i in seq(1, length(teams), by = 2)) { # Gets 2 teams for each match
      team1 <- teams[i]
      team2 <- teams[i+1]
      match_winner <- simulate_match(team1, team2, strengths)
      next_round <- c(next_round, match_winner)
      #cat(paste(team1, "vs", team2, "-> Winner:", match_winner, "\n"))

      # Assign ranks based on the current round
      if (round_number == 1) {  # Quarterfinals
        if (match_winner == team1) {
          if (is.na(ranks[team2])) {
            ranks[team2] <- rank_values["quarterfinalist"]
          }
        } else {
          if (is.na(ranks[team1])) {
            ranks[team1] <- rank_values["quarterfinalist"]
          }
        }
      } else if (round_number == 2) {  # Semifinals
        if (match_winner == team1) {
          if (is.na(ranks[team2])) {
            ranks[team2] <- rank_values["semifinalist"]
          }
        } else {
          if (is.na(ranks[team1])) {
            ranks[team1] <- rank_values["semifinalist"]
          }
        }
      } else if (round_number == 3) {
        if (match_winner == team1) {
          if(is.na(ranks[team2])) {
            ranks[team2] <- rank_values["finalist"]
          }
          if(is.na(ranks[team1])) {
            ranks[team1] <- rank_values["champion"]
          }
        } else {
          if(is.na(ranks[team1])) {
            ranks[team1] <- rank_values["finalist"]
          }
          if(is.na(ranks[team2])) {
            ranks[team2] <- rank_values["champion"]
          }
        }
      }
      # No rank assignment here for the final round; handled after loop
    }

    # If only two teams remain, handle the final match
    #if (length(next_round) == 2 && round_number == 3) {
    #  finalist <- setdiff(teams, next_round)
    #  # Assign ranks for the final match
    #  ranks[next_round[1]] <- rank_values["champion"]
    #  ranks[next_round[2]] <- rank_values["finalist"]
    #}

    teams <- next_round  # Winners move to the next round
    round_number <- round_number + 1
  }

  # Handle the case where only one team remains (champion)
  if (length(teams) == 1 && is.na(ranks[teams])) {
    ranks[teams] <- rank_values["champion"]
  }

  return(list(champion = teams[1], ranks = ranks))  # Return champion and ranks
}

# Define the teams for the tournament
teams <- c("Team 1", "Team 2", "Team 3", "Team 4", "Team 5", "Team 6", "Team 7", "Team 8")

# Run the tournament
cat("Starting tournament with teams:", paste(teams, collapse = ", "), "\n")
result <- simulate_tournament(teams,team_strengths)

# Print the results
cat("\nChampion of the tournament is:", result$champion, "\n")
cat("\nTeam Rankings:\n")
for (team in teams) {
  cat(paste(team, "-> Rank:", ifelse(is.na(result$ranks[team]), "Not Ranked", result$ranks[team]), "\n"))
}
team_strengths
```

Simulating games many times

```{r}
# Initialize the teams you want to simulate
team1 <- "Team 1"  # Replace with the actual team name
team2 <- "Team 2"  # Replace with the actual team name

# Initialize a vector to store match results (1 for team1 win, 0 for team2 win)
match_results <- numeric(1000)

# Run 1000 simulations of the match
for (i in 1:1000000) {
  winner <- simulate_match(team1, team2,team_strengths)
  match_results[i] <- ifelse(winner == team1, 1, 0)  # 1 if team1 wins, 0 if team2 wins
}

# Calculate the mean of the match results
mean_result <- mean(match_results)
cat("Mean result (fraction of wins for team1):", mean_result, "\n")
```

Simulating the tournament many times.

```{r}
# Define the number of simulations
n <- 10000  # Set the number of simulations you want

# Initialize a data frame to store total ranks for each team
total_ranks <- data.frame(matrix(NA, nrow = n, ncol = length(teams)))
colnames(total_ranks) <- teams

# Run multiple simulations
for (i in 1:n) {
  simulation <- simulate_tournament(teams, team_strengths)  # Call the tournament simulation function
  total_ranks[i, ] <- simulation$ranks  # Store the ranks of the current simulation
}
total_ranks
# Calculate average ranks for each team
average_ranks <- colMeans(total_ranks, na.rm = TRUE)

# Print average ranks for each team
for (team in teams) {
  team_strength <- team_strengths$Strength[team_strengths$Team == team]
  cat(paste(team, "-> Avg. Rank:", average_ranks[team], "Strength:", team_strength), "\n")
}

```

As seen here, a completely randomized seeding can lead to a higher strength team losing more often than a lower strength team. Because of the high probability of winning the first game, Team 6 actually averages to the highest ranking.

Now, attempting it seeding the teams based on strength.

```{r}
# Sort the teams by their strengths in descending order (for seeding)
teams_seeded <- team_strengths[order(-team_strengths$Strength), ]  # Strongest at the top
seeded_strengths <- data.frame(
  Team = teams_seeded$Team,
  Strength = teams_seeded$Strength
)
# Display the seeding
#cat("Teams sorted by strength (seeds):\n")
#teams_seeded

simulate_seeded_tournament <- function(teams,strengths) {
  ranks <- rep(NA, length(teams))  # Create empty ranks to be filled later
  names(ranks) <- teams  # Set names for easier access
  round_number <- 1
  while (length(teams) > 1) {
    #cat("\n--- Round", round_number, "---\n")
    next_round <- c()  # Ensures only winners are included in the next round
    # Pair teams based on seeding (1 vs 8, 2 vs 7, etc.)
   for (i in 1:(length(teams) / 2)) {
      team1 <- teams[i]                  # Higher seed
      team2 <- teams[length(teams) - i + 1]  # Lower seed
      match_winner <- simulate_match(team1, team2, strengths)
      next_round <- c(next_round, match_winner)
      #cat(paste(team1, "vs", team2, "-> Winner:", match_winner, "\n\n"))

      # Assign ranks based on the current round
      if (round_number == 1) {  # Quarterfinals
        ranks[team1] <- rank_values["quarterfinalist"]
        ranks[team2] <- rank_values["quarterfinalist"]
      } else if (round_number == 2) {  # Semifinals
        if (match_winner == team1) {
          ranks[team2] <- rank_values["semifinalist"]
        } else {
          ranks[team1] <- rank_values["semifinalist"]
        }
      } else if (round_number == 3) {  # Finals
        if (match_winner == team1) {
          ranks[team1] <- rank_values["champion"]
          ranks[team2] <- rank_values["finalist"]
        } else {
          ranks[team2] <- rank_values["champion"]
          ranks[team1] <- rank_values["finalist"]
        }
      }
    }

    teams <- next_round
    round_number <- round_number + 1
  }

  return(list(champion = teams[1], ranks = ranks))  # Final winner
}

# Define the number of simulations
n <- 1000  # Set the number of simulations you want

# Initialize a data frame to store total ranks for each team
total_ranks <- data.frame(matrix(NA, nrow = n, ncol = length(teams)))
colnames(total_ranks) <- teams

# Run multiple simulations
for (i in 1:n) {
  seeded_simulation <- simulate_seeded_tournament(teams_seeded$Team,seeded_strengths)  # Call the tournament simulation function
  seeded_simulation
  total_ranks[i, names(seeded_simulation$ranks)] <- seeded_simulation$ranks  # Store the ranks of the current simulation
}
# Calculate average ranks for each team
average_ranks <- colMeans(total_ranks, na.rm = TRUE)

# Print average ranks for each team
for (team in teams_seeded$Team) {
  team_strength <- teams_seeded$Strength[teams_seeded$Team == team]
  cat(paste(team, "-> Avg. Rank:", average_ranks[team], "Strength:", team_strength), "\n")
}

```

As seen in this example, the ranks are what you would expect given the strengths. This shows that a seeded single elimination tournament structure will likely produce more accurate depictions of strength than a randomized single elimination structure.

The next step will be to include potential byes to this single elimination tournament and then look into double elimination tournaments.