### 2024-11-22 single_elimination.qmd

---
title: "single_elimination"
format: html
editor: visual
---

# Simulate a Single Match

```{r}
# Function to simulate a match using Bradley-Terry model
simulate_match <- function(team1, team2, strength1, strength2) {
  odds <- exp(strength1 - strength2)/(1+exp(strength1 - strength2))

  winner <- ifelse(runif(1) < odds, team1, team2)  # Randomize outcome
  return(winner)
}
```

# Single Elimination Tournament Structures

```{r}
num_teams <- 8
teams <- paste("Team", 1:num_teams)
team_map <- setNames(1:8, teams)
tourn <- do.call(rbind, args = combinat::permn(1:8))
tourn
dist <- apply(X = tourn, MARGIN = 1, FUN = function(x){
  if (sum(x[5:8] == 1) > 0) {
    x[1:8] <- x[c(5:8, 1:4)]
  }
  if (sum(x[3:4] == min(x[1:4])) > 0) {
    x[1:4] <- x[c(3:4, 1:2)]
  }
  if (sum(x[7:8] == min(x[5:8])) > 0) {
    x[5:8] <- x[c(7:8, 5:6)]
  }
  for (i in c(1, 3, 5, 7)) {
    x[i:(i + 1)] <- sort(x[i:(i + 1)])
  }
return(x)
})

dist <- t(dist)
seeds <- dist[!duplicated(dist), ]

# How to check for a specific seed
which(apply(seeds, 1, function(x) all(x == c(1, 8, 2, 7, 3, 6, 4, 5))))

## Important Structures Rows ##
# 1-8: row 1
# (1,8), (2,7), (3,6), (4,5): row 23
```

# Setting Up the Teams and Strengths

```{r}
# Assign random strengths to each team 
team_strengths <- data.frame(
  Team = teams,
  Strength = c(1,2,3,4,5,6,7,8),
  Wins = rep(0,8),
  Ranks = rep(NA,8),
  Exp_Ranks = c(8,7,6,5,4,3,2,1)
)

normal_strengths <- data.frame(
  Team = teams,
  Strength = round(rnorm(8,0,1),2),
  Wins = rep(0,8),
  Ranks = rep(NA,8)
)
normal_strengths$Exp_Ranks <- rank(-normal_strengths$Strength, ties.method = "average")

unif_strength <- data.frame(
  Team = teams,
  Strength = round(runif(8,0,8),2),
  Wins = rep(0,8),
  Ranks = rep(NA,8)
)
unif_strength$Exp_Ranks <- rank(-unif_strength$Strength, ties.method = "average")

# Store strength data frames in a list
strengths_list <- list(
  "Default" = team_strengths,
  "Normal" = normal_strengths,
  "Uniform" = unif_strength
)

```

# Simulating a Tournament

```{r}
# defining the rank value for each finish
rank_values <- c(
  # Add more ranks as needed
  quarterfinalist = 6.5,
  semifinalist = 3.5,
  finalist = 2,
  champion = 1
)

# Simulate the single-elimination tournament
simulate_tournament <- function(df,seeds = NULL) {

  # Sets seed to that of common best vs worst structure
  if (is.null(seeds)){
    seeds <- matrix(c(1,8,2,7,3,6,4,5), nrow = 1)
  }

  results <- list()  # Initialize the results list

  for (i in 1:nrow(seeds)) {
    # Create a new data frame based on the current permutation in 'test'
    permutation <- seeds[i, ]
    permuted_df <- df
    permuted_df$Team <- df$Team[permutation]  # Update team order 
    permuted_df$Strength <- df$Strength[permutation]
    permuted_df$Ranks <- NA  # Initialize ranks column
    permuted_df$Wins <- 0  # Initialize wins column
    permuted_df$Exp_Rank <- df$Exp_Rank[permutation]

    teams <- permuted_df$Team
    strengths <- permuted_df$Strength

    round_number <- log2(length(teams))

  while (length(teams) > 1) {
    #cat("\n--- Round", round_number, "---\n")
    next_round <- c() # Makes sure only winners are included in next round

    for (j in seq(1, length(teams), by = 2)) { # Gets 2 teams for each match
      team1 <- teams[j]
      team2 <- teams[j+1]

      strength1 <- strengths[permuted_df$Team == team1]
      strength2 <- strengths[permuted_df$Team == team2]

      match_winner <- simulate_match(team1, team2, strength1, strength2)
      next_round <- c(next_round, match_winner)
      #cat(paste(team1, "vs", team2, "-> Winner:", match_winner, "\n"))

      # Can add more rounds by the same code with rounnd_number increasing by 1 and rank_values[round name]

      # Assign ranks based on the current round
      if (round_number == 3) {  # Quarterfinals
        if (match_winner == team1) {
          permuted_df$Ranks[permuted_df$Team == team2] <- 
          rank_values["quarterfinalist"]
        } else {
          permuted_df$Ranks[permuted_df$Team == team1] <-   
          rank_values["quarterfinalist"]
        }
      } else if (round_number == 2) {  # Semifinals
        if (match_winner == team1) {
          permuted_df$Ranks[permuted_df$Team == team2] <-
          rank_values["semifinalist"]
        } else {
          permuted_df$Ranks[permuted_df$Team == team1] <-
          rank_values["semifinalist"]
        }
      } else if (round_number == 1) { # Finals
        if (match_winner == team1) {
          permuted_df$Ranks[permuted_df$Team == team2] <- rank_values["finalist"]
        permuted_df$Ranks[permuted_df$Team == team1] <- rank_values["champion"]
        permuted_df$Wins[permuted_df$Team == team1] <-
        permuted_df$Wins[permuted_df$Team == team1] + 1
        } else {
          permuted_df$Ranks[permuted_df$Team == team1] <- rank_values["finalist"]
          permuted_df$Ranks[permuted_df$Team == team2] <- rank_values["champion"]
          permuted_df$Wins[permuted_df$Team == team2] <-
          permuted_df$Wins[permuted_df$Team == team2] + 1
        }
      }
    }

    teams <- next_round  # Winners move to the next round
    round_number <- round_number - 1
  }

  results[[i]] <- permuted_df # Store the results of this permutation
  }
  return(do.call(rbind, results))
}

```

# Single Tournament Simulation Results

```{r}
result <- simulate_tournament(team_strengths,seeds)
result[order(-result$Strength),]

norm_result <- simulate_tournament(normal_strengths,seeds)
norm_result[order(-norm_result$Strength),]

unif_result <- simulate_tournament(unif_strength,seeds)
unif_result[order(-unif_result$Strength),]
```

# Multiple Simulations

```{r}
tourn_sims <- function(df, seeds = NULL, replicates = 100) {
  tournament_replicates <- replicate(replicates, {
    # Reset wins and ranks for each tournament
    df$Wins <- rep(0, nrow(df))
    df$Ranks <- rep(NA, nrow(df))

    if (is.null(seeds)) {
      seeds <- matrix(c(1, 8, 2, 7, 3, 6, 4, 5), nrow = 1)
    }

    # Run the tournament
    result <- simulate_tournament(df, seeds)

    return(result[, c("Team", "Strength", "Wins", "Ranks", "Exp_Rank")])  
  }, simplify = FALSE)  # Prevents simplification to atomic vector

  combined_results <- do.call(rbind, tournament_replicates)

  # Calculate average ranks and total wins using dplyr
  results_summary <- combined_results %>%
    group_by(Team, Strength, Exp_Rank) %>%  # Track SeedNum
    summarise(
      Average_Rank = mean(Ranks, na.rm = TRUE),
      Total_Wins = sum(Wins, na.rm = TRUE)
    ) %>%
    arrange(desc(Strength))

  return(results_summary)
}

# Run the simulation for each strength configuration
results_sim <- lapply(strengths_list, tourn_sims)

# Print the results for each configuration
for (name in names(results_sim)) {
  cat("\nResults for strength configuration:", name, "\n")
  print(results_sim[[name]])

  # Calculate Kendall tau correlation
  kendall_cor <- cor(results_sim[[name]]$Average_Rank, results_sim[[name]]$Exp_Rank, method = "kendall", use = "complete.obs")
  cat("Kendall Tau Correlation:", kendall_cor, "\n")

  # Calculate Spearman correlation
  spearman_cor <- cor(results_sim[[name]]$Average_Rank, results_sim[[name]]$Exp_Rank, method = "spearman", use = "complete.obs")
  cat("Spearman Correlation:", spearman_cor, "\n")

  # Create a barplot for total wins
  barplot(
    results_sim[[name]]$Total_Wins,
    names.arg = results_sim[[name]]$Strength,
    main = paste("Total Wins for", name),
    xlab = "Strength",
    col = "blue",  # Added color for better visualization
    border = "black"  # Added border color for bars
  )
}

```

# Spearman Correlation Best and Worst Cases

```{r}
all_results <- function (strength){
    all_seeds <- list()
    total_seeds <- nrow(seeds)  # Store the total number of seeds
    for (i in 1:total_seeds) {
        current_seed <- seeds[i, , drop = FALSE]
        result <- suppressMessages(tourn_sims(team_strengths, current_seed))
        result$SeedNum <- i  # Add seed number to the result
        all_seeds[[paste0("Seed_", i)]] <- result
    }
    return(all_seeds)
}

# Store strength data frames in a list
sims_list <- list(
    "Default" = all_results(team_strengths),
    "Normal" = all_results(normal_strengths),
    "Uniform" = all_results(unif_strength)
)

# Initialize a list to store Spearman correlation results separately
spearman_results_list <- list()

# Loop through each simulation and calculate Spearman correlation
for (name in names(sims_list)) {
    sim_results <- do.call(rbind, sims_list[[name]])

    # Calculate the Spearman correlation only once for the unique ranks
    spearman_cor <- cor(sim_results$Average_Rank, sim_results$Exp_Rank, method = "spearman", use = "complete.obs")

    # Store the overall Spearman correlation result
    spearman_results_list[[name]] <- data.frame(Configuration = name, 
                                                  Spearman_Correlation = spearman_cor)

    # Keep only relevant columns and ensure distinct SeedNum
    distinct_results <- sim_results[, c("SeedNum", "Average_Rank", "Exp_Rank")]

    # Calculate the Spearman correlation for each distinct SeedNum
    distinct_spearman_results <- distinct_results %>%
        group_by(SeedNum) %>%
        summarize(Spearman_Correlation = cor(Average_Rank, Exp_Rank, method = "spearman", use = "complete.obs")) %>%
        ungroup()

    # Get top 5 highest distinct Spearman correlations for this configuration
    top_5 <- distinct_spearman_results %>%
        arrange(desc(Spearman_Correlation)) %>%
        slice_head(n = 5)  # Select top 5

    # Get top 5 lowest distinct Spearman correlations for this configuration
    lowest_5 <- distinct_spearman_results %>%
        arrange(Spearman_Correlation) %>%
        slice_head(n = 5)  # Select lowest 5

    # Print results with seed numbers
    cat(paste("\nTop 5 Distinct Spearman Correlations for configuration:", name, "\n"))
    print(top_5)  # Only SeedNum and Spearman_Correlation

    cat(paste("\nLowest 5 Distinct Spearman Correlations for configuration:", name, "\n"))
    print(lowest_5)  # Only SeedNum and Spearman_Correlation
}

# Print overall Spearman results for all configurations
 for (name in names(spearman_results_list)) {
     cat(paste("\nSpearman Correlation for configuration:", name, "\n"))
     print(spearman_results_list[[name]])
 }

```