### 2025-02-20 single_elimination.qmd

---
title: "single_elimination"
format: 
  html:
    embed-resources: true
editor: visual
---

```{r}
library(tidyverse)
library(dplyr)
library(shiny)
```

# Simulate a Single Match

```{r}
# Function to simulate a match using Bradley-Terry model
simulate_match <- function(team1, team2, strength1, strength2) {
  if(is.na(team1)){
    winner <- team2
  }
  else if(is.na(team2)){
    winner <- team1
  }
  else {
  p <- exp(strength1 - strength2)/(1+exp(strength1 - strength2))
  winner <- ifelse(runif(1) < p, team1, team2)  # Randomize outcom
  }
  return(winner)
}
# sample("A","B"),p=(p,1-p)

```

# Single Elimination Tournament Structures

## 8 Teams (all combinations)

```{r}
num_teams <- 8
teams <- paste("Team",num_teams:1)
team_map <- setNames(1:8, teams)
tourn <- do.call(rbind, args = combinat::permn(1:8))
dist <- apply(X = tourn, MARGIN = 1, FUN = function(x){
  if (sum(x[5:8] == 1) > 0) {
    x[1:8] <- x[c(5:8, 1:4)]
  }
  if (sum(x[3:4] == min(x[1:4])) > 0) {
    x[1:4] <- x[c(3:4, 1:2)]
  }
  if (sum(x[7:8] == min(x[5:8])) > 0) {
    x[5:8] <- x[c(7:8, 5:6)]
  }
  for (i in c(1, 3, 5, 7)) {
    x[i:(i + 1)] <- sort(x[i:(i + 1)])
  }
return(x)
})

dist <- t(dist)
seeding_structure <- dist[!duplicated(dist), ]
seeding_structure_char <- apply(seeding_structure,1,function(x) {paste0(x,collapse="")})
# How to check for a specific seed
which(apply(seeding_structure, 1, function(x) all(x == c(1, 8, 4, 5, 2, 7, 3, 6))))

## Important Structures Rows ##
# 1-8: row 1
# (1,8), (2,7), (3,6), (4,5): row 23
```

## 4 Teams (all combinations)

```{r}
num_teams <- 4
teams <- paste("Team",num_teams:1)
team_map <- setNames(1:num_teams, teams)
tourn <- do.call(rbind, args = combinat::permn(1:num_teams))
dist <- apply(X = tourn, MARGIN = 1, FUN = function(x){
  if (sum(x[3:4] == min(x[1:4])) > 0) {
    x[1:4] <- x[c(3:4, 1:2)]
  }
  for (i in c(1, 3)) {
    x[i:(i + 1)] <- sort(x[i:(i + 1)])
  }
return(x)
})
dist <- t(dist)
seeding_structure <- dist[!duplicated(dist), ]
seeding_structure_char <- apply(seeding_structure,1,function(x) {paste0(x,collapse#="")})

#num_teams <- 5
#teams <- paste("Team",num_teams:1)
#team_map <- setNames(1:num_teams, teams)
#tourn <- do.call(rbind, args = combinat::permn(1:num_teams))
#tourn[tourn[,1] == 1,]
```

## Creating the traditional tournament structure for n teams

```{r}
generate_tournament_structure <- function(n) {

  is_power_of_two <- function(x) {
    x > 0 && (x & (x - 1)) == 0
  }

  # If n is not a power of 2, round up to the next power of 2 and add placeholders
  if (!is_power_of_two(n)) {
    next_power_of_two <- 2^ceiling(log2(n))
    extra_teams <- next_power_of_two - n
    message(sprintf("Rounding up to %d teams. Adding %d NA placeholders.", next_power_of_two, extra_teams))
    n <- next_power_of_two
  } else {
    extra_teams <- 0
  }

  rounds <- log(n, base = 2) - 1
  teams <- c(1, 2)  # Initial list of teams

  for (i in 1:rounds) {
    teams <- nextLayer(teams)
  }

  while (extra_teams > 0){
      teams[which.max(teams)] <- NA
      extra_teams <- extra_teams - 1
  }

  return(teams)
}

nextLayer <- function(teams) {
  out <- c()  # Initialize an empty vector
  length <- length(teams) * 2 + 1  # Calculate the length for pairing

  # Generate the next layer by pairing teams
  for (i in teams) {
    out <- c(out, i, length - i)  # Push the current team and its pair
  }

  return(out)
}

```

# Setting Up the Teams and Strengths

```{r}
# Assign random strengths to each team coming from normal and uniform distributions
normal_strengths <- data.frame(
  Team = teams,
  Strength = sapply(num_teams, function(n) { qnorm(1:n/(n+1)) }),
  Wins = rep(0,num_teams),
  Ranks = rep(NA,num_teams)
)
normal_strengths$True_Rank <- rank(-normal_strengths$Strength, ties.method = "average")
normal_strengths <- arrange(normal_strengths, True_Rank)

unif_strength <- data.frame(
  Team = teams,
  Strength = sapply(num_teams, function(n) { qunif(1:n/(n+1) , 0 , sqrt(12)) }),
  Wins = rep(0,num_teams),
  Ranks = rep(NA,num_teams)
)
unif_strength$True_Rank <- rank(-unif_strength$Strength, ties.method = "average")
unif_strength <- arrange(unif_strength, True_Rank)

# Store strength data frames in a list
strengths_list <- list(
  "Normal" = normal_strengths,
  "Uniform" = unif_strength
)

```

# Simulating a Tournament

```{r}
# defining the rank value for each finish
rank_values <- c(
  # Add more ranks as needed
  round_of_sixteen = 12.5,
  quarterfinalist = 6.5,
  semifinalist = 3.5,
  fourth_place = 4,
  third_place = 3,
  finalist = 2,
  champion = 1
)

# Remove ties in rankings
noTies <- function(df) {
  quarterfinals <- df %>%
    filter(Ranks == rank_values["quarterfinalist"]) %>%
    mutate(Ranks = sample(5:8, n(), replace = FALSE))  # Assigns unique ranks
  semifinals <- df %>%
    filter(Ranks == rank_values["semifinalist"]) %>%
    mutate(Ranks = sample(3:min(n(),4), n(), replace = FALSE))

  df <- df %>%
    filter(Ranks != rank_values["quarterfinalist"], Ranks != rank_values["semifinalist"]) %>%
    bind_rows(quarterfinals,semifinals)  # Replace old quarterfinalists with unique ranks

  return(df)
}

# Simulate the single-elimination tournament
simulate_tournament <- function(df,seeding_structure = NULL, ties=T, series=1, third_place = T) {

  # Sets seed to that of common best vs worst structure
  n <- length(unique(df$Team))
  if (is.null(seeding_structure)){
    seeding_structure <- matrix(generate_tournament_structure(n), nrow = 1)
  } 
  if(series %% 2 == 0) {
    stop("Series must be an odd number to avoid potential ties.")
  }

  results <- list()  # Initialize the results list

  for (i in 1:nrow(seeding_structure)) {
    # Create a new data frame based on the current permutation in 'test'
    permutation <- seeding_structure[i, ]
    permuted_df <- df
    permuted_df$Team <- df$Team[permutation]  # Update team order 
    permuted_df$Strength <- df$Strength[permutation]
    permuted_df$Ranks <- NA  # Initialize ranks column
    permuted_df$Wins <- 0  # Initialize wins column
    permuted_df$True_Rank <- df$True_Rank[permutation]

    teams <- permuted_df$Team
    strengths <- permuted_df$Strength

    round_number <- log2(length(teams))

    # Track the losers of the semifinals for the third-place match
    losers_semis <- c()

  while (length(teams) > 1) {
    #cat("\n--- New Round ---\n")
    next_round <- c() # Makes sure only winners are included in next round

    for (j in seq(1, length(teams), by = 2)) { # Gets 2 teams for each match
      team1 <- teams[j]
      team2 <- teams[j+1]

      strength1 <- strengths[permuted_df$Team == team1]
      strength2 <- strengths[permuted_df$Team == team2]

      wins_team1 <- 0
      wins_team2 <- 0
      for (game in 1:series){
      match_winner <- simulate_match(team1, team2, strength1, strength2)
      if (match_winner == team1) {
            wins_team1 <- wins_team1 + 1
          } else {
            wins_team2 <- wins_team2 + 1
          }

          # Determine the overall winner if a majority is reached
          if (wins_team1 > series / 2 || wins_team2 > series / 2) {
            break
          }
        }

        # Determine the match winner based on series results
        series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
      #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, series_winner))
      next_round <- c(next_round, series_winner)

      # Can add more rounds by the same code with round_number increasing by 1 and rank_values[round name]

      # Assign ranks based on the current round
      if (round_number == 3) {  # Quarterfinals
        if (match_winner == team1) {
          permuted_df$Ranks[permuted_df$Team == team2] <- 
          rank_values["quarterfinalist"]
        } else {
          permuted_df$Ranks[permuted_df$Team == team1] <-   
          rank_values["quarterfinalist"]
        }
      } else if (round_number == 2) {  # Semifinals
        if (match_winner == team1) {
          permuted_df$Ranks[permuted_df$Team == team2] <-
          rank_values["semifinalist"]
          losers_semis <- c(losers_semis, team2)
        } else {
          permuted_df$Ranks[permuted_df$Team == team1] <-
          rank_values["semifinalist"]
          losers_semis <- c(losers_semis, team1)
        }

      } else if (round_number == 1) { # Finals
        if (match_winner == team1) {
          permuted_df$Ranks[permuted_df$Team == team2] <- rank_values["finalist"]
        permuted_df$Ranks[permuted_df$Team == team1] <- rank_values["champion"]
        permuted_df$Wins[permuted_df$Team == team1] <-
        permuted_df$Wins[permuted_df$Team == team1] + 1
        } else {
          permuted_df$Ranks[permuted_df$Team == team1] <- rank_values["finalist"]
          permuted_df$Ranks[permuted_df$Team == team2] <- rank_values["champion"]
          permuted_df$Wins[permuted_df$Team == team2] <-
          permuted_df$Wins[permuted_df$Team == team2] + 1
        }
      }
    }

    teams <- next_round  # Winners move to the next round
    round_number <- round_number - 1
  }

  # Third place match between the losers of the semifinals
  if (third_place == TRUE){
    team1 <- losers_semis[1]
    team2 <- losers_semis[2]
    strength1 <- strengths[permuted_df$Team == team1]
    strength2 <- strengths[permuted_df$Team == team2]
    wins_team1 <- 0
    wins_team2 <- 0
    for (game in 1:series) {
      match_winner <- simulate_match(team1, team2, strength1, strength2)
      if (match_winner == team1) {
        wins_team1 <- wins_team1 + 1
      } else {
        wins_team2 <- wins_team2 + 1
      }
     # Determine the overall winner if a majority is reached
      if (wins_team1 > series / 2 || wins_team2 > series / 2) {
      break
    }
   # Assign third place and fourth place ranks
      if (wins_team1 > wins_team2) {
        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team1))
        permuted_df$Ranks[permuted_df$Team == team1] <- rank_values["third_place"]
        permuted_df$Ranks[permuted_df$Team == team2] <- rank_values["fourth_place"]
      } else {
        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team2))
        permuted_df$Ranks[permuted_df$Team == team2] <- rank_values["third_place"]
        permuted_df$Ranks[permuted_df$Team == team1] <- rank_values["fourth_place"]
      }
    }
  }

  results[[i]] <- permuted_df # Store the results of this permutation
  }

  final_results <- do.call(rbind, results)
  # Remove ties if ties = FALSE
  if (!ties) {
    final_results <- noTies(final_results)
  }

  return(final_results)
}

test <- simulate_tournament(normal_strengths,ties=T, series=3)
test[order(-test$Strength),]

# Run the tournament simulation
simulation_results <- replicate(1000, simulate_tournament(unif_strength), simplify = FALSE)

# Combine all replicate results
all_results <- do.call(rbind, simulation_results)

# Summarize total wins
total_wins_summary <- all_results %>%
  group_by(Team) %>%
  summarise(
    Total_Wins = sum(Wins, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(Total_Wins))

# Print the total wins summary
print(total_wins_summary)

# Run the tournament simulation
simulation_series <- replicate(1000, simulate_tournament(unif_strength, series=7), simplify = FALSE)

# Combine all replicate results
series_results <- do.call(rbind, simulation_series)

# Summarize total wins
total_wins_series <- series_results %>%
  group_by(Team) %>%
  summarise(
    Total_Wins = sum(Wins, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(Total_Wins))

# Print the total wins summary
print(total_wins_series)

```

# Single Tournament Simulation Results

```{r}
norm_result <- simulate_tournament(normal_strengths,seeding_structure)
norm_result[order(-norm_result$Strength),]

unif_result <- simulate_tournament(unif_strength,seeding_structure)
unif_result[order(-unif_result$Strength),]
```

# Multiple Simulations for 1 Seeding Structure

```{r}
tourn_sims <- function(df, seeding_structure = NULL, seed_char = NULL, replicates = 100, series=1) {
  # Default seeding structure if none provided
  if (is.null(seeding_structure)) {
    seeding_structure <- matrix(c(1,8,4,5,2,7,3,6), nrow = 1)
    seed_char <- apply(seeding_structure, 1, function(x) paste0(x, collapse = ""))
  }

  # List to store results for each seed configuration
  tournament_replicates <- lapply(1:nrow(seeding_structure), function(seed_num) {
    replicate(replicates, {
      # Reset wins and ranks for each tournament
      df$Wins <- rep(0, nrow(df))
      df$Ranks <- rep(NA, nrow(df))

      # Run the tournament with the current seed configuration
      result <- simulate_tournament(df, seeding_structure[seed_num, , drop = FALSE], series = series)

      # Add Seed_Num column to keep track of the current seed configuration
      result$Seed_Num <- seed_num
      result$Seeding_Structure <- seed_char[seed_num]

      return(result[, c("Team", "Strength", "Wins", "Ranks", "True_Rank", "Seed_Num", "Seeding_Structure")])  
    }, simplify = FALSE)
  })

  # Combine results for all seed configurations
  combined_results <- do.call(rbind, unlist(tournament_replicates, recursive = FALSE))

  # Summarize results
  results_summary <- combined_results %>%
    group_by(Seed_Num, Seeding_Structure, Team, Strength, True_Rank) %>%
    summarise(
      Average_Rank = mean(Ranks, na.rm = TRUE),
      Total_Wins = sum(Wins, na.rm = TRUE),
      .groups = 'drop'
    ) %>%
    arrange(desc(Strength))

  return(results_summary)
}

# Run the simulation for each strength configuration
results_sim <- lapply(strengths_list, tourn_sims)

# Print the results for each configuration
for (name in names(results_sim)) {
  cat("\nResults for strength configuration:", name, "\n")
  print(results_sim[[name]])

  # Calculate Kendall tau correlation
  kendall_cor <- cor(results_sim[[name]]$Average_Rank, results_sim[[name]]$True_Rank, method = "kendall", use = "complete.obs")
  cat("Kendall Tau Correlation:", kendall_cor, "\n")

  # Calculate Spearman correlation
  spearman_cor <- cor(results_sim[[name]]$Average_Rank, results_sim[[name]]$True_Rank, method = "spearman", use = "complete.obs")
  cat("Spearman Correlation:", spearman_cor, "\n")

  # Create a barplot for total wins
  bar <- barplot(
    results_sim[[name]]$Total_Wins,
    names.arg = results_sim[[name]]$True_Rank,
    main = paste("Total Wins for", name),
    xlab = "True Rank",
    ylim = c(0,100),
    col = "blue",  # Added color for better visualization
    border = "black"  # Added border color for bars
  )
  text(
    x = bar,
    y = results_sim[[name]]$Total_Wins + 50,
    labels = results_sim[[name]]$Total_Wins,
    cex = 0.8,
    col = "black"
  )
}

num_games <- 7
series_sim <- lapply(strengths_list, function(strength_config) {
  tourn_sims(strength_config, series = num_games)
})
# Print the results for each configuration
for (name in names(series_sim)) {
  cat("\nResults for strength configuration:", name, "\n")
  print(series_sim[[name]])

  # Calculate Kendall tau correlation
  kendall_cor <- cor(series_sim[[name]]$Average_Rank, series_sim[[name]]$True_Rank, method = "kendall", use = "complete.obs")
  cat("Kendall Tau Correlation:", kendall_cor, "\n")

  # Calculate Spearman correlation
  spearman_cor <- cor(series_sim[[name]]$Average_Rank, series_sim[[name]]$True_Rank, method = "spearman", use = "complete.obs")
  cat("Spearman Correlation:", spearman_cor, "\n")

  # Create a barplot for total wins
  bar <- barplot(
    series_sim[[name]]$Total_Wins,
    names.arg = series_sim[[name]]$True_Rank,
    main = paste("Total Wins for", name),
    xlab = "True Rank",
    ylim = c(0,100),
    col = "blue",  # Added color for better visualization
    border = "black"  # Added border color for bars
  )
  text(
    x = bar,
    y = series_sim[[name]]$Total_Wins + 50,
    labels = series_sim[[name]]$Total_Wins,
    cex = 0.8,
    col = "black"
  )
}
```

# Probability of Each Seed Winning With Each Structure

```{r}
library(dplyr)
library(ggplot2)

sims_norm <- tourn_sims(normal_strengths,seeding_structure, seeding_structure_char, 100)
sims_unif <- tourn_sims(unif_strength,seeding_structure, seeding_structure_char, 100)

graphing <- function(df, distribution_name){
  # Loop over each unique True_Rank
  for (rank in unique(df$True_Rank)){
    # Filter data for the current True_Rank
    team_data <- df %>% filter(True_Rank == rank)

    # Rank the teams based on Total_Wins in descending order, ensuring ties are handled correctly
    best <- team_data %>%
      arrange(desc(Total_Wins)) %>%
      head(25)

    worst <- team_data %>%
      arrange(Total_Wins) %>%
      head(25) %>%
      arrange(desc(Total_Wins))

    # Set Seeding_Structure as a factor to ensure the x-axis ordering by Seeding_Structure
    best$Seeding_Structure <- factor(best$Seeding_Structure, levels = unique(best$Seeding_Structure))
    worst$Seeding_Structure <- factor(worst$Seeding_Structure, levels = unique(worst$Seeding_Structure))

    # Generate the plot for the current True_Rank - Best Plot
    best_plot <- ggplot(best, aes(x = Seeding_Structure, y = Total_Wins, fill = as.factor(True_Rank))) +
      geom_bar(stat = "identity", position = "dodge", width = 0.7) +  
      geom_text(aes(label = Total_Wins), size = 3, vjust = -0.3, color = "black") +
      labs(
        title = paste("Most Total Wins by Seed:", rank, "in Distribution:", distribution_name),
        x = "Seed Number",
        y = "Total Wins",
        fill = "True Rank"
      ) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for readability

    # Print the plot for each True_Rank
    print(best_plot)

    # Generate the plot for the current True_Rank - Worst Plot
    worst_plot <- ggplot(worst, aes(x = Seeding_Structure, y = Total_Wins, fill = as.factor(True_Rank))) +
      geom_bar(stat = "identity", position = "dodge", width = 0.7) + 
      geom_text(aes(label = Total_Wins), size = 3, vjust = -0.3, color = "black") +
      facet_wrap(~ True_Rank, scales = "free_x") +  
      labs(
        title = paste("Least Total Wins by Seed:", rank, "in Distribution:", distribution_name),
        x = "Seed Number",
        y = "Total Wins",
        fill = "True Rank"
      ) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for readability

    # Print the plot for each True_Rank
    print(worst_plot)
  }
}

# Example: Call graphing function for both distributions
graphing(sims_norm, "Normal")
graphing(sims_unif, "Uniform")

```

```{r}
seeding_structure[c(42,86,104,74,41,139,75,117,141,140),]
```

# Individual Spearman Correlations by Seeding Structure

```{r}
spearman_results <- function(df, seeding_structure, replicates = 100) {
  tournament_replicates <- replicate(replicates, {
    # Reset wins and ranks for each tournament
    df$Wins <- rep(0, nrow(df))
    df$Ranks <- rep(NA, nrow(df))  # Ensure Ranks is initialized
    df$True_Rank <- rank(-df$Strength)

    # Run the tournament
    result <- simulate_tournament(df, seeding_structure, ties=FALSE)

    # Calculate Spearman correlation for this replicate
    spearman_cor <- cor(result$Ranks, result$True_Rank, method = "spearman", use = "complete.obs")

    # Return the average Spearman correlation (not the full simulation details)
    return(spearman_cor)
  }, simplify = TRUE)  # Keeps the result as a vector of Spearman correlations

  # Calculate average Spearman correlation for the entire set of replicates
  average_spearman <- mean(tournament_replicates, na.rm = TRUE)

  return(average_spearman)
}

output_spearman <- function(strengths_list, seeding_structure, seeding_structure_char = NULL) {
  # Preallocate results list
  results <- vector("list", length = nrow(seeding_structure) * length(strengths_list))
  counter <- 1  # Index for results

  # Loop over each seed configuration
  for (seed_index in 1:nrow(seeding_structure)) {
    current_seed <- matrix(seeding_structure[seed_index, ], nrow = 1)
    current_structure <- if (!is.null(seeding_structure_char)) {
      seeding_structure_char[seed_index]
    } else {
      paste(seeding_structure[seed_index, ], collapse = "-")
    }

    # Loop over each distribution type
    for (dist_name in names(strengths_list)) {
      df <- strengths_list[[dist_name]]

      # Calculate average Spearman correlation
      avg_spearman <- spearman_results(df, current_seed)

      # Store results
      results[[counter]] <- data.frame(
        Seed = seed_index,
        Seeding_Structure = current_structure,
        Distribution = dist_name,
        Average_Spearman = avg_spearman
      )
      counter <- counter + 1
    }
  }

  # Combine all results into a single data frame
  final_results <- do.call(rbind, results)
  return(final_results)
}

# Example usage
#output_results(strengths_list, seeds = seeds)

```

# Average Spearman Correlation by Seeding Structure

```{r}
spearman_by_seed <- function(df, seeding_structure, seeding_structure_char, replicates = 100) {
  tournament_replicates <- replicate(replicates, {
    # Reset wins and ranks for each tournament
    df$Wins <- rep(0, nrow(df))
    df$Ranks <- rep(NA, nrow(df))  # Ensure Ranks is initialized
    df$True_Rank <- rank(-df$Strength)

    # Run the tournament
    result <- simulate_tournament(df, seeding_structure,ties=FALSE)

    # Calculate Spearman correlation for this replicate
    spearman_cor <- cor(result$Ranks, result$True_Rank, method = "spearman", use = "complete.obs")

    # Return the average Spearman correlation (not the full simulation details)
    return(spearman_cor)
  }, simplify = TRUE)  # Keeps the result as a vector of Spearman correlations

  return(tournament_replicates)
}

output_results <- function(strengths_list, seeding_structure, seeding_structure_char = NULL) {
  # Preallocate results list
  results <- vector("list", length = nrow(seeding_structure) * length(strengths_list))
  counter <- 1  # Index for results

  # Loop over each seed configuration
  for (seed_index in 1:nrow(seeding_structure)) {
    current_seed <- matrix(seeding_structure[seed_index, ], nrow = 1)
    current_structure <- if (!is.null(seeding_structure_char)) {
      seeding_structure_char[seed_index]
    } else {
      paste(seeding_structure[seed_index, ], collapse = "-")
    }

    # Loop over each distribution type
    for (dist_name in names(strengths_list)) {
      df <- strengths_list[[dist_name]]

      # Calculate average Spearman correlation
      avg_spearman <- spearman_by_seed(df, current_seed)

      # Store results
      results[[counter]] <- data.frame(
        Seed = seed_index,
        Seeding_Structure = current_structure,
        Distribution = dist_name,
        Spearman = avg_spearman
      )
      counter <- counter + 1
    }
  }

  # Combine all results into a single data frame
  final_results <- do.call(rbind, results)
  return(final_results)
}

# Example usage
#output_results(strengths_list, seeds = seeds)
```

# Spearman Correlation Best and Worst Cases

```{r}
graph_results <- output_results(strengths_list, seeding_structure = seeding_structure, seeding_structure_char) 
library(ggplot2)
#ggplot(data=graph_results, aes(x=as.factor(Seeding_Structure),y=Spearman)) + geom_boxplot() + facet_grid(~Distribution)

#ggplot(data=graph_results, aes(x=Spearman,fill=as.factor(Seeding_Structure))) + geom_density(alpha=0.5) + facet_grid(~Distribution)

sim_results <- output_spearman(strengths_list, seeding_structure = seeding_structure, seeding_structure_char) 
#ggplot(data=sim_results, aes(x=as.factor(Seeding_Structure),y=Average_Spearman)) + geom_bar(stat="identity", position = "dodge") + facet_grid(~Distribution)
# Sort and extract top 5 highest Spearman correlations
norm_top_5_seeds <- sim_results %>%
  filter(Distribution == "Normal") %>%
  arrange(desc(Average_Spearman)) %>%
  slice_head(n = 5)  # Top 5 highest Spearman correlations

unif_top_5_seeds <- sim_results %>%
  filter(Distribution == "Uniform") %>%
  arrange(desc(Average_Spearman)) %>%
  slice_head(n = 5)  # Top 5 highest Spearman correlations

# Sort and extract top 5 lowest Spearman correlations
norm_worst_5_seeds <- sim_results %>%
  filter(Distribution == "Normal") %>%
  arrange(Average_Spearman) %>%
  slice_head(n = 5)  # Top 5 lowest Spearman correlations

unif_worst_5_seeds <- sim_results %>%
  filter(Distribution == "Uniform") %>%
  arrange(Average_Spearman) %>%
  slice_head(n = 5)  # Top 5 highest Spearman correlations

print(norm_top_5_seeds)
print(norm_worst_5_seeds)
print(unif_top_5_seeds)
print(unif_worst_5_seeds)
```

# Probability of Top 3 Correct

```{r}
# Function to run multiple simulations
run_simulations <- function(df, seeding_structure = NULL, seed_char = NULL, replicates = 1000, series = 1,ties=FALSE) {
  # Default seeding structure
  if (is.null(seeding_structure)) {
    seeding_structure <- matrix(c(1, 8, 4, 5, 2, 7, 3, 6), nrow = 1)
  }

  # Store results for all simulations
  all_simulations <- lapply(1:replicates, function(rep) {
    # Simulate tournament for each seeding structure
    lapply(1:nrow(seeding_structure), function(seed_num) {
      simulate_tournament(df,seeding_structure,series = series,third_place = TRUE
      ,ties=FALSE) %>%
        mutate(Simulation = rep, Seed_Num = seed_num, Seeding_Structure = seed_char[seed_num])
    }) %>%
      do.call(rbind, .)
  }) %>%
    do.call(rbind, .)

  return(all_simulations)
}

# Analyze Top-Three Ranks
analyze_top_three <- function(simulations) {

  # Identify top-3 ranks for each simulation
  top_three_results <- simulations %>%
    group_by(Simulation) %>%
    filter(Ranks %in% 1:3 & True_Rank %in% 1:3) %>%
    summarize(match_count = n_distinct(True_Rank), .groups = 'drop') %>%
    filter(match_count == 3)

  top_three_prob = nrow(top_three_results) / n_distinct(simulations$Simulation)
  return(top_three_prob)
}

simulations <- run_simulations(normal_strengths, replicates = 1000, series = 3, ties=FALSE)

# Analyze top-three occurrences
top_three_summary <- analyze_top_three(simulations)

# Print results
print(top_three_summary)

```

-   Probability that the top 3 are correct (1,2,3) in any order

-   16 team seeding

-   byes and third team game

-   curve showing top i seeds correct (for 2 is 1 and 2 in the final); overlap a curve showing the correct order; probability of 1, 2\|1, 3\|(1,2) or (2,1) etc; use %in%

-   series best of n

    ```{r}
    curve <- function(sim){
      probability <- numeric(num_teams)
      for (i in 1:num_teams){
        if (i==1){
          probability[i] <- nrow(sim[sim$Ranks == 1 & sim$True_Rank == 1, ]) / length(unique(sim$Simulation))
        }
        else{
        top_ranks <- sim[sim$Ranks <= (i-1) & sim$True_Rank <= (i-1), ]
        better_teams <- 1:(i-1)
        numerator <- sum(sapply(unique(top_ranks$Simulation), function(sim_id){
          sim_subset <- sim[sim$Simulation == sim_id, ]

          if(all(better_teams %in% sim_subset$True_Rank[sim_subset$Ranks <= (i - 1)])){
          return(sum(sim_subset$Ranks == i & sim_subset$True_Rank == i))
          }
          else{
            return(0)
          }
          }))

        denominator <- sum(sapply(unique(top_ranks$Simulation), function(sim_id){
            sim_subset <- top_ranks[top_ranks$Simulation == sim_id, ]
            all(better_teams %in% sim_subset$True_Rank)
          }))
        #print(i)
        #print(numerator)
        #print(denominator)
        probability[i] <- numerator / denominator
        }
      }
      return(probability)
    }
    simulations <- run_simulations(normal_strengths, replicates = 2000, series = 7, ties=FALSE)
    sim_probs <- curve(simulations)
    probabilities <- data.frame(x=c(1:num_teams), y=sim_probs)
    probabilities$Line <- "Simulated"

    baseline_prob <- 1/(num_teams:1)
    baseline <- data.frame(x=c(1:num_teams), y=baseline_prob)
    baseline$Line <- "Baseline"

    prob_plot <- rbind(baseline,probabilities)

    ggplot(data=prob_plot, aes(x=x,y=y, color=Line, group=Line)) + geom_point() + geom_line() + ylim(0,1) + labs(
      x = "Rank",
      y = "Conditional Probability"
    )

    # Interesting to note that 3 is almost always greater than 2
    ```