### 2025-02-20 swaps.qmd

---
title: "Swaps"
author: "Zach Culp"
format: html
editor: visual
---

## Number of Swaps to Get Correct Seeding

Notation: PS = number of positions away from the true rank that the predicted rank is

For example, if $R_1 = 3$, then $PS_1 = -2$. PS follows a discrete probability distribution with range of $-n+1 < PS < n-1$, where $n$ is the number of teams in the tournament. My prediction is that E(PS) should be 0 in a fair tournament. Then, if we see multiple tournaments with a predicted PS around 0, the variance can be a determining factor to how well the tournament is structures.

```{r}
library(tidyverse)
se_norm_results <- all_results_norm %>% filter(teams == 8)

se_norm_results$simulation <- rep(1:10000, each = nrow(curves_test) / 10000)

create_ps <- function(df){
  df <- df %>%
  mutate(ps = true_rank - rank_hat) %>%
  group_by(true_rank) %>%
  summarize(

    mean_ps = mean(ps),
    sd_ps = sd(ps)
  )
  return(df)
}
se_test <- create_ps(se_norm_results)
de_test <- create_ps(de_norm_results %>% filter(teams == 8))
rr_test <- create_ps(test_rr_noTies)

library(combinat)
perms_list <- permn(8)
all_data <- list()
# Loop through each permutation and create the data frame
for (i in 1:length(perms_list)) {
  # Create a temporary data frame for the current permutation
  temp_df <- data.frame(
    true_rank = rep(1:8, each = 1),  # Each row gets a true_rank from 1 to 8
    rank_hat = perms_list[[i]],      # The current permutation's ranks
    simulation = rep(i, 8)           # Assign the simulation number to each row
  )
  # Add the temporary data frame to the list
  all_data[[i]] <- temp_df
}
# Combine all the data frames into one
perms_df <- do.call(rbind, all_data)
perms_test <- create_ps(perms_df)
```

### Graph

```{r}
# Combine the two datasets into one for easier plotting
se_test <- se_test %>%
  mutate(type = "Single Elimination")  # Add a label for the type of data

perms_test <- perms_test %>%
  mutate(type = "Baseline")   # Add a label for the type of data

rr_test <- rr_test %>%
  mutate(type = "Round Robin")

de_test <- de_test %>%
  mutate(type = "Double Elimination")

# Combine both data frames into one
combined_swaps_data <- bind_rows(se_test, perms_test, rr_test, de_test)

# Plotting using ggplot2
ggplot(combined_swaps_data, aes(x = true_rank, y = mean_ps, color = type, group = type)) +
  geom_point(size = 3) +  # Plot the points
  geom_line() +  # Connect the points with lines
  geom_errorbar(aes(ymin = mean_ps - sd_ps, ymax = mean_ps + sd_ps), width = 0.2) +  # Add error bars (standard deviation)
  scale_color_manual(values = c("Single Elimination" = "blue", "Baseline" = "red", "Round Robin" = "green", "Double Elimination" = "yellow")) +  # Set colors for the groups
  labs(
    title = "Comparison of Mean PS with Standard Deviation for Simulated vs Baseline",
    x = "True Rank",
    y = "Mean PS",
    color = "Simulation Type"
  ) +
  theme_minimal() + # Apply a clean theme
  theme(
      panel.background = element_rect(fill = "gray75"),  # Set background color to light gray
      plot.background = element_rect(fill = "gray80"),  # Set the plot background color to a lighter gray
      legend.background = element_rect(fill = "gray80")  # Set the legend background color
    )
```

### Baseline Probabilities of Under/Over Performance

An overall graph of the probabilities of correctly predicting the rankings versus overpredicting and underpredicting. If the tournament results were randomly decided the probabilities would be the following:

$P(R_i = i) = \frac{1}{n}$

$P(R_i < i) = \frac{i-1}{n}$

$P(R_i > i) = \frac{n-i}{n}$

```{r}
# Function to calculate probabilities of ps conditions
calculate_probs <- function(df) {
  df %>%
    mutate(ps = true_rank - rank_hat) %>%
    group_by(true_rank) %>%
    summarize(
      ps_equal = sum(ps == 0) / n(),
      ps_less = sum(ps < 0) / n(),
      ps_more = sum(ps > 0) / n()
    )
}

# Apply the function to each data frame
probs_rr_test <- calculate_probs(test_rr_noTies)
probs_se_test <- calculate_probs(se_norm_results)
probs_perms_test <- calculate_probs(perms_df)
probs_de_test <- calculate_probs(de_norm_results_eight)

# Combine the two datasets into one for easier plotting
probs_se_test <- probs_se_test %>%
  mutate(type = "Single Elimination")  # Add a label for the type of data

probs_perms_test <- probs_perms_test %>%
  mutate(type = "Baseline")   # Add a label for the type of data

probs_rr_test <- probs_rr_test %>%
  mutate(type = "Round Robin")

probs_de_test <- probs_de_test %>%
  mutate(type = "Double Elimination")

# Combining datasets
combined_probs_swaps_data <- bind_rows(probs_rr_test, probs_se_test, probs_perms_test, probs_de_test)

# R_i < i plot
ggplot(combined_probs_swaps_data, aes(x = true_rank, y = ps_less, color = type, group = type)) +
  geom_point(size = 3) +  
  geom_line() +
  scale_color_manual(values = c("Single Elimination" = "blue", 
                                "Baseline" = "red", 
                                "Round Robin" = "green",
                                "Double Elimination" = "yellow")) +  
  labs(
    title = "Probability of Simulation Under Predicting Ranks for Each Tournament",
    x = "True Rank",
    y = expression(P(R[i] < i)),
    color = "Simulation Type"
  ) +
  theme_minimal() +  # Apply a clean theme
  theme(
    panel.background = element_rect(fill = "gray75"),  
    plot.background = element_rect(fill = "gray80"), 
    legend.background = element_rect(fill = "gray80")  
  ) +
  scale_y_continuous(limits = c(0, 1))

# R_i = i plot
ggplot(combined_probs_swaps_data, aes(x = true_rank, y = ps_equal, color = type, group = type)) +
  geom_point(size = 3) +  
  geom_line() +
  scale_color_manual(values = c("Single Elimination" = "blue", 
                                "Baseline" = "red", 
                                "Round Robin" = "green",
                                "Double Elimination" = "yellow")) +  
  labs(
    title = "Probability of Simulation Correctly Predicting Ranks for Each Tournament",
    x = "True Rank",
    y = expression(P(R[i] == i)),
    color = "Simulation Type"
  ) +
  theme_minimal() +  # Apply a clean theme
  theme(
    panel.background = element_rect(fill = "gray75"),  
    plot.background = element_rect(fill = "gray80"), 
    legend.background = element_rect(fill = "gray80")  
  ) +
  scale_y_continuous(limits = c(0, 1))

# R_i > i plot
ggplot(combined_probs_swaps_data, aes(x = true_rank, y = ps_more, color = type, group = type)) +
  geom_point(size = 3) +  
  geom_line() +
  scale_color_manual(values = c("Single Elimination" = "blue", 
                                "Baseline" = "red", 
                                "Round Robin" = "green",
                                "Double Elimination" = "yellow")) +  
  labs(
    title = "Probability of Simulation Over Predicting Ranks for Each Tournament",
    x = "True Rank",
    y = expression(P(R[i] > i)),
    color = "Simulation Type"
  ) +
  theme_minimal() +  # Apply a clean theme
  theme(
    panel.background = element_rect(fill = "gray75"),  
    plot.background = element_rect(fill = "gray80"), 
    legend.background = element_rect(fill = "gray80")  
  ) +
  scale_y_continuous(limits = c(0, 1))

```

\

\