### 2025-04-21 mutual_information_curves.qmd

---
title: "Mutual Information Plots"
format: html
editor: visual
---

## Mutual Information Plots

From Wikipedia, weight mutual information formula:

$I(X;Y)=\sum _{y\in Y}\sum _{x\in X}w(x,y)p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}$

```{r}
library(combinat)
library(tidyverse)

mut_info <- function(sims, num_teams, rank_weights = NULL){
  # Handle missing weights (default to equal weights if not provided)
  if (is.null(rank_weights)){
    rank_weights <- rep(1,num_teams)
  }
  # Extract y values from sims, assuming `rank_hat` contains a list of rankings per simulation
  y_df <- sims %>%
    group_by(simulation) %>%
    summarise(rank_hat = list(rank_hat)) %>%  # Ensure rank_hat is a list column
    mutate(rank_hat = map(rank_hat, unlist)) %>%  # Unlist rank_hat properly
    unnest_wider(rank_hat, names_sep = "_")  # Expand into separate columns

  # Rename y columns
  colnames(y_df)[-1] <- paste0("y", 1:num_teams)

  # Create x matrix, representing positions 1:num_teams
  x_matrix <- matrix(rep(1:num_teams, times = nrow(y_df)), ncol = num_teams, byrow = TRUE)
  x_df <- as.data.frame(x_matrix)
  colnames(x_df) <- paste0("x", 1:num_teams)

  # Combine x and y data frames
  perms_df <- cbind(x_df, y_df %>% select(-simulation))
  perms_df <- perms_df %>%
    mutate(across(starts_with("x"), as.numeric)) %>%
    mutate(across(starts_with("y"), as.numeric))

  # Generate all possible permutations of rankings
  all_perms <- as.data.frame(do.call(rbind, permn(1:num_teams)))
  colnames(all_perms) <- paste0("y", 1:num_teams)

  # Count occurrences of each observed ranking
  prob_df <- perms_df %>%
    group_by(across(starts_with("x")),across(starts_with("y")))  %>%
    summarise(count = n(), .groups = "drop")

  # Merge with full permutation set to ensure all permutations exist
  prob_df <- right_join(all_perms, prob_df, by = colnames(all_perms)) %>%
    mutate(count = replace_na(count, 0)) %>%
    mutate(probability = count / sum(count))  # Normalize probabilities

    w <- numeric(nrow(prob_df))  # Initialize weight vector for the number of rows in prob_df

  for (i in 1:num_teams) {
    x_i <- paste0("x", i)
    y_i <- paste0("y", i)

    # Loop through each row of prob_df
    for (j in 1:nrow(prob_df)) {
      # Calculate the squared difference between x_i and y_i for each row j
      diff <- abs(prob_df[[x_i]][j] - prob_df[[y_i]][j])
    w[j] <- w[j] + rank_weights[i]* diff^2  # Sum squared differences for each row
    }
  }
  w <- 1 / (w + 1)  # Inverse of squared differences and manually change 0 probabilities to 1 on the inverse
  w <- w/sum(w)
  # Compute weights dynamically for each row
  weighted_df <- prob_df %>%
    mutate(w = w)

  mutual_info <- factorial(num_teams) * sum(weighted_df$w * weighted_df$probability * log2(weighted_df$probability / (1/factorial(num_teams) * weighted_df$probability)))

  # Return the weighted data frame with mutual information weights
  return(mutual_info)
}
mut_info(all_results_norm,8, c(10,5,1,0,0,0,0,0))
mut_info(de_norm_results_eight,8, c(10,5,1,0,0,0,0,0))
mut_info(test_rr_noTies,8, c(10,5,1,0,0,0,0,0))
mut_info(seven_games_se_results,8, c(10,5,1,0,0,0,0,0))
mut_info(perms_df,8,c(1,0,0,0,0,0,0,0))

de_norm_results_eight <- de_norm_results %>% filter(teams == 8)
de_norm_results_eight$simulation <- rep(1:10000, each = nrow(de_norm_results_eight) / 10000)
test_unif$simulation <- rep(1:10000, each = nrow(test_unif) / 10000)
equal_strengths$simulation <- rep(1:10000, each = nrow(equal_strengths) / 10000)
all_results_norm$simulation <- rep(1:10000, each = nrow(all_results_norm) / 10000)
bad_seeding_structure$simulation <- rep(1:10000, each = nrow(bad_seeding_structure) / 10000)
seven_games_se_results$simulation <- rep(1:10000, each = nrow(seven_games_se_results) / 10000)
# with randomly assigning ties
test_rr_noTies <- test_rr %>%
  group_by(simulation) %>%
  mutate(rank_hat = rank(rank_hat,ties.method="random"))

# Initialize variables
num_vars <- 8  # Total length of the vector
indices <- 1:num_vars  # Store the number of 1's in each step

# Function to compute mutual information for a given dataset
compute_mut_plots <- function(dataset) {
  results <- numeric(num_vars)
  for (i in 1:num_vars) {
    binary_vector <- c(rep(1, i), rep(0, num_vars - i))  # Construct vector
    results[i] <- mut_info(dataset, num_vars, binary_vector)  # Compute MI
  }
  return(results)
}

perms_list <- permn(8)
all_data <- list()
# Loop through each permutation and create the data frame
for (i in 1:length(perms_list)) {
  # Create a temporary data frame for the current permutation
  temp_df <- data.frame(
    true_rank = rep(1:8, each = 1),  # Each row gets a true_rank from 1 to 8
    rank_hat = perms_list[[i]],      # The current permutation's ranks
    simulation = rep(i, 8)           # Assign the simulation number to each row
  )
  # Add the temporary data frame to the list
  all_data[[i]] <- temp_df
}
# Combine all the data frames into one
perms_df <- do.call(rbind, all_data)

# Compute mutual information for each dataset
mi_de_norm <- compute_mut_plots(de_norm_results_eight)
mi_se_norm <- compute_mut_plots(all_results_norm)
mi_test_rr <- compute_mut_plots(test_rr_noTies)
mi_seven_se <- compute_mut_plots(seven_games_se_results)
mi_random <- compute_mut_plots(equal_strengths)
mi_se_unif <- compute_mut_plots(test_unif)
mi_bad <- compute_mut_plots(bad_seeding_structure)
mi_perms <- compute_mut_plots(perms_df)

# Plot results
plot(indices, mi_de_norm, type="b", pch=19, col="blue", ylim=range(c(mi_de_norm, mi_se_norm, mi_test_rr, mi_seven_se, mi_random, mi_se_unif)),
     xlab="Number of 1's in the vector", ylab="Mutual Information Bits", main="Mutual Information vs. Number of 1's")

# Add other datasets
lines(indices, mi_se_norm, type="b", pch=19, col="red")
lines(indices, mi_test_rr, type="b", pch=19, col="green")
lines(indices, mi_seven_se, type = "b", pch=19, col="black")
lines(indices, mi_random, type="b", pch=19, col="purple")
lines(indices, mi_se_unif, type="b", pch=19, col="orange")

# Add legend
legend("topleft", legend=c("Double Elimination", "Single Elimination", "Round Robin", "7 Game Single Elimination", "Random Single Elimination", "Uniform Single Elimination"), 
       col=c("blue", "red", "green", "black", "purple", "orange"), pch=19, lty=1)

mi_plots_df <- data.frame(
  indices = indices,
  DE = mi_de_norm,
  Norm_SE = mi_se_norm,
  Random_SE = mi_random,
  Unif_SE = mi_se_unif,
  Bad_Seeds_SE = mi_bad,
  Perms = mi_perms
)
# Reshape data into long format for ggplot
df_long <- pivot_longer(mi_plots_df, cols = -indices, names_to = "Method", values_to = "MI")
ggplot(df_long, aes(x = indices, y = MI, color = Method, group = Method)) +
  geom_line() +           # Add lines
  geom_point() +          # Add points
  labs(
    title = "Single Elimination vs Double Elimination",
    x = "Number of 1's in the vector",
    y = "Mutual Information Bits"
  ) +
  theme_minimal() +       # Use a clean theme
  theme(legend.position = "right")
```