### 2025-05-01 double_elimination.qmd
update 5/1/25

---
title: "Double Elimination Tournament"
author: "Zach Culp"
format: html
editor: visual
---

# Double Elimination Tournament

```{r}
library(tidyverse)
library(dplyr)
library(shiny)
```

## Simulate a Single Match

```{r}
# Function to simulate a match using Bradley-Terry model
simulate_match <- function(team1, team2, strength1, strength2) {
  if(is.na(team1) && is.na(team2)){
    return(NA)
  }
  if (is.na(team1)) {
    return(team2)
  } else if (is.na(team2)) {
    return(team1)
  } else {
    p <- exp(strength1 - strength2) / (1 + exp(strength1 - strength2))
    return(ifelse(runif(1) < p, team1, team2))  # Randomize outcome
  }
}
```

## Creating the traditional tournament structure for n teams

```{r}
generate_elimination_tournament_structure <- function(n) {

  is_power_of_two <- function(x) {
    x > 0 && (x & (x - 1)) == 0
  }

  # If n is not a power of 2, round up to the next power of 2 and add placeholders
  if (!is_power_of_two(n)) {
    next_power_of_two <- 2^ceiling(log2(n))
    extra_teams <- next_power_of_two - n
    #message(sprintf("Rounding up to %d teams. Adding %d NA placeholders.", next_power_of_two, extra_teams))
    n <- next_power_of_two
  } else {
    extra_teams <- 0
  }

  rounds <- log(n, base = 2) - 1
  teams <- c(1, 2)  # Initial list of teams

  for (i in 1:rounds) {
    teams <- nextLayer(teams)
  }

  left_half <- teams[1:(length(teams)/2)]
  right_half <- teams[(length(teams)/2+1):length(teams)]
  right_half <- rev(right_half)
  teams <- c(left_half,right_half)

  # Ensure that the lower number team is on the left side of each matchup
  teams <- unlist(lapply(seq(1, length(teams), by = 2), function(i) {
    if (teams[i] > teams[i + 1]) {
      c(teams[i + 1], teams[i])  # Swap if needed
    } else {
      c(teams[i], teams[i + 1])
    }
  })
  )

  while (extra_teams > 0){
      teams[which.max(teams)] <- NA
      extra_teams <- extra_teams - 1
  }

  return(teams)
}

nextLayer <- function(teams) {
  out <- c()  # Initialize an empty vector
  length <- length(teams) * 2 + 1  # Calculate the length for pairing

  # Generate the next layer by pairing teams
  for (i in teams) {
    out <- c(out, i, length - i)  # Push the current team and its pair
  }

  return(out)
}

```

## Creating a Single Simulation of a Double Elimination Tournament

```{r}
simulate_double_elimination_tournament <- function(num_teams,distribution,seeding_structure = NULL, ties=T, series=1) {

  if (num_teams <= 3) {
    stop("Number of Teams must be greater than 3.")
  }
  teams <- paste(num_teams:1)
  if (log2(num_teams) %% 1 != 0) {
    power_of_2 <- 2^ceiling(log2(num_teams))
    teams <- c(teams, rep(NA, power_of_2 - num_teams))
  }
  if (distribution == "Normal"){
    strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA,length(teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  }
  else if (distribution == "Uniform"){
    strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1) , 0 , sqrt(12)) })
    unif_strength <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams)),
      rank_hat = rep(NA,length(teams))
    )
    df <- arrange(unif_strength, true_rank)
  }
     else if (distribution == "Manual") {
    cat("Please enter the strengths for each team in descending rank order:\n")
    strengths <- numeric(num_teams)
    for (i in num_teams:1) {
      repeat {
        input <- as.numeric(readline(paste("Strength for Team", num_teams - i + 1, ": ")))
        if (!is.na(input)) {
          strengths[i] <- input
          break
        } else {
          cat("Invalid input. Please enter a numeric value.\n")
        }
      }
    }
    manual_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- arrange(manual_strengths, true_rank)
  }
  else {
    stop("Distribution not found: Enter Manual to input your own strengths")
  }

  # Sets seed to that of common best vs worst structure
  if (is.null(seeding_structure)){
    seeding_structure <- matrix(generate_elimination_tournament_structure(num_teams), nrow = 1)
  } else{
    seeding_structure <- matrix(seeding_structure, nrow = 1)
  }
  if(series %% 2 == 0) {
    stop("Series must be an odd number to avoid potential ties.")
  }

  results <- list()  # Initialize the results list

  for (i in 1:nrow(seeding_structure)) {
    # Create a new data frame based on the current permutation in 'test'
    permutation <- seeding_structure[i, ]
    permuted_df <- df
    permuted_df$true_rank <- df$true_rank[permutation]  # Update team order 
    teams <- permuted_df$true_rank
    permuted_df$true_strength <- df$true_strength[permutation]
    permuted_df$rank_hat <- NA  # Initialize ranks column
    permuted_df$game_wins <- 0  # Initialize wins column
    permuted_df$game_losses <- 0

    team_rank <- num_teams
    winners_bracket <- teams
    strengths <- permuted_df$true_strength

    losers_bracket <- c()

    # gets results of the winners bracket
    while (length(winners_bracket) > 1 || length(losers_bracket > 0)) {
      eliminated <- c()
      if (length(winners_bracket) > 1) {
      #cat("\n--- Winner's Bracket --- \n")
      next_winners <- c() 
      round_losers <- c()

      for (j in seq(1, length(winners_bracket), by = 2)) { # Gets 2 teams for each match
        if (j + 1 > length(winners_bracket)) break
        team1 <- winners_bracket[j]
        team2 <- winners_bracket[j+1]

        strength1 <- strengths[which(permuted_df$true_rank == team1)]
        strength2 <- strengths[which(permuted_df$true_rank == team2)]

        wins_team1 <- 0
        wins_team2 <- 0
        losses_team1 <- 0
        losses_team2 <- 0
        for (game in 1:series){
          match_winner <- simulate_match(team1, team2, strength1, strength2)
          cat(sprintf("Matchup:  %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", team1, team2, strength1, strength2, match_winner))
          if(!is.na(team1)) {
          if (match_winner == team1) {
            wins_team1 <- wins_team1 + 1
            losses_team2 <- losses_team2 + 1
          } else {
            wins_team2 <- wins_team2 + 1
            losses_team1 <- losses_team1 + 1
          }
          }
          else {wins_team2 <- (series+1) / 2}

          # Determine the overall winner if a majority is reached
          if (wins_team1 > series / 2 || wins_team2 > series / 2) {
            break
          }
        }
        # Update the game_wins and game_losses outside of the loop
        if (!any(is.na(c(team1, team2)))) {
          cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
        permuted_df <- permuted_df %>%
          mutate(game_wins = case_when(
            true_rank == team1 ~ game_wins + wins_team1,
            true_rank == team2 ~ game_wins + wins_team2,
            TRUE ~ game_wins
          )) %>%
          mutate(game_losses = case_when(
            true_rank == team1 ~ game_losses + losses_team1,
            true_rank == team2 ~ game_losses + losses_team2,
            TRUE ~ game_losses
          ))
        }
        # Determine the match winner based on series results
        series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
        series_loser <- ifelse(series_winner == team1, team2, team1)
        cat(sprintf("Matchup:  %s vs %s | Winner: %s\n: | Loser: %s\n", team1, team2, series_winner, series_loser))
        next_winners <- c(next_winners, series_winner)
        round_losers <- c(round_losers, series_loser)
    }
     winners_bracket <- next_winners  # Winners move to the next round
     losers_bracket <- c(losers_bracket, round_losers)
     if(length(losers_bracket) %% 2 != 0){losers_bracket <- c(losers_bracket, NA)}

      }

    # Losers' bracket
    if (length(losers_bracket) > 0) {
      if(length(losers_bracket) > 1){
      cat("\n--- Loser's Bracket --- \n")
      }
      next_losers <- c()

      # alternates games for losers bracket
      if(length(losers_bracket) > 2){
        if(length(losers_bracket) == 4 && is.na(losers_bracket[4])){
          losers_bracket <- losers_bracket
        } else{
          top_bracket <- losers_bracket[seq(1,length(losers_bracket), by = 2)]
          bottom_bracket <- losers_bracket[seq(2,length(losers_bracket), by = 2)]
          losers_bracket <- c(top_bracket,bottom_bracket)
        }
      }

      for (j in seq(1, length(losers_bracket), by = 2)) {
        if (j + 1 > length(losers_bracket)) break
        team1 <- losers_bracket[j]
        team2 <- losers_bracket[j + 1]
        strength1 <- strengths[which(permuted_df$true_rank == team1)]
        strength2 <- strengths[which(permuted_df$true_rank == team2)]

        wins_team1 <- 0
        wins_team2 <- 0
        losses_team1 <- 0
        losses_team2 <- 0

        # Play series
        for (game in 1:series) {
          match_winner <- simulate_match(team1, team2, strength1, strength2)
          if(is.na(match_winner)){
            wins_team1 == wins_team1 + 1
          }
          if(!is.na(team1)) {
            if (match_winner == team1) {
              wins_team1 <- wins_team1 + 1
              losses_team2 <- losses_team2 + 1
            } else {
              wins_team2 <- wins_team2 + 1
              losses_team1 <- losses_team1 + 1
            }
          }
          else {wins_team2 <- (series+1) / 2}

          # Determine overall series winner
          if (wins_team1 > series / 2 || wins_team2 > series / 2) {
            break
          }
        }
        if (!any(is.na(c(team1, team2)))) {
          cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
          permuted_df <- permuted_df %>%
          mutate(game_wins = case_when(
            true_rank == team1 ~ game_wins + wins_team1,
            true_rank == team2 ~ game_wins + wins_team2,
            TRUE ~ game_wins
          )) %>%
          mutate(game_losses = case_when(
            true_rank == team1 ~ game_losses + losses_team1,
            true_rank == team2 ~ game_losses + losses_team2,
            TRUE ~ game_losses
          ))
        }

        if(is.na(team1) && is.na(team2)){
          series_winner <- NA
          next_losers <- c(next_losers, NA)
        } else{
        series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
        series_loser <- ifelse(wins_team1 > wins_team2, team2, team1)
        next_losers <- c(next_losers, series_winner)
        eliminated <- c(eliminated, series_loser)
        cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, series_winner))
        }
      }
        team_rank <- team_rank - 1
        permuted_df$rank_hat[permuted_df$true_rank %in% eliminated] <- team_rank
        losers_bracket <- next_losers

    }

      # Final match for loser's bracket (aka third place game)
if (length(winners_bracket) == 1 && length(losers_bracket) == 1) {

# Final matchup between winners and losers
    cat("\n--- Final Match ---\n")
    final_match <- c(winners_bracket, losers_bracket)
    if(length(final_match) == 2){
    team1 <- final_match[1]
    team2 <- final_match[2]
    strength1 <- strengths[which(permuted_df$true_rank == team1)]
    strength2 <- strengths[which(permuted_df$true_rank == team2)]

    wins_team1 <- 0
    wins_team2 <- 0
    losses_team1 <- 0
    losses_team2 <- 0

    for (game in 1:series) {
      match_winner <- simulate_match(team1, team2, strength1, strength2)
      if (match_winner == team1) {
        wins_team1 <- wins_team1 + 1
        losses_team2 <- losses_team2 + 1
      } else {
        wins_team2 <- wins_team2 + 1
        losses_team1 <- losses_team1 + 1
      }

      if (wins_team1 > series / 2 || wins_team2 > series / 2) {
        break
      }
    }
    if (!any(is.na(c(team1, team2)))) {
      cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
      permuted_df <- permuted_df %>%
        mutate(game_wins = case_when(
          true_rank == team1 ~ game_wins + wins_team1,
          true_rank == team2 ~ game_wins + wins_team2,
          TRUE ~ game_wins
        )) %>%
        mutate(game_losses = case_when(
          true_rank == team1 ~ game_losses + losses_team1,
          true_rank == team2 ~ game_losses + losses_team2,
          TRUE ~ game_losses
          ))
    }

    if (wins_team1 > wins_team2) {
      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team1))
      permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
      permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
    } else {
      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team2))
      # second game if the loser's bracket team wins
      for (game in 1:series) {
      match_winner <- simulate_match(team1, team2, strength1, strength2)
      if (match_winner == team1) {
        wins_team1 <- wins_team1 + 1
        losses_team2 <- losses_team2 + 1
      } else {
        wins_team2 <- wins_team2 + 1
        losses_team1 <- losses_team1 + 1
      }

      if (wins_team1 > series / 2 || wins_team2 > series / 2) {
        break
      }
    }
      if (!any(is.na(c(team1, team2)))) {
        cat(sprintf("Wins - %s: %d | %s: %d\n", team1, wins_team1, team2, wins_team2))
        permuted_df <- permuted_df %>%
          mutate(game_wins = case_when(
            true_rank == team1 ~ game_wins + wins_team1,
            true_rank == team2 ~ game_wins + wins_team2 - 1, # wins_team2 = 2 in loop but only need to add one game
            TRUE ~ game_wins
          )) %>%
          mutate(game_losses = case_when(
            true_rank == team1 ~ game_losses + losses_team1 - 1,
            true_rank == team2 ~ game_losses + losses_team2,
            TRUE ~ game_losses
          ))
      }
      if (wins_team1 > wins_team2) {
      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team1))
      permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
      permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
    } else {
      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team2))
      permuted_df$rank_hat[permuted_df$true_rank == team2] <- 1
      permuted_df$rank_hat[permuted_df$true_rank == team1] <- 2
        }
      }
    }
  }
}
    permuted_df$seed <- rep(paste(seeding_structure[i, ], collapse = ", "), nrow(permuted_df))
    results[[i]] <- permuted_df
}
  final_results <- do.call(rbind, results)

  # Remove ties if ties = FALSE
  if (!ties) {
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(true_rank) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "random"))
  }
  else{
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(rank_hat) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "average"))
  }

  final_results <- final_results %>%
    mutate(distribution = distribution) %>%
    mutate(true_rank = as.numeric(true_rank)) %>%
    arrange(true_rank)
  return(final_results)
}

test <- simulate_double_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1)
#test
#manual_test <- simulate_double_elimination_tournament(num_teams = 4, distribution = "Manual", ties=F, series=1)
#manual_test

# Gets all results of a normal distribution num_teams between 4 and 64 single-game elimination
de_norm_results <- list()
for (teams in c(4,8,16,32,64)) {
  de_norm_results[[as.character(teams)]] <- list()
  for (replicate in 1:10000) {
    # Store each simulation result in the list with unique indexing
    de_norm_results[[as.character(teams)]][[replicate]] <- simulate_double_elimination_tournament(
      num_teams = teams,
      distribution = "Normal",
      ties = FALSE,
      series = 1
    )
  }
}

de_norm_results <- do.call(rbind, lapply(names(all_normal_results), function(teams) {
  team_results <- do.call(rbind, all_normal_results[[teams]])
  team_results$teams <- as.numeric(teams)
  return(team_results)
}))
de_norm_results

```

## Shiny User Interface Tournament Simulation

```{r}
ui <- fluidPage(
  titlePanel("Double-Elimination Tournament Simulator"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 4, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      h4("Tournament Results"),
      tableOutput("ranked_results")
    )
  )
)

# Define the server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_elimination_tournament_structure(input$num_teams))
    }
  })

  tournament_results <- eventReactive(input$run_sim, {
    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs

    seed_structure <- custom_seeding()

    simulate_double_elimination_tournament(
      num_teams = input$num_teams,
      distribution = input$distribution,
      ties = input$ties,
      series = input$series,
      seeding_structure = seed_structure
    )
  })

  output$ranked_results <- renderTable({
    results <- tournament_results()
    results[order(results$true_rank), ]  # Order by Strength descending
  })
}

shinyApp(ui = ui, server = server)
```

## Shiny Number of Wins

```{r}
ui <- fluidPage(
  titlePanel("Tournament Simulation with Total Wins Visualization"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      h4("Total Wins Summary"),
      plotOutput("wins_plot")
    )
  )
)

# Shiny server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_elimination_tournament_structure(input$num_teams))
    }
  })

  simulation_results <- eventReactive(input$run_sim, {
    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
    seed_structure <- custom_seeding()
    # Run simulations
    replicate_results <- replicate(
      input$num_simulations,
      simulate_double_elimination_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        seeding_structure = seed_structure
      ),
      simplify = FALSE
    )

    # Combine all results and summarize total wins
    all_results <- do.call(rbind, replicate_results)
    total_wins_summary <- all_results %>%
      group_by(true_rank) %>%
      summarise(
        total_wins = sum(rank_hat==1, na.rm = TRUE),
        .groups = 'drop'
      ) %>%
      arrange(desc(total_wins))
    return(total_wins_summary)
  })

  output$wins_plot <- renderPlot({
    req(simulation_results())
    results <- simulation_results()

    ggplot(results, aes(x = reorder(true_rank, -total_wins), y = total_wins, fill = total_wins)) +
      geom_text(aes(label = total_wins), vjust = -0.5, size = 4, color = "black") +
      geom_bar(stat = "identity") +
      labs(
        title = "Total Wins Across Simulations",
        x = "True Rank",
        y = "Total Wins"
      ) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_fill_gradient(low = "skyblue", high = "blue")
  })
}

# Run the Shiny app
shinyApp(ui = ui, server = server)
```

# Multiple Simulations for 1 Seeding Structure

```{r}
library(shiny)

ui <- fluidPage(
  titlePanel("Tournament Simulation"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", 8, min = 4, step = 1),
      selectInput("distribution", "Distribution Type:", choices = c("Normal", "Uniform")),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      numericInput("replicates", "Number of Replicates:", 100, min = 1, max = 1000, step = 1),
      numericInput("series", "Number of Games in a Series (odd number):", 1, min = 1, max = 11, step = 2),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      actionButton("run_sim", "Run Tournament Simulation")
    ),

    mainPanel(
      tableOutput("summary_results"),
      verbatimTextOutput("kendall_correlation"),
      verbatimTextOutput("spearman_correlation")
    )
  )
)

server <- function(input, output, session) {
 custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_elimination_tournament_structure(input$num_teams))
    }
  })

  # Function to run tournament simulations
  observeEvent(input$run_sim, {
    req(input$num_teams, input$distribution, input$replicates, input$series)

    # Initialize lists to store correlation results
    kendall_correlations <- numeric(input$replicates)
    spearman_correlations <- numeric(input$replicates)

    seed_structure <- custom_seeding()

    # Run tournament simulations for the specified number of replicates
    simulation_results <- list()
    for (replicate in 1:input$replicates) {
      simulation_results[[replicate]] <- simulate_double_elimination_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        seeding_structure = seed_structure
      )

      # Combine all results into one data frame
      all_results <- do.call(rbind, simulation_results)

      # Compute average ranks across all replicates
      average_ranks <- all_results %>%
        group_by(true_rank) %>%
        summarise(avg_rank = mean(rank_hat, na.rm = TRUE), .groups = 'drop') %>%
        arrange(true_rank)

      # Calculate Kendall Tau correlation for the current replicate
      kendall_correlations[replicate] <- cor(all_results$rank_hat, all_results$true_rank, method = "kendall", use = "complete.obs")

      # Calculate Spearman correlation for the current replicate
      spearman_correlations[replicate] <- cor(all_results$rank_hat, all_results$true_rank, method = "spearman", use = "complete.obs")
    }

    # Compute average Kendall Tau and Spearman correlations across all replicates
    avg_kendall_cor <- mean(kendall_correlations, na.rm = TRUE)
    avg_spearman_cor <- mean(spearman_correlations, na.rm = TRUE)

    # Show summary results in table
    output$summary_results <- renderTable({
      average_ranks
    })

    # Display average Kendall Tau correlation
    output$kendall_correlation <- renderPrint({
      cat("Average Kendall Tau Correlation:", avg_kendall_cor, "\n")
    })

    # Display average Spearman correlation
    output$spearman_correlation <- renderPrint({
      cat("Average Spearman Correlation:", avg_spearman_cor, "\n")
    })
  })
}

shinyApp(ui, server)
```

## Shiny Probability of Top 3 Correct

```{r}
ui <- fluidPage(
  titlePanel("Probability of Top 3 Teams in the Correct Order"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      h4("Top 3 Correct"),
      verbatimTextOutput("top3_output")
    )
  )
)

# Shiny server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_elimination_tournament_structure(input$num_teams))
    }
  })

  simulation_results <- eventReactive(input$run_sim, {
    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
    seed_structure <- custom_seeding()

    # Run simulations
    replicate_results <- replicate(
      input$num_simulations,
      simulate_double_elimination_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        seeding_structure = seed_structure
      ),
      simplify = FALSE
    )

    # Combine all results into one dataframe and add a Simulation column
    all_results <- do.call(rbind, replicate_results)
    all_results$simulation <- rep(1:input$num_simulations, each = nrow(all_results) / input$num_simulations)

    return(all_results)
  })

  # Identify top-3 ranks for each simulation and calculate the probability
  top_three_probability <- reactive({
    all_results <- simulation_results()

    # Group by simulation and filter for top 3 true_rank and rank_hat
    top_three_results <- all_results %>%
      group_by(simulation) %>%
      filter(true_rank %in% 1:3 & rank_hat %in% 1:3) %>%
      arrange(simulation, rank_hat) %>%
      summarize(top_3_true_rank = list(true_rank), top_3_rank = list(rank_hat), .groups = 'drop') %>%
      mutate(match = mapply(function(true, rank) setequal(true, rank), top_3_true_rank, top_3_rank))

    # Calculate probability of matching top 3 true_rank with rank_hat
    match_count <- sum(top_three_results$match)
    total_simulations <- n_distinct(all_results$simulation)

    probability <- match_count / total_simulations
    return(probability)
  })

  # Output the probability to the UI
  output$top3_output <- renderText({
    probability <- top_three_probability()

    paste("Probability of Top 3 Seeds with Ranks 1-3: ", round(probability,4))
  })
}

shinyApp(ui=ui,server=server)
```

## Shiny Conditional Probability Curve

```{r}
ui <- fluidPage(
  titlePanel("Conditional Probability of Team Ranks in Tournament"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", 
                placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      plotOutput("probability_plot")
    )
  )
)

# Server
server <- function(input, output) {

  # Helper Function: Parse Custom Seeding Structure
  parse_seeding <- function(seeding_input, num_teams) {
    if (seeding_input != "") {
      seeding_vector <- unlist(strsplit(seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)

      num_teams_next_power_of_two <- 2^ceiling(log2(num_teams))
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(
          paste("Seeding structure length must match", num_teams_next_power_of_two), 
          type = "error"
        )
        return(NULL)
      }
    } else {
      return(NULL)
    }
  }

  # Curve Function to Calculate Conditional Probabilities
  curve <- function(sim, num_teams) {
    probability <- numeric(num_teams)
    for (i in 1:num_teams) {
      if (i == 1) {
        probability[i] <- nrow(sim[sim$rank_hat == 1 & sim$true_rank == 1, ]) / length(unique(sim$simulation))
      } else {
        top_ranks <- sim[sim$rank_hat <= (i - 1) & sim$true_rank <= (i - 1), ]
        better_teams <- 1:(i - 1)
        numerator <- sum(sapply(unique(top_ranks$simulation), function(sim_id) {
          sim_subset <- sim[sim$simulation == sim_id, ]
          if (all(better_teams %in% sim_subset$true_rank[sim_subset$rank_hat <= (i - 1)])) {
            return(sum(sim_subset$rank_hat == i & sim_subset$true_rank == i))
          } else {
            return(0)
          }
        }))
        denominator <- sum(sapply(unique(top_ranks$simulation), function(sim_id) {
          sim_subset <- top_ranks[top_ranks$simulation == sim_id, ]
          all(better_teams %in% sim_subset$true_rank)
        }))
        probability[i] <- numerator / denominator
      }
    }
    return(probability)
  }

  # Reactive Event: Run Simulations
  simulation_data <- eventReactive(input$run_sim, {
    num_teams <- input$num_teams
    distribution <- input$distribution
    series <- input$series
    num_simulations <- input$num_simulations

    # Parse custom seeding structure
    seeding_structure <- parse_seeding(input$seeding_input, num_teams)

    # Run simulations
    replicate_results <- replicate(
      num_simulations,
      simulate_double_elimination_tournament(
        num_teams = num_teams,
        distribution = distribution,
        series = series,
        ties = FALSE,
        seeding_structure = seeding_structure
      ),
      simplify = FALSE
    )

    # Combine all results into one dataframe
    #all_results <- do.call(rbind, replicate_results)
    #all_results$simulation <- rep(1:num_simulations, each = nrow(all_results) / num_simulations)
    all_results <- list_rbind(x=replicate_results, names_to = "simulation")

    # Calculate conditional probabilities
    sim_probs <- curve(all_results, num_teams)
    probabilities <- data.frame(x = 1:num_teams, y = sim_probs)
    probabilities$line <- "Simulated"

    # Baseline probabilities
    baseline_prob <- 1 / (num_teams:1)
    baseline <- data.frame(x = 1:num_teams, y = baseline_prob)
    baseline$line <- "Baseline"

    # Combine for plotting
    rbind(baseline, probabilities)
  })

  # Plot Output
  output$probability_plot <- renderPlot({
    req(simulation_data())
    ggplot(data = simulation_data(), aes(x = x, y = y, color = line, group = line)) +
      geom_point() +
      geom_line() +
      ylim(0, 1) +
      labs(
        x = "Rank",
        y = "Conditional Probability",
        title = "Simulated vs Baseline Conditional Probabilities"
      ) +
      theme_minimal()
  })
}

# Run the App
shinyApp(ui = ui, server = server)
```