### 2025-05-01 group_stage.qmd

---
title: "Group Stage Play"
author: "Zach Culp"
format: html
editor: visual
---

## Group Stage

### Simulate Game

```{r}
simulate_match <- function(team1, team2, strength1, strength2) {
  if(is.na(team1) && is.na(team2)){
    return(NA)
  }
  if (is.na(team1)) {
    return(team2)
  } else if (is.na(team2)) {
    return(team1)
  } else {
    p <- exp(strength1 - strength2) / (1 + exp(strength1 - strength2))
    return(ifelse(runif(1) < p, team1, team2))  # Randomize outcome
  }
}
```

### Simulate Each Group

```{r}
simulate_group_play <- function(strengths, rounds=1){
  teams <- strengths$true_rank
  n <- length(teams)
  win_counts <- numeric(n)
  loss_counts <- numeric(n)
  for (i in 1:rounds){
    for (i in seq_along(teams)){
      for (j in seq_along(teams)){
        if (i != j){
          team1 <- teams[i]
          team2 <- teams[j]
          strength1 <- strengths$true_strength[strengths$true_rank == team1]
          strength2 <- strengths$true_strength[strengths$true_rank == team2]
          match_winner <- simulate_match(teams[i],teams[j],strength1, strength2)
          if (match_winner == teams[i]) {
            win_counts[i] <- win_counts[i] + 1
            loss_counts[j] <- loss_counts[j] + 1
          } else {
            win_counts[j] <- win_counts[j] + 1
            loss_counts[i] <- loss_counts[i] + 1
          }
        }
      }
    }
  }
  results <- data.frame(
    true_rank = teams,
    game_wins = win_counts,
    game_loses = loss_counts,
    true_strength = strengths$true_strength,
    groups = strengths$groups
  )

  # Ranks teams based on number of wins (includes ties)
  results$rank_hat <- rank(-results$game_wins, ties.method = "random")
  results <- results[order(results$rank_hat),]

  return(results)
}

simulate_groups <- function(group_length, num_teams, distribution, num_advance){
  teams <- paste(num_teams:1)
  if (distribution == "Normal"){
    strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA,length(teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  }
  else if (distribution == "Same"){
    strengths <- sapply(num_teams, function(n) {rep(0,n)})
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA,length(teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  }
  else if (distribution == "Uniform"){
    strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1) , 0 , sqrt(12)) })
    unif_strength <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams)),
      rank_hat = rep(NA,length(teams))
    )
    df <- arrange(unif_strength, true_rank)
  }

  num_groups <- num_teams / group_length
  # Create tiers based on sorted true_strength
  df <- df %>% arrange(desc(true_strength))
  tiers <- split(df, ceiling(seq_along(df$true_rank) / num_groups))

  # Initialize group vector
  df$groups <- NA

  # Assign one team from each tier to each group randomly
  for (tier in tiers) {
    group_ids <- sample(1:num_groups)  # random group order for this tier
    df$groups[match(tier$true_rank, df$true_rank)] <- group_ids
  }

  all_results <- list()
  for (g in unique(df$groups)){
    group_df <- df[df$groups == g, ]
    group_results <- simulate_group_play(strengths = group_df)
    group_results$groups <- g
    all_results[[g]] <- group_results
  }
  print(all_results)
  final_df <- do.call(rbind, all_results)
  teams_in_knockout <- final_df %>%
    arrange(rank_hat, game_wins) %>%
    head(num_advance * num_groups)
  return(teams_in_knockout)
}
create_knockout_matches <- function(df) {
  group_list <- sort(unique(df$groups))
  matches <- list()

  for (i in seq(1, length(group_list), by = 2)) {
    g1 <- group_list[i]
    g2 <- group_list[i + 1]

    # Get the winner from group g1 and runner-up from group g2
    team1 <- df[df$groups == g1 & df$rank_hat == 1, ]
    team2 <- df[df$groups == g2 & df$rank_hat == 2, ]

    # Get the winner from group g2 and runner-up from group g1
    team3 <- df[df$groups == g2 & df$rank_hat == 1, ]
    team4 <- df[df$groups == g1 & df$rank_hat == 2, ]

    matches[[length(matches) + 1]] <- data.frame(
      Match = paste("Match", length(matches) + 1),
      TeamA = team1$true_rank,
      TeamB = team2$true_rank
    )

    matches[[length(matches) + 1]] <- data.frame(
      Match = paste("Match", length(matches) + 1),
      TeamA = team3$true_rank,
      TeamB = team4$true_rank
    )
  }

  do.call(rbind, matches)
}
simulate_knockout <- function(df, series=1, third_place = T){

  seeding_structure <- create_knockout_matches(df)

  for (i in nrow(df)){
    for (g in unique(df$groups)){

    }
  }

  results <- list()  # Initialize the results list

  for (i in 1:nrow(seeding_structure)) {
    # Create a new data frame based on the current permutation in 'test'
    permutation <- seeding_structure[i, ]
    permuted_df <- df
    permuted_df$true_rank <- df$true_rank[permutation]  # Update team order 
    teams <- permuted_df$true_rank
    permuted_df$true_strength <- df$true_strength[permutation]
    permuted_df$rank_hat <- NA  # Initialize ranks column
    permuted_df$game_wins <- 0  # Initialize wins column
    permuted_df$game_losses <- 0

    strengths <- permuted_df$true_strength
    round_number <- ceiling(log2(length(teams)))

    # Track the losers of the semifinals for the third-place match
    losers_semis <- c()

    while (length(teams) > 1) {
      #cat("\n--- New Round ---\n")
      next_round <- c() # Makes sure only winners are included in next round

      for (j in seq(1, length(teams), by = 2)) { # Gets 2 teams for each match
        team1 <- teams[j]
        team2 <- teams[j+1]

        strength1 <- strengths[which(permuted_df$true_rank == team1)]
        strength2 <- strengths[which(permuted_df$true_rank == team2)]

        wins_team1 <- 0
        wins_team2 <- 0
        losses_team1 <- 0
        losses_team2 <- 0
        for (game in 1:series){
          match_winner <- simulate_match(team1, team2, strength1, strength2)
          #cat(sprintf("Matchup:  %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", team1, team2, strength1, strength2, match_winner))
          if(!is.na(team1)) {
          if (match_winner == team1) {
            wins_team1 <- wins_team1 + 1
            losses_team2 <- losses_team2 + 1
          } else {
            wins_team2 <- wins_team2 + 1
            losses_team1 <- losses_team1 + 1
          }
          }
          else {wins_team2 <- series / 2 + 1}

          # Determine the overall winner if a majority is reached
          if (wins_team1 > series / 2 || wins_team2 > series / 2) {
            break
          }
        }

        # adds wins and losses of the series to the teams unless they have a bye
        if (!any(is.na(c(team1, team2)))) {
          permuted_df <- permuted_df %>%
            mutate(game_wins = ifelse(true_rank == team1, game_wins + wins_team1, game_wins)) %>%
            mutate(game_wins = ifelse(true_rank == team2, game_wins + wins_team2, game_wins)) %>%
            mutate(game_losses = ifelse(true_rank == team1, game_losses + losses_team1, game_losses)) %>%
            mutate(game_losses = ifelse(true_rank == team2, game_losses + losses_team2, game_losses)) 
           }

        # Determine the match winner based on series results
        series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
        #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, series_winner))
        next_round <- c(next_round, series_winner)

        if (round_number > 2){
          # Assign ranks based on the current round
          n <- 2^(round_number)
          #print(series_winner)
            if (!is.na(team1) & series_winner == team1) {
                permuted_df$rank_hat[permuted_df$true_rank == team2] <- n
            } else {
                permuted_df$rank_hat[permuted_df$true_rank == team1] <- n
            }
        } else if (round_number == 2) {  # Semifinals
          n <- 2^(round_number)
          if (match_winner == team1) {
              permuted_df$rank_hat[permuted_df$true_rank == team2] <- n
            losers_semis <- c(losers_semis, team2)
          } else {
              permuted_df$rank_hat[permuted_df$true_rank == team1] <- n
            losers_semis <- c(losers_semis, team1)
          }
        }else if (round_number == 1) { # Finals
          permuted_df <- permuted_df %>%
            filter(!is.na(true_rank))
          n <- 2^(round_number)
          if (series_winner == team1) {
            permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
            permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
          } else {
            permuted_df$rank_hat[permuted_df$true_rank == team1] <- 2
            permuted_df$rank_hat[permuted_df$true_rank == team2] <- 1
          }
        }
      }

      teams <- next_round  # Winners move to the next round
      round_number <- round_number - 1
    }

    # Third place match between the losers of the semifinals
    if (third_place == TRUE){
      team1 <- losers_semis[1]
      team2 <- losers_semis[2]
      strength1 <- permuted_df$true_strength[permuted_df$true_rank == team1]
      strength2 <- permuted_df$true_strength[permuted_df$true_rank == team2]
      wins_team1 <- 0
      wins_team2 <- 0
      for (game in 1:series) {
        match_winner <- simulate_match(team1, team2, strength1, strength2)
        if (match_winner == team1) {
          wins_team1 <- wins_team1 + 1
        } else {
          wins_team2 <- wins_team2 + 1
        }
        # Determine the overall winner if a majority is reached
        if (wins_team1 > series / 2 || wins_team2 > series / 2) {
          break
        }
      }
      # Assign third place and fourth place ranks
      if (wins_team1 > wins_team2) {
        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team1))
        permuted_df$rank_hat[permuted_df$true_rank == team1] <- 3
        permuted_df$rank_hat[permuted_df$true_rank == team2] <- 4
      } else {
        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team2))
        permuted_df$rank_hat[permuted_df$true_rank == team2] <- 3
        permuted_df$rank_hat[permuted_df$true_rank == team1] <- 4
      }
    }

    permuted_df$seed <- rep(paste(seeding_structure[i, ], collapse = ", "), nrow(permuted_df))
    results[[i]] <- permuted_df # Store the results of this permutation
  }

  final_results <- do.call(rbind, results)

  # Remove ties if ties = FALSE
  if (!ties) {
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(true_rank) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "random"))
  } else{
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(rank_hat) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "average"))
  }

  final_results <- final_results %>%
    mutate(distribution = distribution) %>%
    mutate(true_rank = as.numeric(true_rank)) %>%
    arrange(true_rank)
  return(final_results)
}

knockout_teams <- simulate_groups(4,8,"Normal",2)
create_knockout_matches(knockout_teams)
```