### 2025-05-01 mutual_information_curves.qmd

---
title: "Mutual Information Plots"
format: html
editor: visual
---

## Mutual Information Plots

From Wikipedia, weighted mutual information formula:

$I(X;Y)=\sum _{y\in Y}\sum _{x\in X}w(x,y)p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}$

Rajski's distance: $1 - \frac{I(X,Y)}{H(X,Y)}$

Interpretation: the fraction of uncertainty not explained by mutual information

<https://en.wikipedia.org/wiki/Inverse_distance_weighting>

Every permutation is not possible to elimination tournaments, so should we divide by every possible permutation for 8 teams, or just every possible permutation in the tournament structure?

```{r}
library(combinat)
library(tidyverse)

mut_info <- function(sims, num_teams, rank_weights = NULL){
  # Handle missing weights (default to equal weights if not provided)
  if (is.null(rank_weights)){
    rank_weights <- rep(1,num_teams)
  }
  # Extract y values from sims, assuming `rank_hat` contains a list of rankings per simulation
  y_df <- sims %>%
    group_by(simulation) %>%
    summarise(rank_hat = list(rank_hat)) %>%  # Ensure rank_hat is a list column
    mutate(rank_hat = map(rank_hat, unlist)) %>%  # Unlist rank_hat properly
    unnest_wider(rank_hat, names_sep = "_")  # Expand into separate columns

  # Rename y columns
  colnames(y_df)[-1] <- paste0("y", 1:num_teams)

  # Create x matrix, representing positions 1:num_teams
  x_matrix <- matrix(rep(1:num_teams, times = nrow(y_df)), ncol = num_teams, byrow = TRUE)
  x_df <- as.data.frame(x_matrix)
  colnames(x_df) <- paste0("x", 1:num_teams)

  # Combine x and y data frames
  perms_df <- cbind(x_df, y_df %>% select(-simulation))
  perms_df <- perms_df %>%
    mutate(across(starts_with("x"), as.numeric)) %>%
    mutate(across(starts_with("y"), as.numeric))

  # Generate all possible permutations of rankings
  all_perms <- as.data.frame(do.call(rbind, permn(1:num_teams)))
  colnames(all_perms) <- paste0("y", 1:num_teams)

  # Count occurrences of each observed ranking
  prob_df <- perms_df %>%
    group_by(across(starts_with("x")),across(starts_with("y")))  %>%
    summarise(count = n(), .groups = "drop")

  # Merge with full permutation set to ensure all permutations exist
  prob_df <- right_join(all_perms, prob_df, by = colnames(all_perms)) %>%
    mutate(count = replace_na(count, 0)) %>%
    mutate(probability = count / sum(count))  # Normalize probabilities

  w <- numeric(nrow(prob_df))  # Initialize weight vector for the number of rows in prob_df

  for (i in 1:num_teams) {
    x_i <- paste0("x", i)
    y_i <- paste0("y", i)

    # Loop through each row of prob_df
    for (j in 1:nrow(prob_df)) {
      #x_vals <- as.numeric(prob_df[j, paste0("x", 1:num_teams)])
      #y_vals <- as.numeric(prob_df[j, paste0("y", 1:num_teams)])
      # Calculate the squared difference between x_i and y_i for each row j
      diff <- abs(prob_df[[x_i]][j] - prob_df[[y_i]][j])
      #rank_diff <- abs(rank(x_vals) - rank(y_vals))
      w[j] <- w[j] + rank_weights[i] * diff^2  # Sum squared differences for each row
      # Count inversions (pairs where order is flipped)
      #inversions <- sum(outer(x_vals, x_vals, ">") & outer(y_vals, y_vals, "<"))
      #w[j] <- sum(exp(rank_diff) - 1)
      #w[j] <- w[j] + rank_weights[i] * (exp(inversions) - 1)
      #w[j] <- cor(x_vals, y_vals, method = "spearman", use = "complete.obs")
      # Kendal's Tau as weights?
    }
  }
  #w <- (w + 1) / 2
  #w[w == Inf] <- 0
  w <- 1 / (w+1)  # Inverse of squared differences and manually change 0 probabilities to 1 on the inverse
  #w <- w/sum(w)
  # Compute weights dynamically for each row
  weighted_df <- prob_df %>%
    mutate(w = w)
  #print(weighted_df)

  numerator <- sum(weighted_df$w * weighted_df$probability * log2(weighted_df$probability / (1/factorial(num_teams) * (1/factorial(num_teams)))))
  denominator <- log2(factorial(num_teams))

  Raj_dist <- 1 - numerator/denominator

  # Return the weighted data frame with mutual information weights
  return(Raj_dist)
}

de_norm_results_eight <- de_norm_results %>% filter(teams == 8)
de_norm_results_eight$simulation <- rep(1:10000, each = nrow(de_norm_results_eight) / 10000)

all_results_norm$simulation <- rep(1:10000, each = nrow(all_results_norm) / 10000)
seven_games_se_results$simulation <- rep(1:10000, each = nrow(seven_games_se_results) / 10000)
# with randomly assigning ties
test_rr_noTies <- test_rr %>%
  group_by(simulation) %>%
  mutate(rank_hat = rank(rank_hat,ties.method="random"))
perms_list <- permn(8)
all_data <- list()
# Loop through each permutation and create the data frame
for (i in 1:length(perms_list)) {
# Create a temporary data frame for the current permutation
temp_df <- data.frame(
true_rank = rep(1:8, each = 1),  # Each row gets a true_rank from 1 to 8
rank_hat = perms_list[[i]],      # The current permutation's ranks
simulation = rep(i, 8)           # Assign the simulation number to each row
)
# Add the temporary data frame to the list
all_data[[i]] <- temp_df
}
# Combine all the data frames into one
perms_df <- do.call(rbind, all_data)

best <- data.frame(
  true_rank = rep(1:8, times = 10000),  # Repeat 1-8 for 10,000 simulations
  rank_hat = rep(1:8, times = 10000),   # Ensure rank_hat matches true_rank
  simulation = rep(1:10000, each = 8)   # Assign each set to a simulation
)
worst <- data.frame(
  true_rank = rep(1:8, times = 10000),  # Repeat 1-8 for 10,000 simulations
  rank_hat = rep(8:1, times = 10000),   # Ensure rank_hat matches true_rank
  simulation = rep(1:10000, each = 8)   # Assign each set to a simulation
)

# Function to compute mutual information for a given dataset
compute_mut_plots <- function(dataset) {
  num_teams <- length(unique(dataset$true_rank))
  results <- numeric(num_teams)
  for (i in 1:num_teams) {
    binary_vector <- c(rep(1, i), rep(0, num_teams - i))  # Construct vector
    results[i] <- mut_info(dataset, num_teams, binary_vector)  # Compute MI
  }
  return(results)
}

# Compute mutual information for each dataset
mi_de_norm <- compute_mut_plots(de_norm_results_eight)
mi_se_norm <- compute_mut_plots(all_results_norm)
mi_test_rr <- compute_mut_plots(test_rr_noTies)
mi_seven_se <- compute_mut_plots(seven_games_se_results)
mi_random <- compute_mut_plots(equal_strengths)
mi_se_unif <- compute_mut_plots(test_unif)
mi_bad <- compute_mut_plots(bad_seeding_structure)
mi_perms <- compute_mut_plots(perms_df)
mi_best <- compute_mut_plots(best)
mi_worst <- compute_mut_plots(worst)

mi_plots_df <- data.frame(
indices = 1:8,
"Double Elimination" = (1-mi_de_norm),
"Single Elimination" = (1-mi_se_norm),
"Equal Strengths Single Elimination" = (1-mi_random),
#"Unif_SE" = (1-mi_se_unif),
"Bad Seeding Structure Single Elimination" = (1-mi_bad),
"Every Possible Permutation" = (1-mi_perms),
"Round Robin" = (1-mi_test_rr),
"Worst Possible" = (1-mi_worst),
"Best Possible" = (1-mi_best)
)

# Reshape data into long format for ggplot
df_long <- pivot_longer(mi_plots_df, cols = -indices, names_to = "Method", values_to = "MI")
ggplot(df_long, aes(x = indices, y = MI, color = Method, group = Method)) +
geom_line() +           # Add lines
geom_point() +          # Add points
labs(
title = "Comparison of Tournament Structures with Normal Strengths",
x = "Top i Number of Teams",
y = "Mutual Information \\ Entropy (log10 scale)"
) +
theme_minimal() +       # Use a clean theme
theme(legend.position = "right") +
  scale_y_log10(
    labels = scales::label_number() 
  )

normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}

# Apply normalization to all columns except "indices"
mi_norm_plot <- mi_plots_df
mi_norm_plot[ , -1] <- t(apply(mi_plots_df[ , -1], 1, normalize))
# Reshape data into long format for ggplot
df_long_norm <- pivot_longer(mi_norm_plot, cols = -indices, names_to = "Method", values_to = "MI")
windowsFonts(Arial = windowsFont("Arial"))
ggplot(df_long_norm, aes(x = indices, y = MI, color = Method, group = Method)) +
geom_line(linewidth = 2.5) +           # Add lines
geom_point(size=3) +          # Add points
labs(
title = "Comparison of Tournament Structures with Normal Strengths",
x = "Top i Number of Teams",
y = "Normalized Mutual Information \\ Entropy (log10 scale)"
) +
theme_minimal() +       # Use a clean theme
theme(text = element_text(family = "Arial", size = 32),
  legend.position = "bottom",
  legend.box = "vertical",
  plot.title = element_text(hjust = 0.5, size = 36),
  axis.title = element_text(size = 30),
  axis.text = element_text(size=30)) +
  guides(color = guide_legend(nrow = 4)) +
  scale_y_log10(
    labels = scales::label_number() 
  ) +
  scale_color_viridis_d(option = "H")
ggsave(
  filename = "tournament_comparison_plot.png",  # Change file type if needed
  plot = last_plot(),  # Saves the last ggplot created
  dpi = 600,        
  bg = "white",
  units = "in",
  width = 15.28,
  height = 13.43
)
```

include number of games (might be a range)

Single Elimination - third place and without third place; series = 3,5,7

Double Elimination

-   Repechage

-   True Double Elimination

-   Consolation Bracket (loser's bracket best place is third)

Triple Elimination

Deterministic Outcomes

Round Robin - with k rounds (1,2,3,5,7)

Stepladder - 8 vs 7, winner plays 6, etc.

Group Stage - winners of group advance, top 2 advance

Staged Round Robin/Jacobian Ladder - top 2 in each group form new group, bottom 2 new group (each new group is a new round robin)

High School Tennis Ladder - 1 vs 2, 3 vs 4, etc., Re-rank 2 vs 3, 4 vs 5, 6 vs 7, number of teams - 1 iterations