### 2025-05-01 manuscript/paper.Rmd
paper update with Results section

---
title: |
  |  \large Tournaments 
author: | 
  | \large Zach Culp \vspace{-1.1mm}
  | \normalsize  \vspace{-1mm}
  | \normalsize Loyola Chicago \vspace{-1mm}
  | \normalsize Chicago \vspace{-1mm}
  |
  | \large Josie Peterburgs \vspace{-1.1mm}
  | \normalsize College of Education \vspace{-1mm}
  | \normalsize University of Texas at Austin \vspace{-1mm}
  | \normalsize Austin, TX 78712 \vspace{-1mm}
  |
  | \large Ryan McShane \vspace{-1.1mm}
  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
  | \normalsize Loyola University Chicago \vspace{-1mm}
  | \normalsize Chicago, IL 60660 \vspace{-1mm}
  |
  | \large Gregory J. Matthews \vspace{-1.1mm}
  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
  | \normalsize Center for Data Science and Consulting \vspace{-1mm}
  | \normalsize Loyola University Chicago \vspace{-1mm}
  | \normalsize Chicago, IL 60660 \vspace{-1mm}
  | \normalsize [`email`](mailto:ypu@something.edu) \vspace{-1mm}
  | \normalsize [`brian.mills@austin.utexas.edu`](mailto:brian.mills@austin.utexas.edu) \vspace{-1mm}
  | \normalsize [`lderango@luc.edu`](mailto:lderango@luc.edu) \vspace{-1mm}
  | \normalsize [`gmatthews1@luc.edu`](mailto:gmatthews1@luc.edu) \vspace{-1mm}
abstract: | 
 | Here is the abstract.
 | 
  \vspace{2mm}
  | Keywords: survival analysis, hockey
bibliography: references.bib
fontsize: 12pt
link-citations: true
linkcolor: cyan
urlcolor: cyan
output:
  pdf_document:
    df_print: kable
    number_sections: true
    keep_tex: true
header-includes:
 \usepackage{setspace}
 \setstretch{1.15}
 \usepackage{float}
 \floatplacement{figure}{t}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)

gjm <- function(x, color = "red") {
  sprintf("\\textcolor{%s}{%s}", color, x)}
```

```{r pkgs}
library(tidyverse)
library(dplyr)
library(stringr)
theme_set(theme_minimal())
```

# Introduction

There are hundreds of different tournament structures that dictate how a
set of teams compete against each other to determine an overall ranking.
Different tournament structures have unique strengths and weaknesses,
affecting how well they reflect the true rankings of the teams. The goal
of a tournament is to find the best teams, but how can we quantify how
well a tournament performs? In some instances, the tournament organizers
only award the overall winner, but other times, the top three or ten
teams are awarded. Because of this, organizers may prefer one structure
over another based on their needs. Tournament organizers must also take
into account factors like cost, timeliness, entertainment value, and
fairness when choosing a tournament structure.

To evaluate a tournament’s effectiveness, we propose a numerical metric
that quantifies how accurately it orders teams based on their true
rankings. Tournament results can be viewed as a "message" attempting to
convey the true team rankings. However, various factors introduce noise,
leading to information loss. We use principles from information theory
to measure this information loss across multiple tournament simulations
to assess the reliability of different formats.

# Methods

Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with
an associated true strength parameter
$\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that
$(\theta_1 > \theta_2 > \ldots > \theta_n)$. Next, let
$r(\boldsymbol{\theta})$ be the indexes of $\boldsymbol{\theta}$ when
theta is sorted from largest to smallest so that
$r(\boldsymbol{\theta}) = \{1, 2, \ldots, n-1, n\}$. Next define
$T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure
$s$. $\phi$ contains all the information about the scheduling of teams,
which could be fully known prior to the tournament (i.e. round robin) or
determined as the tournament progresses (i.e. single elimination
tournament). We then let $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be an
$n$- dimensional vector valued random variable that gives the results of
a tournament as a vector of the indexes of the vector
$\boldsymbol{\theta}$. The index in the first position of the vector
$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the
team that finished first, the index in the second position of the vector
$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the
team that finished second, and so on.

For example, in a 4 team tournament,
$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) indicates that
the team with index 3 (i.e. the true third best team) won the
tournament, the true second best team finished second, the true best
team finished third and the true 4th team finished 4th in that
particular tournament. If $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ =
(1, 2, 3, 4), this means that the random outcome of the tournament was
the same as the true ordering of the teams.

If one views the true ranking of the teams in a tournament as a message
to be sent to a receiver and the outcome of the tournament as a message
that is received, we can measure the "goodness" of a tournament in terms
of its ability to accurately transmit the true ranking. We can then use
concepts from information theory to assess the ability of a tournament
to correctly rank teams. Specifically, we start with the concept of
mutual information (CITE). For two random variables $X$ and $Y$ mutual
information is defined as:

$$
I(X;Y)=\sum_{y\in Y}\sum_{x\in X}p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}
$$.

In our setting, we replace $X$ and $Y$ with $r(\boldsymbol{\theta})$ and
$\hat{r}(T(\phi, s),\boldsymbol{\theta})$, and we could seek to estimate
$I(r(\boldsymbol{\theta}),\hat{r}(T(\phi, s),\boldsymbol{\theta}))$. For
simplicity, we drop the arguments from the functions $r$ and $\hat{r}$
for ease of exposition.

We note here that $r$ is not a random variable, however, since the
indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary
(we use 1, 2, 3, etc. for convenience, but any indexing is valid), we
can view this a random variable where all permutation so of the $n$
teams are equally likely. Therefore when calculating the mutual
information in this setting, we set $p(r) = \frac{1}{n!}$. By a similar
argument we set $p(\hat{r}) = \frac{1}{n!}$.

We define the mutual information of $r$ and $\hat{r}$ to be:

$$
I(r,\hat{r})=\sum_{r}\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})} = n!\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{(\frac{1}{n!})^2}
$$.

While there are $n!$ different permutations for the result of $r$, we
don't need to sum across these as any specific choice of $r$ is just an
arbitrary labeling of the true strength parameters vector
$\boldsymbol{\theta}$. So given $r$, the distribution of $\hat{r}$ is
the same up to the labeling. Therefore, we only need to consider a
single permutation of $r$, we compute the the quantity inside the
summation, and then multiple by $n!$ (effectively summing across $r$).

In order to compute this quantity, we need to compute $p(r,\hat{r})$.
This probability is found by calculating the probability of a given
permutation of $\hat{r}$ and $r$ is assumed to be the permutation from
$\{1, 2, \ldots, n\}$. We estimate $p(r,\hat{r})$ empirically through
simulation.

However, using mutual information in this form does not suit our needs
in this setting. The problem is that mutual information in this form
will yield high values of mutual information when the output from the
tournament $\hat{r}$ is highly consistent *even if the ranking from the
tournament is incorrect*. As an example, if $n=4$ and the true order of
\boldsymbol{\theta} is $r = \{1, 2, 3, 4\}$ and
$p(\hat{r} = \{4, 3, 2, 1\}) = 1$ will be the same mutual information as
when $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{1, 2, 3, 4\}) = 1$ and in
both of these cases the mutual information will be maximized at:

$$
I(r,\hat{r})= n!\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{\left(\frac{1}{n!}\right)^2} = n! log((n!)^2)
$$ and for the specific case when $n=4$ would be: $$
4! log_2(4!^2) = 24*log_2(24^2) = 220.0782
$$.

In order to alleviate this problem, we instead consider *weighted*
mutual information. We want to give more weight to permutations from
$\hat{r}$ that are "closer" to $r$. There are any number of reasonable
choices for this weighting function, and we choose to use squared error.
Specifically, \$\$ w(r, \hat{r}) =

\\left{\

\begin{array}{cc}
    \begin{array}{cc}
      \frac{1}{\hat{r}'r} & r \ne \hat{r} \\
      1 & r = \hat{r} \\
    \end{array}
\end{array}

\right. $$
In general, though any loss function $l(r, \hat{r})$ can be used:
$$ w(r, \hat{r}) =

\\left{\

\begin{array}{cc}
    \begin{array}{cc}
      \frac{1}{l(r, \hat{r})} & r \ne \hat{r} \\
      1 & r = \hat{r} \\
    \end{array}
\end{array}

\right. \$\$

Note that one is 1/N!

And this does work because consistently getting the order incorrect is
perfect information.

$$
I(X;Y)=\sum _{y\in Y}\sum _{x\in X}w(x,y)p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}
$$

Let $x$ be

x is the true ranking and y is the output

Rajski's distance: $$
1 - \frac{I(X,Y)}{H(X,Y)}
$$

<!-- # ryan way  -->

<!-- Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with an associated true strength parameter $\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that $(\theta_1 < \theta_2 < \ldots < \theta_n)$.  Next, let $r(\boldsymbol{\theta})$ be the rank of the vector $\boldsymbol{\theta}$ from largest to smallest so that $r(\boldsymbol{\theta}) = \{n, n-1, \ldots, 2, 1\}$.  Next define $T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure $s$.  $\phi$ contains all the information about the scheduling of teams, which could be fully known prior to the tournament (i.e. round robin) or determined as the tournament progresses (i.e. single elimination tournament).  We then let  $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be a vector valued random variable that gives the results of a tournament as a vector of ranks.   -->

<!-- Figure out what $\hat{r}$ should look like.   -->

$r = \{1,2,3,4\}$ which is the

Let $\boldsymbol{x} = (x_1, x_2, \ldots, x_n)$ be the permutation of
indexes of $\boldsymbol{\theta}$ such that
$x_1 = {i | \theta_i = max(\boldsymbol{\theta})}$

Let $x$ be

x is the true ranking and y is the output

$$
I(X;Y)=\sum _{y\in Y}\sum _{x\in X}w(x,y)p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}
$$

Rajski's distance: $$
1 - \frac{I(X,Y)}{H(X,Y)}
$$

# Results

In order to fairly compare tournament structures, we assume the true
strengths of the teams follow a normal distribution, where each team is
assigned an equally spaced quantile and each team's strength is the
z-score of its respective quantile. Using this assumption, 10,000
simulations, all with 8 total teams, were ran for a single game round
robin (each team plays each other once), a single elimination structure
with the usual seeding structure (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7), a
single elimination structure with a poor seeding structure (1 vs 2, 3 vs
4, 5 vs 6, 7 vs 8). For a better understanding of the results, the best
possible results (every structure has the best in rank 1, the 2nd best
in rank 2, etc), the worst possible results (every structure has the
worst in rank 1, the 2nd worst in rank 2, etc), every possible
permutation of the ranks once, and a single elimination tournament with
the usual seeding structure where every team is of equal strengths.
Because the metric found is unitless, we were able to normalize the
values using the best and worst results to make the range from 0 to 1,
with 0 as the worst and 1 as the best. After normalizing, we get the
following graph (note the y-axis is on a log10 scale):

![](images/Normal_Information_Curve_Graph.png){width="408"}

Based on the graph, the intuitive best tournament, round robin, does
show to be the best. Double elimination is slightly better than single
elimination, with the gap growing wider as the top number of teams
looked at increases. However, those are under the assumption that the
true strengths were ordered correctly in the seeding. If there is a poor
seeding structure, as seen in the graph, it can drastically impact the
effectiveness of the tournament. In the extreme case seen, the
tournament can actually become worse than deciding games by random
chance (like flipping a coin). Another important discovery is that the
single elimination structure with equal strengths is better than every
possible permutation. The line for every possible permutation would
likely be similar to a round robin structure with all equal strengths,
so perhaps the closer the teams are in true strength, the closer round
robin and single elimination become.

# Conclusion {#sec:conclusion}

# Acknowledgements {.unnumbered}

# Supplementary Material {.unnumbered}

# References