### 2025-09-06 manuscript/paper.Rmd
paper edits

---
title: |
  | \large Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information
author: | 
  | \large Zach Culp \vspace{-1.1mm}
  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
  | \normalsize Loyola University Chicago \vspace{-1mm}
  | \normalsize Chicago, IL 60660 \vspace{-1mm}
  |
  | \large Josie Peterburs \vspace{-1.1mm}
  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
  | \normalsize Loyola University Chicago \vspace{-1mm}
  | \normalsize Chicago, IL 60660 \vspace{-1mm}
  |
  | \large Ryan McShane \vspace{-1.1mm}
  | \normalsize University of Chicago \vspace{-1mm}
  | \normalsize Chicago, IL 60660 \vspace{-1mm}
  |
  | \large Gregory J. Matthews \vspace{-1.1mm}
  |	\normalsize  \vspace{-1mm}
  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
  | \normalsize Center for Data Science and Consulting \vspace{-1mm}
  | \normalsize Loyola University Chicago \vspace{-1mm}
  | \normalsize Chicago, IL 60660 \vspace{-1mm}
  | \normalsize [`email`](mailto:ypu@something.edu) \vspace{-1mm}
  | \normalsize [`zculp@luc.edu`](mailto:zculp@luc.edu) \vspace{-1mm}
  | \normalsize [`jpeterburs@luc.edu`](mailto:jpeterburs@luc.edu) \vspace{-1mm}
  | \normalsize [`rmcshane@uchicago.edu`](mailto:rmcshane@uchicago.edu) \vspace{-1mm}
  | \normalsize [`gmatthews1@luc.edu`](mailto:gmatthews1@luc.edu) \vspace{-1mm}
abstract: | 
 | We evaluate multiple competitions structures to assess their effectiveness in accurately reproducing the true underlying ranking of participating teams. Using repeated simulations for each structure, we compare teams’ true performance ranks to their simulated outcomes. To quantify the accuracy of each format, we calculate weighted mutual information between true and simulated rankings across a large number of replicates. "Salient" weights are applied to look at only the top i out of n total teams/competitors in a given competition structure. This information-theoretic approach allows us to objectively compare formats and identify which structures most effectively preserve rank order. Results are visualized through comparative plots, providing clear insights into the trade-offs and strengths of each competition design.
 | 
  \vspace{2mm}
  | Keywords: tournament structures, mutual information
bibliography: references.bib
fontsize: 12pt
link-citations: true
linkcolor: cyan
urlcolor: cyan
output:
  pdf_document:
    df_print: kable
    number_sections: true
    keep_tex: true
header-includes:
 \usepackage{setspace}
 \setstretch{1.15}
 \usepackage{float}
 \floatplacement{figure}{t}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)

gjm <- function(x, color = "red") {
  sprintf("\\textcolor{%s}{%s}", color, x)}
```

```{r pkgs}
library(tidyverse)
library(dplyr)
library(stringr)
theme_set(theme_minimal())
```

# Introduction

<!-- @Rajski1961 -->

<!-- @McShane2019 -->

<!-- Csato 2020: The incentive (in)compatibility of group-based qualification systems. World cup qualifying has problems where teams are better off exerting lower effort -->

<!-- @Sziklai2022 -->

There are hundreds of different competitions structures that dictate how
a set of teams compete against each other to determine an overall
ranking. Common structures include bracket tournaments where if a team
wins their competition they advance to the next round. Bracket
tournaments can be run as single elimination where competitors advance
or are eliminated based on a single match (e.g. NCAA College Basketball
Tournament, National Football League (NFL) playoffs, major tennis
tournaments) or where each round is contested as a best of $k$ series
(e.g. $k$ = 7: National Basketball Association (NBA) playoffs, National
Hockey League (NHL) playoffs, $k$ varies by round: Major League Baseball
(MLB), Women's National Basketball Association (WNBA)) and the series
winner advance while the series loser is eliminated. Additional styles
of the bracket tournaments include variations where competitors are not
eliminated after a single loss such as double elimination tournaments
where competitors continue until they have lost twice. Within double
elimination tournament, there are several variations. For instance, the
NCAA College Wrestling tournament is a double elimination tournament,
however, if a competitor loses a match they fall into the consolation
bracket and the best they can finish after a loss before the finals is
3rd place making this not a "true" double elimination format. Olympic
judo uses a variant of this system with single elimination until the
quarterfinals followed by a consolation bracket consisting of the four
quarterfinal losers who still have a chance for a bronze medal[^1].

[^1]: in Olympic judo they award two bronze medals to the two
    competitors who each win their side of the consolation bracket

"True" double elimination - a bracket tournament where a competitor can
still finish ranked first as long as a competitor has not lost two
games - is used in NCAA baseball and softball for certain rounds in
their postseason. Additional variants of double elimination include
double elimination brackets with a "true third" match which includes an
additional match contested between the competitors ranked 2nd (loser of
the finals) and 3rd (winner of the consolation bracket) at the end of
the tournament *if they have not already competed against each other*.
There is also a variation of double elimination where only those
competitors who have lost to a competitor that makes the finals are
placed into the consolation bracket and given the chance to compete for
third place. This style of competition is used in Olympic wrestling
(both freestyle and Greco-Roman)[^2].

[^2]: As in Olympic judo, Olympic wrestling also award two bronze medals
    to the winners of each side of the consolation bracket

<!-- A paragraph about Swiss, Dutch, and Monrad systems -->

In bracket style competitions, only a handful of the match-ups that are
possible actually occur. When every possible competitive pairing occurs,
this is referred to as a round-robin structure where every competitor
plays every other competitor the same number of times. Single
round-robin, where was possible match-up occurs exactly once, is used
in, for examples, the group stage of the FIFA World Cup to create a
ranking to determine which teams advance to the next round. Double
round-robin, where each possible pairing occurs twice, is used in
association football in English Premier League (EPL) and in chess during
the FIDE candidates tournament (a tournament that decides who will
challenge the world champion). For the sake of brevity, we review only a
small number of possible competition structures, but the interested
reader can find a more complete list of competition structures in
@Goosens2024, @Appleton1995-zn and @Csato2021.

A natural question of interest after reviewing different types of
competition structures is: which competition structure is best? As with
most complicated questions, the answer is "it depends". One commonsense
conceptualization for the best competition structure is that the final
rankings produced by a competition structure should closely match the
true rankings of the strength of the competitors. That is, a good
competition structure should rank the true best team first, the true
second best team second, and so on, until finally ranking the true worst
team last. We refer to this property of a competition structure as the
efficacy of the structure (@Lasek2018-de, @Sziklai2022-kf). If one were
only interested in correctly ranking the true best competitor as first,
with the ranks of the remaining competitors viewed as inconsequential,
we refer to this here are effectivity (@Glenn1960-mn). While there are
many competition structures where efficacy and effectivity are useful
measures[^3], there are many other competitions where the correct
ranking of more than one competitor, but not all the competitors is of
import. For instance, in Olympic competition, the goal of a competition
structure is to correctly rank the top 3 competitors to award gold,
silver, and bronze medals to[^4]. In many American sports, for instance
the NFL, the goal of the competition structure of the regular season is
to rank the top 7 teams in each conference to choose the playoff teams
with the ranking of the teams 8 through 16 in each conference
inconsequential in determining a Super Bowl champion[^5]. We coin the
term "salient ranks" to refer to the ranks of a competition structure
that matter. That is the ranks 1, 2, and 3 in Olympic competition are
the salient ranks in that competition structure. In this notation,
efficacy is a measure with with salient ranks 1 through $n$ where $n$ is
the number of competitors, and effectivity is a measure where 1 is the
only salient rank.\
<!-- Efficacy: @Lasek2018-de, @Sziklai2022-kf -->
<!-- Effectivity: @Glenn1960-mn -->

[^3]: efficacy is a useful measure for the EPL, for example, as the
    ranking of the teams matters at both the high end (i.e. entry into
    special tournaments) and the low end (i.e. relegation status) and
    effectivity is a useful measure for the NCAA College Basketball
    tournament (i.e. main goal is to crown a champion)

[^4]: Some sports award two bronze medals (e.g. wrestling, judo, boxing,
    tae kwon do)

[^5]: The ranking of teams outside of the playoffs does matter for draft
    purposes though

<!-- From Goosens 2024: The ability of a tournament to rank contestants according to their true strength is called efficacy (Lasek and Gagolewski, -->

<!-- 2018; Sziklai et al., 2022). Efficacy with respect to only the best player is also known as effectivity (Glenn, 1960) -->

<!-- or predictive power (Ryvkin and Ortmann, 2008; Vu and Shoham, 2011). -->

If the only goal of a competition structure was high efficacy of the
salient ranks, the best way to achieve this would be to play each pair
of possible match ups of competitors as many times as possible. Of
course, this is impossible due to many factors including time
constraints, cost constraints, and wear and tear on competitors in high
impact sports (e.g. boxing, American football, etc.). In addition,
competitions where the true best team is guaranteed or nearly guaranteed
to win may be less appealing to fans and may drive down demand as
discussed in @Johnson2022-wz.

In light of this, many different competition structures have been
proposed, several of which are mentioned previously. Given these many
competition structures, ideally there would be a way comparing the
performance of different competition structures to each other.

Here, we propose a novel metric for evaluating different competition
structures using an information theoretic approach. As an analogy, we
view the results of a tournament as a transmitted message, and we use
information theory to assess how accurately the message (i.e. the
tournament results) matches the true transmission (i.e. the true ranking
of competitors).

Specifically, we use weighted mutual information with a weights based on
squared error loss[^6] between the ranks generated by a competition
structure and the true ranks of the competitors. Additionally, we
include an indicator vector to specify the salient ranks, allowing us to
compare different competition structures with different choices for
salient ranks.

[^6]: other loss functions can easily be incorporated into our method

The remainder of this manuscript presents the mathematical foundation of
the proposed measure followed by a simulation study estimating out
proposed measure for different types of competition structures with
varying choices of salient ranks. We close by offering a summary of the
work and ideas for future directions.

<!-- @Marchand2002-tq only cares about the probability of winning in seeded vs random draw knockout tournemnts.   -->

<!-- @Johnson2022-wz: Match outcome uncertainty and the Rottenberg hypothesis -->

<!-- Different competition structures have unique strengths and weaknesses, affecting how well they reflect the true rankings of the teams. The goal of a tournament is to find the best teams, but how can we quantify how well a tournament performs? In some instances, the tournament organizers only award the overall winner, but other times, the top three or ten teams are awarded. Because of this, organizers may prefer one structure over another based on their needs. Tournament organizers must also take into account factors like cost, timeliness, entertainment value, and fairness when choosing a tournament structure. -->

<!-- To evaluate a tournament’s effectiveness, we propose a numerical metric that quantifies how accurately it orders teams based on their true rankings. Tournament results can be viewed as a "message" attempting to convey the true team rankings. However, various factors introduce noise, leading to information loss. We use principles from information theory to measure this information loss across multiple tournament simulations to assess the reliability of different formats. -->

# Methods

Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with
an associated true strength parameter
$\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that
$(\theta_1 > \theta_2 > \ldots > \theta_n)$. Next, let
$r(\boldsymbol{\theta})$ be the indexes of $\boldsymbol{\theta}$ when
$\theta$ is sorted from largest to smallest so that
$r(\boldsymbol{\theta}) = \{1, 2, \ldots, n-1, n\}$. Next define
$T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure
$s$. $\phi$ contains all the information about the scheduling of teams,
which could be fully known prior to the tournament (*e.g.* round robin)
or determined as the tournament progresses (*e.g.* single elimination
tournament). We then let $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be an
$n$-dimensional vector-valued random variable that gives the results of
a tournament as a vector of the indexes of the vector
$\boldsymbol{\theta}$. The index in the first position of the vector
$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the
team that finished first, the index in the second position of the vector
$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the
team that finished second, and so on.

For example, in a $n$ = 4 team tournament,
$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) indicates that
the team with index 3 (i.e. the true third best team) won the
tournament, the true second best team finished second, the true best
team finished third and the true 4th team finished 4th in that
particular tournament. If $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ =
(1, 2, 3, 4), this means that the random outcome of the tournament was
the same as the true ordering of the teams.

If one views the true ranking of the teams in a tournament as a message
to be sent to a receiver and the outcome of the tournament as a message
that is received, we can measure the "goodness" of a tournament in terms
of its ability to accurately transmit the true ranking. We can then use
concepts from information theory to assess the ability of a tournament
to correctly rank teams. Specifically, we start with the concept of
mutual information (@Guiasu1977-nu). For two random variables $X$ and
$Y$ mutual information is defined as:

$$
I(X;Y)=\sum_{y\in Y}\sum_{x\in X}p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}
$$

In our setting, we replace $X$ and $Y$ with $r(\boldsymbol{\theta})$ and
$\hat{r}(T(\phi, s),\boldsymbol{\theta})$, and we could seek to estimate
$I(r(\boldsymbol{\theta}),\hat{r}(T(\phi, s),\boldsymbol{\theta}))$. For
simplicity, we drop the arguments from the functions $r$ and $\hat{r}$
for ease of exposition.

We note here that $r$ is not a random variable, however, since the
indexing of the teams in the vector $\boldsymbol{\theta}$ is arbitrary
(we use 1, 2, 3, etc. for convenience, but any indexing is valid), we
can view this a random variable where all permutation so of the $n$
teams are equally likely. Therefore when calculating the mutual
information in this setting, we set $p(r) = \frac{1}{n!}$. By a similar
argument we set $p(\hat{r}) = \frac{1}{n!}$.

We define the mutual information of $r$ and $\hat{r}$ to be:

$$
I(r,\hat{r})=\sum_{r}\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})} = n!\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{(\frac{1}{n!})^2}
$$

While there are $n!$ different permutations for the result of $r$, we
don't need to sum across these as any specific choice of $r$ is just an
arbitrary labeling of the true strength parameters vector
$\boldsymbol{\theta}$. So given $r$, the distribution of $\hat{r}$ is
the same up to the labeling. Therefore, we only need to consider a
single permutation of $r$, we compute the the quantity inside the
summation, and then multiple by $n!$ (effectively summing across $r$).

In order to compute this quantity, we need to compute $p(r,\hat{r})$.
This probability is found by calculating the probability of a given
permutation of $\hat{r}$ and $r$ is assumed to be the permutation from
$\{1, 2, \ldots, n\}$. We estimate $p(r,\hat{r})$ empirically through
simulation.

However, using mutual information in this form does not suit our needs
in this setting. The problem is that mutual information in this form
will yield high values of mutual information when the output from the
tournament $\hat{r}$ is highly consistent *even if the ranking from the
tournament is incorrect*. As an example, if $n=4$ and the true order of
$\boldsymbol{\theta}$ is $r = \{1, 2, 3, 4\}$ and
$p(\hat{r} = \{4, 3, 2, 1\}) = 1$ will be the same mutual information as
when $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{1, 2, 3, 4\}) = 1$ and in
both of these cases the mutual information will be maximized at:

$$
I(r,\hat{r})= n!\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{\left(\frac{1}{n!}\right)^2} = n! log((n!)^2)
$$ and for the specific case when $n=4$ would be: $$
4! log_2(4!^2) = 24*log_2(24^2) = 220.0782
$$

In order to alleviate this problem, we instead consider *weighted*
mutual information (@Guiasu1977-nu). We want to give more weight to
permutations from $\hat{r}$ that are "closer" to $r$.

In general, any loss function $l(r, \hat{r})$ can be used to define the
weigting function:

$$
w(r, \hat{r}) =
\begin{cases}
\frac{1}{l(r, \hat{r})}, & r \ne \hat{r} \\
1, & r = \hat{r}
\end{cases}
$$

Here though, we choose to use squared error loss:

$$
w(r, \hat{r}) =
\begin{cases}
\frac{1}{\hat{r}'r}, & r \ne \hat{r} \\
1, & r = \hat{r}
\end{cases}
$$ <!-- Note that one is 1/N! -->

<!-- And this does work because consistently getting the order incorrect is perfect information. -->

However, because the number of teams that need to be accurately ordered
differs based on the choice of salient ranks, a second set of weights is
defined, $m$, is used to indicate which ranks are salient. We call these
the "salient weights" to differentiate from the other weighting
function. As an example, for a four-team tournament, if the organizer
decides that only the winner is important, then $m = \{1,0,0,0\}$ (i.e.
essentially measuring effectivity). While here we only consider
indicators (i.e. 0's or 1's) in the salient weights vector, there is no
reason why you couldn't weight the ranks by relative levels of
importance. For instance, letting $m = \{3,1,0,0\}$ indicates that the
salient ranks are 1 and 2, but getting the true best competitor as the
winner is more important than ranking the true second best competitor
second.

With the addition of the weighting functions, the formula for mutual
information becomes:

$$
wI_m(r, \hat{r})=\sum_{r}\sum_{\hat{r}} w(r,\hat{r}) * m * p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})}
$$

Because this mutual information is unitless, the formula can be
standardized to be between 0 and 1 as follows: $$
\frac{wI_m(r,\hat{r})}{\text{max\{}H(r), H(\hat{r})\}}
$$

where $H(.)$ is the usual Shannon entropy. Competition structures with
higher levels of efficacy are indicated when our measure is closer to 1.
Likewise, competition structures that are near 0 by our proposed metric
indicate low levels of efficacy.

<!-- $$ -->

<!-- max\{H(X), H(Y)\} = \sum_{r}  p(r) \log p(r) = n! \log n! -->

<!-- $$ -->

<!-- # ryan way  -->

<!-- Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with an associated true strength parameter $\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that $(\theta_1 < \theta_2 < \ldots < \theta_n)$.  Next, let $r(\boldsymbol{\theta})$ be the rank of the vector $\boldsymbol{\theta}$ from largest to smallest so that $r(\boldsymbol{\theta}) = \{n, n-1, \ldots, 2, 1\}$.  Next define $T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure $s$.  $\phi$ contains all the information about the scheduling of teams, which could be fully known prior to the tournament (i.e. round robin) or determined as the tournament progresses (i.e. single elimination tournament).  We then let  $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be a vector valued random variable that gives the results of a tournament as a vector of ranks.   -->

<!-- Figure out what $\hat{r}$ should look like.   -->

<!-- Let $\boldsymbol{x} = (x_1, x_2, \ldots, x_n)$ be the permutation of indexes of $\boldsymbol{\theta}$ such that $x_1 = {i | \theta_i = max(\boldsymbol{\theta})}$ -->

# Results

```{=html}
<!--

| Structure/Ranks | 1 | 2 | 3 | 4 | 5 | 6 | 7 | 8 |
|:----------------------:|:----:|:----:|:----:|:----:|:----:|:----:|:----:|:----:|
| Double Elimination | 0.4067 | 0.2517 | 0.1443 | 0.0916 | 0.0532 | 0.0390 | 0.0340 | 0.0290 |
| Single Elimination | 0.4023 | 0.2439 | 0.1338 | 0.0837 | 0.0478 | 0.0356 | 0.0312 | 0.0270 |
| Equal Strengths Single Elimination | 0.1297 | 0.0474 | 0.0224 | 0.0129 | 0.0072 | 0.0051 | 0.0050 | 0.0050 |
| Bad Seeding Single Elimination | 0.2679 | 0.0146 | 0.0102 | 0.0059 | 0.0032 | 0.0022 | 0.0020 | 0.0026 |
| Every Possible Permutation | 0.1036 | 0.0302 | 0.0113 | 0.0047 | 0.0018 | 0.0008 | 0.0011 | 0.0014 |
| Round Robin | 0.8851 | 0.8473 | 0.8096 | 0.7708 | 0.7337 | 0.7013 | 0.6755 | 0.6697 |
| Worst Possible | 0 | 0 | 0 | 0 | 0 | 0 | 0 | 0 |
| Best Possible | 1 | 1 | 1 | 1 | 1 | 1 | 1 | 1 |

-->
```

We examined the following structure:

-   Single elimination, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7)
    and normally distributed strengths

-   Single elimination, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7)
    and equal strengths (effectively each game is determined by a coin
    toss)

-   Single elimination, bad seeding (1 vs 2, 3 vs 4, 5 vs 6, 7 vs 8) and
    normally distributed strengths

-   Every Possible Permutation: One simulation for each $8! = 40320$
    possible combinations that final rankings may occur

-   Double elimination, usual seeding (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7)

-   Round Robin, normally distributed strengths

In order to fairly compare tournament structures, we assume the true
strengths of the teams follow a normal distribution, where each team is
assigned an equally spaced quantile and each team's strength is the
z-score of its respective quantile. Using this assumption, 10,000
simulations, all with 8 total teams, were ran for a single game round
robin (each team plays each other once), a single elimination structure
with the usual seeding structure (1 vs 8, 4 vs 5, 3 vs 6, 2 vs 7), a
single elimination structure with a poor seeding structure (1 vs 2, 3 vs
4, 5 vs 6, 7 vs 8). For a better understanding of the results, the best
possible results (every structure has the best in rank 1, the 2nd best
in rank 2, etc), the worst possible results (every structure has the
worst in rank 1, the 2nd worst in rank 2, etc), every possible
permutation of the ranks once, and a single elimination tournament with
the usual seeding structure where every team is of equal strengths.
Because the metric found is unitless, we were able to normalize the
values using the best and worst results to make the range from 0 to 1,
with 0 as the worst and 1 as the best. After normalizing, we get the
following graph (note the y-axis is on a log10 scale):

![This chart compares the normalized mutual information (on log10 scale)
of various tournament structures against the number of top teams being
considered, where the team strengths are equally distanced based on a
normal distribution.
\label{fig:fig1}](images/Normal_Information_Curve_Graph.png){width="408"}

Based on the Figure 1 \ref{fig:fig1} , the intuitive best tournament,
round robin, does show to be the best. It is important to note that, due
to the weighting structure, the worst possible score approaches 0 as the
$i$ number of teams increases, but never reaches 0. So, the plot is
normalized so that the worst possible is 0 and the best possible is 1.
This explains the increase for a few of the lines at the end of the
plot. If there is a poor seeding structure, as seen in the graph, it can
drastically impact the effectiveness of the tournament. In the extreme
case seen, the tournament can actually become worse than deciding games
by random chance (like flipping a coin). Another important discovery is
that the single elimination structure with equal strengths is better
than every possible permutation. This is likely because, for a single
elimination structure, teams that play in the first round cannot both
advance to the next round in each simulation, leading to slightly more
information. Additionally, the bad seeding structure overlaps with the
equal strengths single elimination. This shows that, given that the best
team is the 1-seed, even a poorly structured single elimination
structure still results in the best team winning more than by random
chance. However, because the 2-seed plays the 1-seed in the first round,
the poor structure results in a less effective result because the top 2
can never be truly accurate according to the known strengths.

![Plot to further explore the difference between single elimination and
double elimination competition structures
\label{fig:fig2}](images/tournament_comparison_plot_se_de.png){width="409"}

Double elimination is slightly better than single elimination, with the
gap growing wider as the top number of teams looked at increases.
However, those are under the assumption that the true strengths were
ordered correctly in the seeding. While double elimination was expected
to be slightly better than single elimination, the increase in the gap
is likely due to noise that comes with how ties are handled. Ties led to
randomly assigning teams to a ranking, but with double elimination,
there were less ties that had to be dealt with, leading to more accurate
rankings.

# Conclusion {#sec:conclusion}

This study introduced a weighted mutual information framework to
evaluate the effectiveness of different tournament formats in preserving
the true underlying rankings of teams. By simulating tournaments with
varying seeding structures and applying an information-theoretic
approach, we quantified how accurately each format conveys ranking
information. Our findings confirm the intuitive advantage of round robin
tournaments, which consistently outperformed single and double
elimination formats in ranking accuracy. Additionally, we observed that
seeding plays a critical role: poor seeding structures can significantly
degrade performance, in some cases making outcomes less informative than
random chance.

These results underscore the trade-off between tournament accuracy and
practical constraints such as time, cost, and entertainment value. While
round robin offers the highest fidelity to true rankings, it is often
impractical for large tournaments, highlighting the need for hybrid or
adaptive structures that balance accuracy with logistical feasibility.

Future work should extend this analysis to scenarios with more
tournament structures, more teams, incorporate probabilistic strength
models reflecting real-world uncertainty, and explore alternative
weighting schemes for different competitive priorities (e.g., top-4
accuracy versus full ranking). Applying these methods to actual
tournament data could further validate their usefulness for organizers
aiming to design fair and informative competitions.

# Acknowledgements {.unnumbered}

We thank the Department of Mathematics and Statistics at Loyola
University Chicago for their support and resources in conducting this
study. We thank Kailey Marie Lum for the suggestion of the term "salient
ranks". No external funding was received for this research.

# Supplementary Material {.unnumbered}

All supplementary material available at
<https://github.com/gjm112/tournaments>.

# References