### 2025-10-10 CMSAC_presentation.Rmd
CMSAC presentation

---
title: "Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information"
author: "Zach Culp, Josie Peterburs, Dr. Gregory Matthews, Dr. Ryan McShane"
date: "`r Sys.Date()`"
bibliography: references.bib
fontsize: 12pt
output:
  beamer_presentation:
    theme: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tinytex)
library(tidyverse)
load("curves_data.RData")
#load("tournament_comparison_plot.png")
#load("tournament_comparison_plot_close.png")
```

## Introduction
<!-- Overall comments 
I think the presentation is too long for 12 minutes (+ questions), but
I feel like I didn't cover everything we want. What is plausable to cover
in a 12-minute period? 
Also, we need to figure out the notation for the final metric because I think
the weight should have the subscript m, not I. And I probably was not consistent
with the usage of theta or calling it true rank/true strength. -->

-   There exists numerous competition formats across all professional sports
-   Which competition structure is best?
-   The "best" competition structure depends on how well it reflects the competitors' true strengths
-   We will propose a metric that quantifies how effectively different tournament structures convey information about competitors' true strengths

<!-- Should we talk about Efficacy vs Effectivity here -->

## Game Simulation

-   Given the assigned "known strength", $\theta$, we simulate a game using the Bradley-Terry Model

$$\Pr(i > j) = \frac{e^{\theta_i}}{e^{\theta_i} + e^{\theta_j}} = \frac{1}{1+e^{\theta_i - \theta_j}}$$

## Tournament Simulations
- Each competition structure is simulated given the number of teams, $n$,  and $\theta$
- Below is an example of the output from one simulation of a single elimination tournament structure

```{r}
renamed_df <- all_results_norm %>% 
  select(-c(simulation, distribution, seed, teams)) %>%
  rename(theta = true_strength, r = true_rank, r_hat = rank_hat) %>%
  mutate(theta = round(theta,3))

head(renamed_df,8)
```

## Mutual Information

- If one views the true ranking of the teams in a tournament as a message to be sent to a receiver and the outcome of the tournament as a message that is received, we can measure the “goodness” of a tournament in terms of its ability to accurately transmit the true ranking.

- The notation for mutual information is: 
$$ I(X;Y) = \sum_{y \in Y}\sum_{x \in X} p(x,y) \log{\frac{p(x,y)}{p(x)p(y)}} $$
- In our case, it can be simplified to:
$$ I(r, \hat{r}) = n! \sum_{\hat{r}} p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{(\frac{1}{n!})^2}} = n! \log{((n!)^2)} $$

## Weighted Mutual Information
- A limitation of mutual information is that it does not distinguish between the direction of information
- In other words, a ranking of $(1,2,3,4)$ is equivalent to $(4,3,2,1)$
- Because of this, we proposed a weighting function $w(r, \hat{r})$
- So, the new formula is:
$$ wI_m(r,\hat{r}) = \sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}} $$

## Weighting Function
- In our original paper submission, we used squared error loss as the weighting function:
$$ w(r, \hat{r}) =
\begin{cases}
\frac{1}{\hat{r}'r}, & r \neq \hat{r} \\
1, & r = \hat{r}
\end{cases} $$
- However, after comments and reviews from people at CSAS and CMSAC, we have decided to use a weighted Kendall's Tau

## Weighted Kendall
$$ \tau_w = \frac{\sum_{i\neq j} W_{ij} \times \text{sgn}(i-j)\times \text{sgn}(R_i - R_j)}{\sum_{i,j} W_{ij} - \sum_i W_{ii}}$$

- $W$ is a $n\times n$ lower-triangular matrix, where

<!-- Should this be capital W??? -->

$$W_{ij} = w_{\min(i,j)}, \text{ for } i,j = m,2m, \dots, n\times m$$

- $m$: vector of length $n$ representing the "salient weights", or the importance of correctly getting each ranking

- If $\tau_w = 1$, then the ranking is identical to the true ranking; if $\tau_w = 0$, then the order of teams is perfectly inversely proportional to the true rankings

## Weighted Kendall Example
- If $\hat{r} = (2,1,3,4)$ and m = $(3,1,1,0)$, then:
$$ W = \begin{pmatrix}
3 & 3 & 3 & 3 \\
3 & 1 & 1 & 1 \\
3 & 1 & 1 & 1 \\
3 & 1 & 1 & 0
\end{pmatrix} $$

$$\sum_{i,j} W_{ij} = 29; \sum_{i} W_{ii} = 5; \sum_{i\neq j} W_{ij} \times \text{sgn}(i-j)\times \text{sgn}(R_i - R_j) = 18 $$

- So, $\tau_w$ for this example would be $\frac{18}{29-5} = 0.75$

## Normalizing Mutual Information
- Mutual Information can be normalized by dividing by the joint entropy

- Entropy: $-p(x)\log{p(x)}$

- Assuming $r$ and $\hat{r}$ are independent, joint entropy can be written as $H(r, \hat{r}) = H(r) + H(\hat{r})$

- $H(r)$ and $H(\hat{r})$ are theoretically equivalent, leaving $H(r,\hat{r}) = -2\sum_{r} p(r)\log(p(r)) = -2(n!) \times (1/n!) \times\log_2(1/n!) = \log_2((n!)^2)$
<!-- Is this true?? I can't remember if they are equivalent or their maximums were equivalent -->

- So, the final formula for the metric is:
$$ \frac{wI_m(r,\hat{r})}{H(r) + H(\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$

## Results

<!-- Should these plots be on separate slides? -->
```{r out.width="49%", fig.show='hold', fig.align='center'}
knitr::include_graphics("tournament_comparison_plot.png")
knitr::include_graphics("tournament_comparison_plot_close.png")
```

## Limitations
- The "known" strengths are assumptions that can be estimated but never truly known
- Ties are not addressed in any of the games
- Ties in the $\hat{r}$ are randomly assigned

## Future Work
- Evaluate structures using various distributions and "known" weights
- Estimate the $\theta$s in a real-world example and compare recent changes to the old format

## References
<!-- I don't know if the references.bib would work here or not -->
- Appleton, David R. 1995. “May the Best Man Win?” Statistician 44 (4): 529.
- Csato, L. 2021. Tournament Design: How Operations Research Can Improve Sports Rules.
1st Ed. Cham, Switzerland: Palgrave Macmillan.
- Devriesere, Csató, K., and D. Goossens. 2025. “Tournament Design: A Review from an
Operational Research Perspective.” European Journal of Operational Research 324 (1):
1–21.
- Glenn, W A. 1960. “A Comparison of the Effectiveness of Tournaments.” Biometrika 47
(3-4): 253–62.

## References
- Guiasu, Silviu. 1977. Information Theory with Applications. New York, NY: McGraw-Hill.
- Johnson, Sidney, and Rodney Fort. 2022. “Match Outcome Uncertainty and Sports Fan
Demand: An Agnostic Review and the Standard Economic Theory of Ports Leagues.”
Int. J. Emp. Econ. 01 (02).
- Lasek, Jan, and Marek Gagolewski. 2018. “The Efficacy of League Formats in Ranking
Teams.” Stat. Modelling 18 (5-6): 411–35.
- Sziklai, Balázs R, Péter Biró, and László Csató. 2022. “The Efficacy of Tournament Designs.”
Comput. Oper. Res. 144 (105821): 105821.