### 2025-10-10 josie slides.qmd
update intro slides

---
title: "Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information"
author: "Zach Culp, Josie Peterburs, Dr. Gregory J. Matthews, Dr. Ryan P.A. McShane"
format: revealjs
editor: visual
date: "`r Sys.Date()`"
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tinytex)
library(tidyverse)
library(combinat)
load("curves_data.RData")
#load("tournament_comparison_plot.png")
#load("tournament_comparison_plot_close.png")
```

##

```{r}
knitr::include_graphics("4 bracket good.png")
```

##

:::: {.columns}

::: {.column width="50%"}
Good 
:::

::: {.column width="50%"}
Bad
:::

::::

```{r}
knitr::include_graphics("4 seeded tourn w arrow.png")
```

::: {.notes}
- 4 team single elimination tournament:
  - Dream Team vs Work Group vs Middle School Girls Team vs Babies
- If we only care about the top team winning, seeding structure does not matter
- If we care about who gets second and third, seeding becomes important
  - Would not pair up Work Group and Dream Team in the first round
:::

## Comparing Structures

- Which competition structure is best at reflecting the true strengths of competitors?

- **Efficacy**^[Lasek and Gagolewsk (2018), Sziklai, Biró, and Csató (2022)]: Final rankings should match the true rankings of the strength of the competitors

- **Effectivity**^[Glenn (1960)]: Interested in correctly ranking the best competitor as first

## Efficacy and Effectivity

- Decide how many rankings are important
  - Olympic competition ranks the top 3 competitors
  - Premier League: top 4 (Champions League) and bottom 3 (Relegation)

- **Salient ranks**: the ranks that matter
  - Efficacy: 1 through $n$ matter
  - Effectivity: 1 is only salient rank

## Goal

- Want to quantify how well a structure can accurately order the competitors
- We will propose a metric that quantifies how effectively different tournament structures convey information about competitors’ true strengths

## Distribution of $\hat{r}$

```{r}
ranks <- seq(1:3)
permn(ranks) %>% unlist() %>% matrix(ncol = 3, byrow = TRUE)

```