### 2025-10-17 CMSAC_presentation/josie_slides.qmd
josie slides but having issues

---
title: "Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information"
author: "Zach Culp, Josie Peterburs, Dr. Gregory J. Matthews, Dr. Ryan P.A. McShane"
format: revealjs
editor: visual
date: "`r Sys.Date()`"
bibliography: ../references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tinytex)
library(tidyverse)
library(combinat)
# load("curves_data.RData")
#load("tournament_comparison_plot.png")
#load("tournament_comparison_plot_close.png")
```

##

::::: columns
::: {.column width="50%"}
```{r}
knitr::include_graphics("4 bracket good.png")
```
:::

::::: columns
::: {.column width="50%"}
```{r}
knitr::include_graphics("Mens-Basketball-Olympics-Getty-H-2024.jpeg")
knitr::include_graphics("five_rec_league_men_playing_basketball.png")
knitr::include_graphics("jv_girls_basketball.png")
knitr::include_graphics("five_babies_playing_basketball.png")

```
:::
:::::

##

::::: columns
::: {.column width="50%"}
Good
:::

::: {.column width="50%"}
Bad
:::
:::::

```{r}
knitr::include_graphics("4 seeded tourn w arrow.png")
```

::: {.notes}
-   4 team single elimination tournament:
    -   Dream Team vs Work Group vs Middle School Girls Team vs Babies
-   If we only care about the top team winning, seeding structure does not matter
-   If we care about who gets second and third, seeding becomes important
    -   Would not pair up Work Group and Dream Team in the first round
:::

## Comparing Structures

-   Which competition structure is best at reflecting the true strengths of competitors?

-   **Efficacy**[^1]: Final rankings should match the true rankings of the strength of the competitors

-   **Effectivity**[^2]: Interested in correctly ranking the best competitor as first

[^1]: Lasek and Gagolewsk (2018), Sziklai, Biró, and Csató (2022)

[^2]: Glenn (1960)

## Efficacy and Effectivity

-   Decide how many rankings are important
    -   Olympic competition ranks the top 3 competitors
    -   Premier League: top 4 (Champions League) and bottom 3 (Relegation)
-   **Salient ranks**: the ranks that matter
    -   Efficacy: 1 through $n$ matter
    -   Effectivity: 1 is only salient rank

## Goal

-   Want to quantify how well a structure can accurately order the competitors
-   We will propose a metric that quantifies how effectively different tournament structures convey information about competitors’ true strengths

## Distribution of $\hat{r}$

```{r}
ranks <- seq(1:3)
permn(ranks) %>% unlist() %>% matrix(ncol = 3, byrow = TRUE)

```