### 2025-10-17 CMSAC_presentation/Zach_slides.qmd

---
title: "Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information"
author: "Zach Culp, Josie Peterburs, Dr. Gregory J. Matthews, Dr. Ryan P.A. McShane"
format: revealjs
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tinytex)
library(tidyverse)
load("curves_data.RData")
#load("tournament_comparison_plot.png")
#load("tournament_comparison_plot_close.png")
```

## Mutual Information

::: notes
If one views the true ranking of the teams in a tournament as a message to be sent to a receiver and the outcome of the tournament as a message that is received, we can measure the "goodness" of a tournament in terms of its ability to accurately transmit the true ranking.
:::

-   The notation for mutual information is: $$ I(X;Y) = \sum_{y \in Y}\sum_{x \in X} p(x,y) \log{\frac{p(x,y)}{p(x)p(y)}} $$

-   In our case, it can be simplified to:

    $$ I(r, \hat{r}) = n! \sum_{\hat{r}} p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{(\frac{1}{n!})^2}} = n! \log{((n!)^2)} $$

## Weighted Mutual Information

-   A limitation of mutual information is that it does not distinguish between the direction of information
-   In other words, given equal probabilities, a ranking of $(1,2,3,4)$ is equivalent to $(4,3,2,1)$
-   Because of this, we proposed a new weighting function $w(r, \hat{r})$\
    $$ wI_m(r,\hat{r}) = \sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}} $$

## Weighting Function

-   In our original paper submission, we used squared error loss as the weighting function: $w(r, \hat{r}) = \begin{cases} \frac{1}{(r-\hat{r})^2}, & r \neq \hat{r} \\ 1, & r = \hat{r} \end{cases}$
-   However, after comments and reviews from people at CSAS and CMSAC, we have decided to use a weighted Kendall's Tau

## Weighted Kendall

$$ \tau_w = \frac{\sum_{i\neq j} W_{ij} \text{ sgn}(i-j) \text{ sgn}(R_i - R_j)}{\sum_{i,j} W_{ij} - \sum_i W_{ii}}$$

-   If $\hat{r} = (2,1,3,4)$, then:

$$\sum_{i\neq j} W_{ij} \text{ sgn}(i-j) \text{ sgn}(R_i - R_j) = 10$$

-   So, $\tau_w$ for this example would be $\frac{10}{15-3} = 0.833$

## Top K-teams Weighted Kendall

```{r}
rhat_df %>% 
  select(-squared_error) %>%   
  filter(permutation %in% c("1234", "4321", "1324", "2143")) %>%   
  arrange(permutation,k) %>%   
  rename("Weighted Kendall" = weighted_kendall)
```

$$ \tau_w = \frac{\sum_{i\neq j} W_{ij} \text{ sgn}(i-j) \text{ sgn}(R_i - R_j)}{\sum_{i,j} W_{ij} - \sum_i W_{ii}}$$

## Comparing Weighted Kendall's Tau to Squared Error

```{r}
rhat_df %>%    
  filter (k==4) %>%   
  select(-k) %>%   
  arrange(desc(weighted_kendall)) %>%    
  rename("Weighted Kendall"= weighted_kendall, "Squared Error" = squared_error)
```

## Normalizing Mutual Information

-   To make comparisons across different values of $n$, we normalize by the joint entropy

-   So, the final formula for the metric is:

$$\frac{wI_m(r,\hat{r})}{H(r,\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$

-   Assuming $r$ and $\hat{r}$ are independent, joint entropy can be written as $H(r, \hat{r}) = H(r) + H(\hat{r})$

## Tournament Simulations

-   Given the assigned "known strength", $\theta$, we simulate a game using the Bradley-Terry Model: $\Pr(i > j) = \frac{e^{\theta_i}}{e^{\theta_i} + e^{\theta_j}} = \frac{1}{1+e^{\theta_i - \theta_j}}$
-   Below is output from one simulation of a single elimination tournament structure

```{r}
renamed_df <- all_results_norm %>%    
  select(-c(simulation, distribution, seed, teams)) %>%   
  rename(theta = true_strength, r = true_rank, r_hat = rank_hat) %>%   
  mutate(theta = round(theta,3))  
head(renamed_df,8)
```

## Results

```{r out.width="49%", fig.show='hold', fig.align='center'}
knitr::include_graphics("tournament_comparison_plot.png")
```

## Close-Up Results

```{r out.width="49%", fig.show='hold', fig.align='center'}
knitr::include_graphics("tournament_comparison_plot_close.png")
```

## Limitations

-   The "known" strengths are assumptions that can be estimated but never truly known
-   Ties are not addressed in any of the games
-   Ties in the $\hat{r}$ are randomly assigned

## Future Work

-   Evaluate structures using various distributions and "known" weights
-   Estimate the $\theta$s in a real-world example and compare recent changes to the old format

## References

<!-- I don't know if the references.bib would work here or not -->

-   Appleton, David R. 1995. “May the Best Man Win?” Statistician 44 (4): 529.
-   Csato, L. 2021. Tournament Design: How Operations Research Can Improve Sports Rules. 1st Ed. Cham, Switzerland: Palgrave Macmillan.
-   Devriesere, Csató, K., and D. Goossens. 2025. “Tournament Design: A Review from an Operational Research Perspective.” European Journal of Operational Research 324 (1): 1–21.
-   Glenn, W A. 1960. “A Comparison of the Effectiveness of Tournaments.” Biometrika 47 (3-4): 253–62.

## References

-   Guiasu, Silviu. 1977. Information Theory with Applications. New York, NY: McGraw-Hill.
-   Johnson, Sidney, and Rodney Fort. 2022. “Match Outcome Uncertainty and Sports Fan Demand: An Agnostic Review and the Standard Economic Theory of Ports Leagues.” Int. J. Emp. Econ. 01 (02).
-   Lasek, Jan, and Marek Gagolewski. 2018. “The Efficacy of League Formats in Ranking Teams.” Stat. Modelling 18 (5-6): 411–35.
-   Sziklai, Balázs R, Péter Biró, and László Csató. 2022. “The Efficacy of Tournament Designs.” Comput. Oper. Res. 144 (105821): 105821.