### 2025-10-22 single_elimination_shiny.qmd

---
title: "single_elimination"
format: 
  html:
    embed-resources: true
editor: visual
---

```{r}
library(tidyverse)
library(dplyr)
library(shiny)
```

### Things To Do

-   Add a re-seeded option

-   Custom strengths (likely as an input to the function)

# Simulate a Single Match

```{r}
# Function to simulate a match using Bradley-Terry model
simulate_match <- function(team1, team2, strength1, strength2) {
  if (is.na(team1)) {
    return(team2)
  } else if (is.na(team2)) {
    return(team1)
  } else {
    p <- exp(strength1 - strength2) / (1 + exp(strength1 - strength2))
    return(ifelse(runif(1) < p, team1, team2))  # Randomize outcome
  }
}

```

# Single Elimination Tournament Structures

## Creating the traditional tournament structure for n teams

```{r}
generate_elimination_tournament_structure <- function(n) {

  is_power_of_two <- function(x) {
    x > 0 && (x & (x - 1)) == 0
  }

  # If n is not a power of 2, round up to the next power of 2 and add placeholders
  if (!is_power_of_two(n)) {
    next_power_of_two <- 2^ceiling(log2(n))
    extra_teams <- next_power_of_two - n
    #message(sprintf("Rounding up to %d teams. Adding %d NA placeholders.", next_power_of_two, extra_teams))
    n <- next_power_of_two
  } else {
    extra_teams <- 0
  }

  rounds <- log(n, base = 2) - 1
  teams <- c(1, 2)  # Initial list of teams

  for (i in 1:rounds) {
    teams <- nextLayer(teams)
  }

  left_half <- teams[1:(length(teams)/2)]
  right_half <- teams[(length(teams)/2+1):length(teams)]
  right_half <- rev(right_half)
  teams <- c(left_half,right_half)

  # Ensure that the lower number team is on the left side of each matchup
  teams <- unlist(lapply(seq(1, length(teams), by = 2), function(i) {
    if (teams[i] > teams[i + 1]) {
      c(teams[i + 1], teams[i])  # Swap if needed
    } else {
      c(teams[i], teams[i + 1])
    }
  })
  )

  while (extra_teams > 0){
      teams[which.max(teams)] <- NA
      extra_teams <- extra_teams - 1
  }

  return(teams)
}

nextLayer <- function(teams) {
  out <- c()  # Initialize an empty vector
  length <- length(teams) * 2 + 1  # Calculate the length for pairing

  # Generate the next layer by pairing teams
  for (i in teams) {
    out <- c(out, i, length - i)  # Push the current team and its pair
  }

  return(out)
}
generate_elimination_tournament_structure(8)
generate_elimination_tournament_structure(16)

```

# Simulating a Tournament

```{r}
# Simulate the single-elimination tournament
simulate_single_elimination_tournament <- function(num_teams, distribution, seeding_structure = NULL, ties = T, series = 1, third_place = T, reseeded = F) {

  if (num_teams <= 3) {
    stop("Number of Teams must be greater than 3.")
  }
  teams <- paste(num_teams:1)
  if (log2(num_teams) %% 1 != 0) {
    power_of_2 <- 2^ceiling(log2(num_teams))
    teams <- c(teams, rep(NA, power_of_2 - num_teams))
  }

  if (distribution == "Normal"){
    strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA,length(teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  }
  else if (distribution == "Same"){
    strengths <- sapply(num_teams, function(n) {rep(0,n)})
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA,length(teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  }
  else if (distribution == "Uniform"){
    strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1) , 0 , sqrt(12)) })
    unif_strength <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams)),
      rank_hat = rep(NA,length(teams))
    )
    df <- arrange(unif_strength, true_rank)
  }
  else if (distribution == "Manual") {
    cat("Please enter the strengths for each team in descending rank order:\n")
    strengths <- numeric(num_teams)
    for (i in num_teams:1) {
      repeat {
        input <- as.numeric(readline(paste("Strength for Team", num_teams - i + 1, ": ")))
        if (!is.na(input)) {
          strengths[i] <- input
          break
        } else {
          cat("Invalid input. Please enter a numeric value.\n")
        }
      }
    }
    manual_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- arrange(manual_strengths, true_rank)
  }
  else {
    stop("Distribution not found: Enter Manual to input your own strengths")
  }

  # Sets seed to that of common best vs worst structure
  if (is.null(seeding_structure)){
    seeding_structure <- matrix(generate_elimination_tournament_structure(num_teams), nrow = 1)
  } else{
    seeding_structure <- matrix(seeding_structure, nrow = 1)
  }
  if(series %% 2 == 0) {
    stop("Series must be an odd number to avoid potential ties.")
  }

  results <- list()  # Initialize the results list

  for (i in 1:nrow(seeding_structure)) {
    # Create a new data frame based on the current permutation in 'test'
    permutation <- seeding_structure[i, ]
    permuted_df <- df
    permuted_df$true_rank <- df$true_rank[permutation]  # Update team order 
    teams <- permuted_df$true_rank
    permuted_df$true_strength <- df$true_strength[permutation]
    permuted_df$rank_hat <- NA  # Initialize ranks column
    permuted_df$game_wins <- 0  # Initialize wins column
    permuted_df$game_losses <- 0

    strengths <- permuted_df$true_strength
    round_number <- ceiling(log2(length(teams)))

    # Track the losers of the semifinals for the third-place match
    losers_semis <- c()

    while (length(teams) > 1) {
      #cat("\n--- New Round ---\n")

      # RESEEDING LOGIC: If reseeded is TRUE, sort teams by their true rank before each round
      if (reseeded && length(teams) > 1) {
        # Remove NA teams for sorting
        valid_teams <- teams[!is.na(teams)]
        na_teams <- teams[is.na(teams)]

        # Sort valid teams by their true rank (best to worst)
        sorted_teams <- valid_teams[order(as.numeric(valid_teams))]

        # Create reseeded matchups: 1 vs worst, 2nd vs 2nd worst, etc.
        n_valid <- length(sorted_teams)
        if (n_valid > 1) {
          reseeded_teams <- c()
          for (j in 1:(n_valid/2)) {
            reseeded_teams <- c(reseeded_teams, sorted_teams[j], sorted_teams[n_valid - j + 1])
          }
          # Add back any NA teams (byes)
          teams <- c(reseeded_teams, na_teams)
        }
      }

      next_round <- c() # Makes sure only winners are included in next round

      for (j in seq(1, length(teams), by = 2)) { # Gets 2 teams for each match
        team1 <- teams[j]
        team2 <- teams[j+1]

        strength1 <- strengths[which(permuted_df$true_rank == team1)]
        strength2 <- strengths[which(permuted_df$true_rank == team2)]

        wins_team1 <- 0
        wins_team2 <- 0
        losses_team1 <- 0
        losses_team2 <- 0
        for (game in 1:series){
          match_winner <- simulate_match(team1, team2, strength1, strength2)
          #cat(sprintf("Matchup:  %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", team1, team2, strength1, strength2, match_winner))
          if(!is.na(team1)) {
            if (match_winner == team1) {
              wins_team1 <- wins_team1 + 1
              losses_team2 <- losses_team2 + 1
            } else {
              wins_team2 <- wins_team2 + 1
              losses_team1 <- losses_team1 + 1
            }
          }
          else {wins_team2 <- series / 2 + 1}

          # Determine the overall winner if a majority is reached
          if (wins_team1 > series / 2 || wins_team2 > series / 2) {
            break
          }
        }

        # adds wins and losses of the series to the teams unless they have a bye
        if (!any(is.na(c(team1, team2)))) {
          permuted_df <- permuted_df %>%
            mutate(game_wins = ifelse(true_rank == team1, game_wins + wins_team1, game_wins)) %>%
            mutate(game_wins = ifelse(true_rank == team2, game_wins + wins_team2, game_wins)) %>%
            mutate(game_losses = ifelse(true_rank == team1, game_losses + losses_team1, game_losses)) %>%
            mutate(game_losses = ifelse(true_rank == team2, game_losses + losses_team2, game_losses)) 
        }

        # Determine the match winner based on series results
        series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
        #cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, series_winner))
        next_round <- c(next_round, series_winner)

        if (round_number > 2){
          # Assign ranks based on the current round
          n <- 2^(round_number)
          #print(series_winner)
          if (!is.na(team1) & series_winner == team1) {
            permuted_df$rank_hat[permuted_df$true_rank == team2] <- n
          } else {
            permuted_df$rank_hat[permuted_df$true_rank == team1] <- n
          }
        } else if (round_number == 2) {  # Semifinals
          n <- 2^(round_number)
          if (match_winner == team1) {
            permuted_df$rank_hat[permuted_df$true_rank == team2] <- n
            losers_semis <- c(losers_semis, team2)
          } else {
            permuted_df$rank_hat[permuted_df$true_rank == team1] <- n
            losers_semis <- c(losers_semis, team1)
          }
        } else if (round_number == 1) { # Finals
          permuted_df <- permuted_df %>%
            filter(!is.na(true_rank))
          n <- 2^(round_number)
          if (series_winner == team1) {
            permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
            permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
          } else {
            permuted_df$rank_hat[permuted_df$true_rank == team1] <- 2
            permuted_df$rank_hat[permuted_df$true_rank == team2] <- 1
          }
        }
      }

      teams <- next_round  # Winners move to the next round
      round_number <- round_number - 1
    }

    # Third place match between the losers of the semifinals
    if (third_place == TRUE){
      team1 <- losers_semis[1]
      team2 <- losers_semis[2]
      strength1 <- permuted_df$true_strength[permuted_df$true_rank == team1]
      strength2 <- permuted_df$true_strength[permuted_df$true_rank == team2]
      wins_team1 <- 0
      wins_team2 <- 0
      losses_team1 <- 0
      losses_team2 <- 0
      for (game in 1:series) {
        match_winner <- simulate_match(team1, team2, strength1, strength2)
        if (match_winner == team1) {
          wins_team1 <- wins_team1 + 1
          losses_team2 <- losses_team2 + 1

        } else {
          wins_team2 <- wins_team2 + 1
          losses_team1 <- losses_team1 + 1
        }
        # Determine the overall winner if a majority is reached
        if (wins_team1 > series / 2 || wins_team2 > series / 2) {
          break
        }
      }

      # Update game_wins / game_losses in permuted_df for both teams
        permuted_df <- permuted_df %>%
          mutate(game_wins = ifelse(true_rank == team1, game_wins + wins_team1, game_wins)) %>%
          mutate(game_wins = ifelse(true_rank == team2, game_wins + wins_team2, game_wins)) %>%
          mutate(game_losses = ifelse(true_rank == team1, game_losses + losses_team1, game_losses)) %>%
          mutate(game_losses = ifelse(true_rank == team2, game_losses + losses_team2, game_losses))

      # Assign third place and fourth place ranks
      if (wins_team1 > wins_team2) {
        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team1))
        permuted_df$rank_hat[permuted_df$true_rank == team1] <- 3
        permuted_df$rank_hat[permuted_df$true_rank == team2] <- 4
      } else {
        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team2))
        permuted_df$rank_hat[permuted_df$true_rank == team2] <- 3
        permuted_df$rank_hat[permuted_df$true_rank == team1] <- 4
      }
    }

    permuted_df$seed <- rep(paste(seeding_structure[i, ], collapse = ", "), nrow(permuted_df))
    results[[i]] <- permuted_df # Store the results of this permutation
  }

  final_results <- do.call(rbind, results)

  # Remove ties if ties = FALSE
  if (!ties) {
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(true_rank) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "random"))
  } else{
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(rank_hat) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "average"))
  }

  final_results <- final_results %>%
    mutate(distribution = distribution) %>%
    mutate(true_rank = as.numeric(true_rank)) %>%
    arrange(true_rank)
  return(final_results)
}

  test <- list()
for (teams in c(8)) {
  test[[as.character(teams)]] <- list()
  for (replicate in 1:10000) {
    # Store each simulation result in the list with unique indexing
    test[[as.character(teams)]][[replicate]] <- simulate_single_elimination_tournament(
      num_teams = teams,
      distribution = "Same",
      ties = FALSE,
      series = 1,
      third_place = TRUE
    )
  }
}

single_run_example <- simulate_single_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1, third_place=T, reseeded=F)

#manual_test <- simulate_single_elimination_tournament(num_teams = 4, distribution = "Manual", ties=T, series=1, third_place=T)
#manual_test

# Gets all results of a normal distribution num_teams between 4 and 64 single-game elimination
all_normal_results <- list()
for (teams in c(8)) {
  all_normal_results[[as.character(teams)]] <- list()
  for (replicate in 1:10000) {
    # Store each simulation result in the list with unique indexing
    all_normal_results[[as.character(teams)]][[replicate]] <- simulate_single_elimination_tournament(
      num_teams = teams,
      distribution = "Normal",
      ties = FALSE,
      series = 1,
      third_place = TRUE
    )
  }
}

all_results_norm <- do.call(rbind, lapply(names(all_normal_results), function(teams) {
  team_results <- do.call(rbind, all_normal_results[[teams]])
  team_results$teams <- as.numeric(teams)
  return(team_results)
}))

all_results_norm

# Gets all results of a uniform distribution num_teams between 4 and 64 single-game elimination
all_unif_results <- list()
for (teams in c(4,8,16,32,64)) {
  all_unif_results[[as.character(teams)]] <- list()
  for (replicate in 1:100) {
    # Store each simulation result in the list with unique indexing
    all_unif_results[[as.character(teams)]][[replicate]] <- simulate_single_elimination_tournament(
      num_teams = teams,
      distribution = "Uniform",
      ties = FALSE,
      series = 1,
      third_place = TRUE
    )
  }
}

all_results_unif <- do.call(rbind, lapply(names(all_unif_results), function(teams) {
  team_results <- do.call(rbind, all_unif_results[[teams]])
  team_results$teams <- as.numeric(teams)
  return(team_results)
}))

all_results_unif

seven_games_se_results <- list()
for (teams in c(8)) {
  seven_games_se_results[[as.character(teams)]] <- list()
  for (replicate in 1:10000) {
    # Store each simulation result in the list with unique indexing
    seven_games_se_results[[as.character(teams)]][[replicate]] <- simulate_single_elimination_tournament(
      num_teams = teams,
      distribution = "Normal",
      ties = FALSE,
      series = 7,
      third_place = TRUE
    )
  }
}

seven_games_se_results <- do.call(rbind, lapply(names(seven_games_se_results), function(teams) {
  team_results <- do.call(rbind, seven_games_se_results[[teams]])
  team_results$teams <- as.numeric(teams)
  return(team_results)
}))

reseeded_results <- list()
for (teams in c(8)) {
  reseeded_results[[as.character(teams)]] <- list()
  for (replicate in 1:10000) {
    # Store each simulation result in the list with unique indexing
    reseeded_results[[as.character(teams)]][[replicate]] <- simulate_single_elimination_tournament(
      num_teams = teams,
      distribution = "Normal",
      ties = FALSE,
      series = 1,
      third_place = TRUE,
      reseeded = TRUE
    )
  }
}

reseeded_results <- do.call(rbind, lapply(names(reseeded_results), function(teams) {
  team_results <- do.call(rbind, reseeded_results[[teams]])
  team_results$teams <- as.numeric(teams)
  return(team_results)
}))

reseeded_results

```

## Shiny User Interface Tournament Simulation

```{r}
# Define the Shiny app UI
ui <- fluidPage(
  titlePanel("Single-Elimination Tournament Simulator"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 4, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      uiOutput("manual_strengths_ui"), # Dynamic UI for Manual strengths
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      h4("Tournament Results"),
      tableOutput("ranked_results")
    )
  )
)

# Define the server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_elimination_tournament_structure(input$num_teams))
    }
  })

  tournament_results <- eventReactive(input$run_sim, {
    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs

    seed_structure <- custom_seeding()

    simulate_single_elimination_tournament(
      num_teams = input$num_teams,
      distribution = input$distribution,
      ties = input$ties,
      series = input$series,
      third_place = input$third_place,
      seeding_structure = seed_structure
    )
  })

  output$ranked_results <- renderTable({
    results <- tournament_results()
    results[order(results$true_rank), ]  # Order by Strength descending
  })
}

shinyApp(ui = ui, server = server)
```

## Shiny Number of Wins

```{r}
# Shiny app UI
ui <- fluidPage(
  titlePanel("Tournament Simulation with Total Wins Visualization"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      h4("Total Wins Summary"),
      plotOutput("wins_plot")
    )
  )
)

# Shiny server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_elimination_tournament_structure(input$num_teams))
    }
  })

  simulation_results <- eventReactive(input$run_sim, {
    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
    seed_structure <- custom_seeding()
    # Run simulations
    replicate_results <- replicate(
      input$num_simulations,
      simulate_single_elimination_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        third_place = input$third_place,
        seeding_structure = seed_structure
      ),
      simplify = FALSE
    )

    # Combine all results and summarize total wins
    all_results <- do.call(rbind, replicate_results)
    total_wins_summary <- all_results %>%
      group_by(true_rank) %>%
      summarise(
        total_wins = sum(rank_hat==1, na.rm = TRUE),
        .groups = 'drop'
      ) %>%
      arrange(desc(total_wins))
    return(total_wins_summary)
  })

  output$wins_plot <- renderPlot({
    req(simulation_results())
    results <- simulation_results()

    ggplot(results, aes(x = reorder(true_rank, -total_wins), y = total_wins, fill = total_wins)) +
      geom_text(aes(label = total_wins), vjust = -0.5, size = 4, color = "black") +
      geom_bar(stat = "identity") +
      labs(
        title = "Total Wins Across Simulations",
        x = "True Rank",
        y = "Total Wins"
      ) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_fill_gradient(low = "skyblue", high = "blue")
  })
}

# Run the Shiny app
shinyApp(ui = ui, server = server)
```

# Multiple Simulations for 1 Seeding Structure

```{r}
ui <- fluidPage(
  titlePanel("Tournament Simulation"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", 8, min = 4, step = 1),
      selectInput("distribution", "Distribution Type:", choices = c("Normal", "Uniform")),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      numericInput("replicates", "Number of Replicates:", 100, min = 1, max = 1000, step = 1),
      numericInput("series", "Number of Games in a Series (odd number):", 1, min = 1, max = 11, step = 2),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Tournament Simulation")
    ),

    mainPanel(
      tableOutput("summary_results"),
      verbatimTextOutput("kendall_correlation"),
      verbatimTextOutput("spearman_correlation")
    )
  )
)

server <- function(input, output, session) {
 custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_elimination_tournament_structure(input$num_teams))
    }
  })

  # Function to run tournament simulations
  observeEvent(input$run_sim, {
    req(input$num_teams, input$distribution, input$replicates, input$series)

    # Initialize lists to store correlation results
    kendall_correlations <- numeric(input$replicates)
    spearman_correlations <- numeric(input$replicates)

    seed_structure <- custom_seeding()

    # Run tournament simulations for the specified number of replicates
    simulation_results <- list()
    for (replicate in 1:input$replicates) {
      simulation_results[[replicate]] <- simulate_single_elimination_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        third_place = input$third_place,
        seeding_structure = seed_structure
      )

      # Combine all results into one data frame
      all_results <- do.call(rbind, simulation_results)

      # Compute average ranks across all replicates
      average_ranks <- all_results %>%
        group_by(true_rank) %>%
        summarise(avg_rank = mean(rank_hat, na.rm = TRUE), .groups = 'drop') %>%
        arrange(true_rank)

      # Calculate Kendall Tau correlation for the current replicate
      kendall_correlations[replicate] <- cor(all_results$rank_hat, all_results$true_rank, method = "kendall", use = "complete.obs")

      # Calculate Spearman correlation for the current replicate
      spearman_correlations[replicate] <- cor(all_results$rank_hat, all_results$true_rank, method = "spearman", use = "complete.obs")
    }

    # Compute average Kendall Tau and Spearman correlations across all replicates
    avg_kendall_cor <- mean(kendall_correlations, na.rm = TRUE)
    avg_spearman_cor <- mean(spearman_correlations, na.rm = TRUE)

    # Show summary results in table
    output$summary_results <- renderTable({
      average_ranks
    })

    # Display average Kendall Tau correlation
    output$kendall_correlation <- renderPrint({
      cat("Average Kendall Tau Correlation:", avg_kendall_cor, "\n")
    })

    # Display average Spearman correlation
    output$spearman_correlation <- renderPrint({
      cat("Average Spearman Correlation:", avg_spearman_cor, "\n")
    })
  })
}

shinyApp(ui, server)

```

## Shiny Probability of Top 3 Correct

```{r}
ui <- fluidPage(
  titlePanel("Probability of Top 3 Teams in the Correct Order"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      h4("Top 3 Correct"),
      verbatimTextOutput("top3_output")
    )
  )
)

# Shiny server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_elimination_tournament_structure(input$num_teams))
    }
  })

  simulation_results <- eventReactive(input$run_sim, {
    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
    seed_structure <- custom_seeding()

    # Run simulations
    replicate_results <- replicate(
      input$num_simulations,
      simulate_single_elimination_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        third_place = input$third_place,
        seeding_structure = seed_structure
      ),
      simplify = FALSE
    )

    # Combine all results into one dataframe and add a Simulation column
    all_results <- do.call(rbind, replicate_results)
    all_results$simulation <- rep(1:input$num_simulations, each = nrow(all_results) / input$num_simulations)

    return(all_results)
  })

  # Identify top-3 ranks for each simulation and calculate the probability
  top_three_probability <- reactive({
    all_results <- simulation_results()

    # Group by simulation and filter for top 3 true_rank and rank_hat
    top_three_results <- all_results %>%
      group_by(simulation) %>%
      filter(true_rank %in% 1:3 & rank_hat %in% 1:3) %>%
      arrange(simulation, rank_hat) %>%
      summarize(top_3_true_rank = list(true_rank), top_3_rank = list(rank_hat), .groups = 'drop') %>%
      mutate(match = mapply(function(true, rank) setequal(true, rank), top_3_true_rank, top_3_rank))

    # Calculate probability of matching top 3 true_rank with rank_hat
    match_count <- sum(top_three_results$match)
    total_simulations <- n_distinct(all_results$simulation)

    probability <- match_count / total_simulations
    return(probability)
  })

  # Output the probability to the UI
  output$top3_output <- renderText({
    probability <- top_three_probability()

    paste("Probability of Top 3 Seeds with Ranks 1-3: ", round(probability,4))
  })
}

shinyApp(ui=ui,server=server)
```

## Shiny Conditional Probability Curve

```{r}
ui <- fluidPage(
  titlePanel("Conditional Probability of Team Ranks in Tournament"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", 
                placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

    mainPanel(
      plotOutput("probability_plot")
    )
  )
)

# Server
server <- function(input, output) {

  # Helper Function: Parse Custom Seeding Structure
  parse_seeding <- function(seeding_input, num_teams) {
    if (seeding_input != "") {
      seeding_vector <- unlist(strsplit(seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)

      num_teams_next_power_of_two <- 2^ceiling(log2(num_teams))
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(
          paste("Seeding structure length must match", num_teams_next_power_of_two), 
          type = "error"
        )
        return(NULL)
      }
    } else {
      return(NULL)
    }
  }

  # Curve Function to Calculate Conditional Probabilities
  curve <- function(sim, num_teams) {
    probability <- numeric(num_teams)
    for (i in 1:num_teams) {
      if (i == 1) {
        probability[i] <- nrow(sim[sim$rank_hat == 1 & sim$true_rank == 1, ]) / length(unique(sim$simulation))
      } else {
        top_ranks <- sim[sim$rank_hat <= (i - 1) & sim$true_rank <= (i - 1), ]
        better_teams <- 1:(i - 1)
        numerator <- sum(sapply(unique(top_ranks$simulation), function(sim_id) {
          sim_subset <- sim[sim$simulation == sim_id, ]
          if (all(better_teams %in% sim_subset$true_rank[sim_subset$rank_hat <= (i - 1)])) {
            return(sum(sim_subset$rank_hat == i & sim_subset$true_rank == i))
          } else {
            return(0)
          }
        }))
        denominator <- sum(sapply(unique(top_ranks$simulation), function(sim_id) {
          sim_subset <- top_ranks[top_ranks$simulation == sim_id, ]
          all(better_teams %in% sim_subset$true_rank)
        }))
        probability[i] <- numerator / denominator
      }
    }
    return(probability)
  }

  # Reactive Event: Run Simulations
  simulation_data <- eventReactive(input$run_sim, {
    num_teams <- input$num_teams
    distribution <- input$distribution
    series <- input$series
    num_simulations <- input$num_simulations
    third_place <- input$third_place

    # Parse custom seeding structure
    seeding_structure <- parse_seeding(input$seeding_input, num_teams)

    # Run simulations
    replicate_results <- replicate(
      num_simulations,
      simulate_single_elimination_tournament(
        num_teams = num_teams,
        distribution = distribution,
        series = series,
        ties = FALSE,
        third_place = third_place,
        seeding_structure = seeding_structure
      ),
      simplify = FALSE
    )

    # Combine all results into one dataframe
    all_results <- do.call(rbind, replicate_results)
    all_results$simulation <- rep(1:num_simulations, each = nrow(all_results) / num_simulations)
    # list_rbind(x=all_results, names_to = "simulation")

    # Calculate conditional probabilities
    sim_probs <- curve(all_results, num_teams)
    probabilities <- data.frame(x = 1:num_teams, y = sim_probs)
    probabilities$line <- "Simulated"

    # Baseline probabilities
    baseline_prob <- 1 / (num_teams:1)
    baseline <- data.frame(x = 1:num_teams, y = baseline_prob)
    baseline$line <- "Baseline"

    # Combine for plotting
    rbind(baseline, probabilities)
  })

  # Plot Output
  output$probability_plot <- renderPlot({
    req(simulation_data())
    ggplot(data = simulation_data(), aes(x = x, y = y, color = line, group = line)) +
      geom_point() +
      geom_line() +
      ylim(0, 1) +
      labs(
        x = "Rank",
        y = "Conditional Probability",
        title = "Simulated vs Baseline Conditional Probabilities"
      ) +
      theme_minimal()
  })
}

# Run the App
shinyApp(ui = ui, server = server)
```