### 2025-11-07 nfl_playoffs.qmd

---
title: "NFL Playoffs"
format: html
editor: visual
---

## Standard NFL Playoff Format

```{r}
# Initialize storage for results
nfl_results <- vector("list", 10000)

for (sim in 1:10000){
  nfc <- simulate_single_elimination_tournament(num_teams = 6, distribution = "Manual", 
                                                 ties=T, series=1, 
                                                 seeding_structure = c(1,NA,5,4,6,3,2,NA), 
                                                 third_place=F, reseeded=T, theta_hat = nfc_theta)

  afc <- simulate_single_elimination_tournament(num_teams = 6, distribution = "Manual", 
                                                 ties=T, series=1, 
                                                 seeding_structure = c(1,NA,2,6,4,5,3,NA), 
                                                 third_place=F, reseeded=T, theta_hat = afc_theta)

  nfc_theta_df <- nfc %>%
    left_join(theta, by = c("true_strength" = "theta")) %>%
    select(c(true_strength, true_rank, rank_hat, team))

  afc_theta_df <- afc %>%
    left_join(theta, by = c("true_strength" = "theta")) %>%
    select(c(true_strength, true_rank, rank_hat, team))

  winner_nfc <- nfc_theta_df %>% filter(rank_hat == 1)
  winner_afc <- afc_theta_df %>% filter(rank_hat == 1)
  sb_winner <- simulate_match(winner_nfc$team, winner_afc$team, 
                              winner_nfc$true_strength, winner_afc$true_strength)

  # Combine and adjust rankings
  nfl_playoff_results <- rbind(nfc_theta_df, afc_theta_df) %>%
    mutate(
      # Super Bowl winner gets rank 1, both conference winners get rank 2, others shift down
      rank_hat = case_when(
        team == sb_winner ~ 1,
        rank_hat == 1 ~ 2,  # losing conference champion
        TRUE ~ rank_hat + 1
      )
    ) %>%
    arrange(-true_strength) %>%
    mutate(
      true_rank = 1:12,
      sim_id = sim
    )

  all_results[[sim]] <- nfl_playoff_results
}

# Combine all simulations
nfl_results <- bind_rows(all_results)
```