### 2025-11-13 mutual_information_curves.qmd
Recommit changes without large RData file

---
title: "Mutual Information Plots"
format: html
editor: visual
---

## Mutual Information Plots

From Wikipedia, weighted mutual information formula:

$I(X;Y)=\sum _{y\in Y}\sum _{x\in X}w(x,y)p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}$

Rajski's distance: $1 - \frac{I(X,Y)}{H(X,Y)}$

Interpretation: the fraction of uncertainty not explained by mutual information

<https://en.wikipedia.org/wiki/Inverse_distance_weighting>

Every permutation is not possible to elimination tournaments, so should we divide by every possible permutation for 8 teams, or just every possible permutation in the tournament structure?

## Weight Structure (Partial Kendall or Weighted Kendall)

##

```{r}
r <- c(1,2,3,4,5)
rhat <- c(4,1,1,1,5)
n <- length(r)
k <- sum(rhat == 1)
count <- 0
for (i in 1:(length(r)-1)){
  for (j in ((i+1):(length(r)))){
    print(sign(r[i] - r[j])*sign(rhat[i] - rhat[j]))
    count <- count + sign(r[i] - r[j])*sign(rhat[i] - rhat[j])
  }
}

count
#Correlation
count / (choose(n,2) - choose(k,2))
#Weight
((count / (choose(n,2) - choose(k,2))) + 1)/2

```

```{r}
library(gtools)
library(tidyverse)
library(combinat)

squared_error <- function(vec, weights = NULL) {
  n <- length(vec)
  ref <- 1:n

  # Default weights = 1 for each element
  if (is.null(weights)) {
    weights <- rep(1, n)
  }

  # Initialize sum
  summ <- 0

  # Accumulate weighted squared error
  for (i in 1:n) {
    summ <- summ + (weights[i] * (vec[i] - ref[i]))^2
  }

  # Handle near-zero sum safely
  if (abs(summ) < 1e-10) {
    out <- 1
  } else {
    out <- 1 / summ
  }

  return(out)
}

partial_kendall <- function(vec, k, weights = NULL){
  n <- length(vec) 
  if (k == n){k <- n-1} 
  ref <- c(1:n)

  if(is.null(weights)){
    weights <- rep(1,n)
  }

  count <- 0 
  #denom <- sum(n - (1:k))

  for (i in 1:k){ 
    for (j in (i+1):n){
      #pair_weight <- weights[i] + weights[j]

        add <- (((vec[i] < vec[j]) & (ref[i] < ref[j])) | 
                  ((vec[i] > vec[j]) & (ref[i] > ref[j]))) + 0
        #print(c(i, j, add, pair_weight)) 
        #if (add == 1){
          count <- count + add #+ pair_weight
        #} 
        denom <- denom + pair_weight
    } 
  } 
  weight <- count / denom

  return(weight) 
  }

weighted_kendall <- function(rhat, k, wtype = "symmetric_salience", salient_weights = NULL) {
  n <- length(rhat)
  #r <- 1:n
  if (is.null(salient_weights)) salient_weights <- rep(1, k)

  # Weight matrix
  w <- matrix(0, ncol = n, nrow = n)
  if (is.null(wtype) || wtype == "symmetric_salience") {
    for (q in k:1) {
      w[q, ] <- salient_weights[q]
      w[, q] <- salient_weights[q]
    }
  }
  if (wtype == "rydog") {
    for (q in k:1) {
      w[1:q, 1:q] <- salient_weights[q]
    }
  }

  # Compute weighted Kendall numerator and denominator consistently
  summ <- 0
  denom <- 0
  for (i in 1:(n-1)) {
    for (j in (i+1):n) {
      if (i != j) {
        sign_ij <- sign(i - j)
        sign_rhat <- sign(rhat[i] - rhat[j])
        sign_rhat <- ifelse(sign_rhat == 0, -1, sign_rhat)
        summ <- summ + w[i, j] * sign_ij * sign_rhat
        denom <- denom + w[i, j]
        #print(c(sign_r, sign_rhat, summ, denom))
      }
    }
  }

  out <- summ / denom
  weight <- (out + 1) / 2
  return(weight)
}

weighted_kendall(rhat=c(2,1,3,4), k=4)
weighted_kendall(c(1.5,3,1.5), k=2)
weighted_kendall(c(1,3,4,2), k=3)
weighted_kendall(c(1,1,4,1), k=3)

# Define the function calls
results <- data.frame(
  rhat = rep(c("(1,1,3)", "(1,3,1)", "(3,1,1)"), each = 3),
  k = rep(1:3, times = 3),
  weighted_kendall = c(
    weighted_kendall(c(1,1,3), k=1),
    weighted_kendall(c(1,1,3), k=2),
    weighted_kendall(c(1,1,3), k=3),
    weighted_kendall(c(1,3,1), k=1),
    weighted_kendall(c(1,3,1), k=2),
    weighted_kendall(c(1,3,1), k=3),
    weighted_kendall(c(3,1,1), k=1),
    weighted_kendall(c(3,1,1), k=2),
    weighted_kendall(c(3,1,1), k=3)
  )
)

results

weighted_kendall(c(1,1,1,1,5,6), k = 3, r = c(1,1,1,1,5,6))
# Only works for k = # of top teams tied

make_weights <- function(k) {
  w <- rep(0, 4)
  w[1:k] <- 1
  return(w)
}

# Store all results
all_results <- list()

perms_list <- permn(1:4)

# Loop over k = 1 to 4
for (k in 1:4) {
  weights <- make_weights(k)

  # Weighted Kendall (using provided function)
  kendall_results <- sapply(perms_list, function(p) {
    weighted_kendall(rhat = p, k = k, wtype = "symmetric_salience")
  })

  # Squared error using current weights
  sq_results <- sapply(perms_list, function(p) {
    squared_error(vec = p, weights = weights)
  })

  # Combine into one dataframe for this k
  rhat_df <- data.frame(
    k = k,
    permutation = sapply(perms_list, paste, collapse = ""),
    weighted_kendall = kendall_results,
    squared_error = sq_results
  )

  all_results[[k]] <- rhat_df
}

# Combine across all k
rhat_df <- bind_rows(all_results)
```

```{r}
library(combinat)
library(tidyverse)
library(viridis)

mut_info <- function(sims, num_teams, rank_weights = NULL, wtype = "symmetric_salience"){
  # Handle missing weights (default to equal weights if not provided)
  if (is.null(rank_weights)) {
    rank_weights <- rep(1, num_teams)
  }

  # Extract rank_hat lists, widen into columns y1..yN
  y_df <- sims %>%
    group_by(simulation) %>%
    summarise(rank_hat = list(rank_hat), .groups = "drop") %>%
    mutate(rank_hat = map(rank_hat, unlist)) %>%
    unnest_wider(rank_hat, names_sep = "_")

  colnames(y_df)[-1] <- paste0("y", 1:num_teams)

  # Probability of each observed permutation
  prob_df <- y_df %>%
    select(starts_with("y")) %>%
    group_by(across(everything())) %>%
    summarise(count = n(), .groups = "drop") %>%
    mutate(probability = count / sum(count))

  w <- numeric(nrow(prob_df))  # Initialize weight vector for the number of rows in prob_df

  #for (i in 1:num_teams) {
    #x_i <- paste0("x", i)
    #y_i <- paste0("y", i)

    #max_footrule <- sum(rank_weights * abs((1:num_teams) - (num_teams:1)))
    # Loop through each row of prob_df
    for (j in 1:nrow(prob_df)) {
      #x_vals <- as.numeric(prob_df[j, paste0("x", 1:num_teams)])
      y_vals <- as.numeric(prob_df[j, paste0("y", 1:num_teams)])

      # Calculate the squared difference between x_i and y_i for each row j
      #diff <- abs(prob_df[[x_i]][j] - prob_df[[y_i]][j])
      #rank_diff <- abs(rank(x_vals) - rank(y_vals))
      #w[j] <- w[j] + rank_weights[i] * diff^2  # Sum squared differences for each row

      # Weighted Kendall's Tau correlation
      #x_rep <- rep(x_vals, rank_weights)
      #y_rep <- rep(y_vals, rank_weights)
      #tau <- cor(x_rep, y_rep, method="kendall")
      #if(!is.na(tau)){
        #w[j] <- (1 - tau) / 2 * choose(length(x_rep),2)
      #} else{
      #  diff <- abs(prob_df[[x_i]][j] - prob_df[[y_i]][j])
      #  w[j] <- w[j] <- w[j] + rank_weights[i] * diff^2
      #}

      # Truncated Normalized Kendall's Tau Distance
      # Function to compute normalized "swap similarity" for top-K
      w[j] <- weighted_kendall(rhat = y_vals, k=num_teams, salient_weights = rank_weights, wtype = wtype)

      # Initialize weighted concordance / discordance sum
      #weighted_inversions <- 0

      #for (l in 1:i) {
      #  for (k in (l+1):num_teams) {
      #    concordant <- (y_vals[l] - y_vals[k]) > 0
      #    discordant <- (y_vals[l] - y_vals[k]) < 0

          # Weight each pair by sum of their rank_weights
      #    pair_weight <- rank_weights[l] + rank_weights[k]

      #    if (discordant) weighted_inversions <- weighted_inversions + pair_weight
      #  }
      #}
    #w[j] <- weighted_inversions

        # Calculate NDCG (Normalized discounted cummulative gain)
      # x_vals tells us the true ranking positions
      # y_vals tells us the predicted ranking positions
      #relevance <- rank_weights * (num_teams - x_vals + 1)
      # Calculate DCG based on predicted ranking (y_vals)
      #dcg <- 0
      # for (pos in 1:num_teams) {
      #   # Find which team is at position 'pos' in the predicted ranking
      #   team_idx <- which(y_vals == pos)
      #   if (length(team_idx) > 0) {
      #     # Use that team's relevance weight
      #     dcg <- dcg + relevance[team_idx] / log2(pos + 1)
      #   }
      # }
      # # Calculate ideal DCG (teams sorted by their rank_weights in descending order)
      # ideal_order <- order(relevance, decreasing = TRUE)
      # idcg <- 0
      # for (pos in 1:num_teams) {
      #   idcg <- idcg + relevance[ideal_order[pos]] / log2(pos + 1)
      # }
      # # NDCG = DCG / IDCG
      # if (idcg > 0) {
      #   ndcg <- dcg / idcg
      #   # Convert NDCG (0 to 1, where 1 is perfect) to a distance measure
      #   # Use (1 - NDCG) so that perfect match = 0 distance
      #   w[j] <- 1 - ndcg
      # } else {
      #   w[j] <- 1  # Maximum distance if IDCG is 0
      # }

      # Calculate weighted Spearman's footrule distance
      # Footrule = sum of |true_rank - predicted_rank| for each item
      # Weighted by rank_weights for each team

      #footrule_distance <- 0
      #for (i in 1:num_teams) {
        # Team i has true rank x_vals[i] and predicted rank y_vals[i]
        #footrule_distance <- footrule_distance + rank_weights[i] * abs(x_vals[i] - y_vals[i])
      #}

      # Normalize by maximum possible distance
      #if (max_footrule > 0) {
      #  w[j] <- footrule_distance / max_footrule
      #} else {
      #  w[j] <- 0  # All weights are 0, so distance is 0
      #}

    }
  #}
  #w <- 1 / (w+1)  # Inverse of squared differences and manually change 0 probabilities to 1 on the inverse
  #w <- w/sum(w)
  # Compute weights dynamically for each row
  weighted_df <- prob_df %>%
    mutate(w = w)

  numerator <- sum(weighted_df$w * weighted_df$probability * log2(weighted_df$probability / (1/factorial(num_teams) * (1/factorial(num_teams)))))
  denominator <- log2(factorial(num_teams)^2)

  Raj_dist <- numerator/denominator
  # Return the weighted data frame with mutual information weights
  return(Raj_dist)
}

# Single Elimination with reseeding
reseeded_results_eight <- reseeded_results %>% filter(teams == 8)
reseeded_results_eight$simulation <- rep(1:10000, each = nrow(reseeded_results_eight) / 10000)

# Group Stage (like World Cup)
group_stage_results_eight <- group_stage_results %>% filter(teams == 8)
group_stage_results_eight$simulation <- rep(1:10000, each = nrow(group_stage_results_eight) / 10000)

# Double Elimination
de_norm_results_eight <- de_norm_results %>% filter(teams == 8)
de_norm_results_eight$simulation <- rep(1:10000, each = nrow(de_norm_results_eight) / 10000)

# Single Elimination with Normal Distribution
all_results_norm$simulation <- rep(1:10000, each = nrow(all_results_norm) / 10000)

# ROund Robin with randomly assigning ties
test_rr_noTies <- test_rr %>%
  group_by(simulation) %>%
  mutate(rank_hat = rank(rank_hat,ties.method="random"))

# All possible orderings of teams one time
perms_list <- permn(8)
all_data <- list()
# Loop through each permutation and create the data frame
for (i in 1:length(perms_list)) {
# Create a temporary data frame for the current permutation
temp_df <- data.frame(
true_rank = rep(1:8, each = 1),  # Each row gets a true_rank from 1 to 8
rank_hat = perms_list[[i]],      # The current permutation's ranks
simulation = rep(i, 8)           # Assign the simulation number to each row
)
# Add the temporary data frame to the list
all_data[[i]] <- temp_df
}
# Combine all the data frames into one
perms_df <- do.call(rbind, all_data)

# Best Possible (every team in correct order)
best <- data.frame(
  true_rank = rep(1:8, times = 10000),  # Repeat 1-8 for 10,000 simulations
  rank_hat = rep(1:8, times = 10000),   # Ensure rank_hat matches true_rank
  simulation = rep(1:10000, each = 8)   # Assign each set to a simulation
)
# Worst Possible (worst team always wins; best last)
worst <- data.frame(
  true_rank = rep(1:8, times = 10000),  # Repeat 1-8 for 10,000 simulations
  rank_hat = rep(8:1, times = 10000),   # Ensure rank_hat matches true_rank
  simulation = rep(1:10000, each = 8)   # Assign each set to a simulation
)

# Single Elimination with 7 game series
seven_games_se_results$simulation <- rep(1:10000, each = nrow(seven_games_se_results) / 10000)

# Function to compute mutual information for a given dataset
compute_mut_plots <- function(dataset) {
  num_teams <- length(unique(dataset$true_rank))
  results <- numeric(num_teams)
  for (i in 1:num_teams) {
    binary_vector <- c(rep(1, i), rep(0, num_teams - i))  # Construct vector
    results[i] <- mut_info(dataset, num_teams, binary_vector)  # Compute MI
  }
  return(results)
}

# Compute mutual information for each dataset
mi_se_reseeded <- compute_mut_plots(reseeded_results_eight)
mi_group_stage <- compute_mut_plots(group_stage_results_eight)
mi_de_norm <- compute_mut_plots(de_norm_results_eight)
mi_se_norm <- compute_mut_plots(all_results_norm)
mi_test_rr <- compute_mut_plots(test_rr_noTies)
mi_seven_se <- compute_mut_plots(seven_games_se_results)
mi_random <- compute_mut_plots(equal_strengths)
mi_se_unif <- compute_mut_plots(test_unif)
mi_bad <- compute_mut_plots(bad_seeding_structure)
mi_perms <- compute_mut_plots(perms_df)
mi_best <- compute_mut_plots(best)
mi_worst <- compute_mut_plots(worst)

mi_plots_df <- data.frame(
indices = 1:8,
"Single Elimination (Reseeded)" = mi_se_reseeded,
"Group Stage" = mi_group_stage,
"Double Elimination" = mi_de_norm,
"Single Elimination (Set Structure)" = mi_se_norm,
"Coin Flip Single Elimination" = mi_random,
#"Unif_SE" = mi_se_unif,
"Bad Seeding Structure Single Elimination" = mi_bad,
"Coin Flip Round Robin" = mi_perms,
"Round Robin" = mi_test_rr,
"Worst Possible" = mi_worst,
"Best Possible" = mi_best,
check.names = FALSE
)

# Reshape data into long format for ggplot
df_long <- pivot_longer(mi_plots_df, cols = -indices, names_to = "Method", values_to = "MI")
df_long <- df_long %>% mutate(Method = recode( Method, "Single.Elimination..Reseeded." = "Single Elimination (Reseeded)", "Group.Stage" = "Group Stage", "Double.Elimination" = "Double Elimination", "Single.Elimination..Set.Structure." = "Single Elimination (Set Structure)", "Coin.Flip.Single.Elimination" = "Coin Flip Single Elimination", "Bad.Seeding.Structure.Single.Elimination" = "Bad Seeding Structure Single Elimination", "Coin.Flip.Round.Robin" = "Coin Flip Round Robin", "Round.Robin" = "Round Robin", "Worst.Possible" = "Worst Possible", "Best.Possible" = "Best Possible" ))

okabe_ito <- c(
  "#E69F00", # orange
  "#56B4E9", # sky blue
  "#009E73", # bluish green
  "#F0E442", # yellow
  "#0072B2", # blue
  "#D55E00", # vermillion
  "#CC79A7", # reddish purple
  "#999999", # gray
  "#8A2BE2", # violet (extra)
  "#20B2AA"  # teal (extra)
)
# Compute the ordering of methods by mean MI (descending) 
method_order <- df_long %>% 
  group_by(Method) %>% 
  summarize(mean_MI = mean(MI, na.rm = TRUE)) %>% 
  arrange(desc(mean_MI)) %>% 
  pull(Method)

# Reorder Method factor levels by MI 
df_long <- df_long %>% 
  mutate(Method = factor(Method, levels = method_order))

# Define color mapping in the new order 
methods <- levels(df_long$Method) 
method_colors <- setNames(okabe_ito[seq_along(methods)], methods)

windowsFonts(Arial = windowsFont("Arial"))
ggplot(df_long, aes(x = indices, y = MI, color = Method, group = Method)) +
geom_line(linewidth = 2.5) +           # Add lines
geom_point(size=3) +          # Add points
labs(
title = "Comparison of Tournament Structures with Normal Strengths",
x = "Top i Number of Teams",
y = "Normalized Mutual Information \\ Entropy"
) +
theme_minimal() +       # Use a clean theme
theme(text = element_text(family = "Arial", size = 32),
  legend.position = "bottom",
  legend.box = "vertical",
  plot.title = element_text(hjust = 0.5, size = 36),
  axis.title = element_text(size = 30),
  axis.text = element_text(size=30)) +
  guides(color = guide_legend(nrow = 5)) +
  scale_color_manual(values = method_colors)
ggsave(
  filename = "tournament_comparison_plot.png",  # Change file type if needed
  plot = last_plot(),  # Saves the last ggplot created
  dpi = 600,        
  bg = "white",
  units = "in",
  width = 15.28,
  height = 13.43
)

# Apply normalization to all columns except "indices"
df_close <- df_long %>% filter(Method %in% c("Single Elimination (Reseeded)", "Double Elimination", "Group Stage", "Single Elimination (Set Structure)" ))
windowsFonts(Arial = windowsFont("Arial"))
ggplot(df_close, aes(x = indices, y = MI, color = Method, group = Method)) +
geom_line(linewidth = 2.5) +           # Add lines
geom_point(size=3) +          # Add points
labs(
title = "Comparison of Tournament Structures with Normal Strengths",
x = "Top i Number of Teams",
y = "Normalized Mutual Information \\ Entropy"
) +
theme_minimal() +       # Use a clean theme
theme(text = element_text(family = "Arial", size = 32),
  legend.position = "bottom",
  legend.box = "vertical",
  plot.title = element_text(hjust = 0.5, size = 36),
  axis.title = element_text(size = 30),
  axis.text = element_text(size=30)) +
  guides(color = guide_legend(nrow = 5)) +
  scale_color_manual(values = method_colors)
ggsave(
  filename = "tournament_comparison_plot_close.png",  # Change file type if needed
  plot = last_plot(),  # Saves the last ggplot created
  dpi = 600,        
  bg = "white",
  units = "in",
  width = 15.28,
  height = 13.43
)

```

## MI Calculation and Plotting

These functions can take a list of dataframes and create mutual information that can then be plotted.

```{r}
compute_mi_for_datasets <- function(dataset_list) {
  # Validate input
  if (!is.list(dataset_list)) {
    stop("dataset_list must be a list of data frames")
  }

  # Get dataset names from the list
  if (!is.null(names(dataset_list)) && all(names(dataset_list) != "")) {
    dataset_names <- names(dataset_list)
  } else {
    dataset_names <- paste0("Dataset_", seq_along(dataset_list))
  }

  # Helper function to add simulation number if needed
  add_simulation_number <- function(df) {
    if (!"simulation" %in% colnames(df)) {
      n_rows <- nrow(df)
      n_sims <- 10000
      rows_per_sim <- round(n_rows / n_sims)

      # Memory efficient: create simulation numbers
      df$simulation <- rep(1:n_sims, each = rows_per_sim)
    }
    return(df)
  }

  # Compute MI for all datasets and store as list column
  mi_list <- lapply(seq_along(dataset_list), function(i) {
    df <- dataset_list[[i]]
    df <- add_simulation_number(df)

    # Compute MI - returns a vector or dataframe
    result <- compute_mut_plots(df)

    # Convert to vector if it's a dataframe
    if (is.data.frame(result)) {
      result <- unlist(result)
    }

    return(result)
  })

  # Return dataframe with list column
  data.frame(
    dataset = dataset_names,
    stringsAsFactors = FALSE
  ) %>%
    mutate(mi = mi_list)
}

# Example usage:
mi_ex <- compute_mi_for_datasets(list(
   singleElimination_normal_noTies_1 = singleElimination_normal_noTies_1,
   singleElimination_normal_noTies_1_badseed = singleElimination_normal_noTies_1_badseed
 ))

plot_mi <- function(mi_results, dataset_names = NULL) {
  # If no dataset names specified, plot all datasets
  if (is.null(dataset_names)) {
    dataset_names <- mi_results$dataset
  }

  # Filter to only requested datasets
  filtered_results <- mi_results %>%
    filter(dataset %in% dataset_names)

  if (nrow(filtered_results) == 0) {
    stop("No matching datasets found in results.")
  }

  # Convert list column to long format for plotting
  plot_df <- filtered_results %>%
    rowwise() %>%
    mutate(
      index = list(1:length(mi)),
      mi_values = list(mi)
    ) %>%
    unnest(cols = c(index, mi_values)) %>%
    ungroup()

  # Create plot
  ggplot(plot_df, aes(x = index, y = mi_values, color = dataset, group = dataset)) +
    geom_line(linewidth = 1) +
    geom_point(size = 2) +
    labs(
      title = "Mutual Information Comparison",
      x = "Top K Teams",
      y = "Mutual Information (MI)",
      color = "Dataset"
    ) +
    scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
    scale_x_continuous(breaks = 1:max(plot_df$index)) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      legend.position = "bottom",
      legend.title = element_text(size = 11, face = "bold"),
      legend.text = element_text(size = 9)
    ) +
    scale_color_brewer(palette = "Set1")
}

plot_mi(mi_ex)
```

```{r}
make_perms_df <- function(num_teams){
  perms_list <- permn(1:num_teams)
  all_data <- vector("list", length(perms_list))

  for (i in seq_along(perms_list)) {
    all_data[[i]] <- data.frame(
      true_rank  = 1:num_teams,     # True rank positions
      rank_hat   = perms_list[[i]], # The current permutation
      simulation = i
    )
  }
  do.call(rbind, all_data)
}

# Generate datasets for 4, 6, and 10 teams
res <- data.frame(
  num_teams = 2:8,
  mutual_info = sapply(2:8, function(i){
    df <- make_perms_df(i)
    mut_info(df, num_teams = i)
  })
)

res

```

Why is it always 0.25?

$$
I(R;\hat{R}) = \sum_{R\in R} \sum_{\hat{R} \in \hat{R}} w(R,\hat{R})\times p(R,\hat{R})\times \log_2(\frac{p(R,\hat{R})}{p(R)\times p(\hat{R})})
$$

since every outcome is equally likely: $p(R,\hat{R}) = \frac{1}{n!}$

because $R$ and $\hat{R}$ are independent and have equal probabilities,

$H(R,\hat{R}) = H(R) + H(\hat{R}) = -2\sum_{x \in X} p(x)\log(p(x)) = -2(n!) \times (1/n!) \times\log_2(1/n!) = \log_2((n!)^2)$

$$
\frac{I(R,\hat{R})}{H(R,\hat{R})} = \frac{\sum \sum w(R,\hat{R}) \times \frac{1}{n!} \times \log_2 (\frac{1/n!}{(1/n!)^2})} {\log_2((n!)^2)}
$$

$\sum \sum w(R,\hat{R}) = 0.5\times n!$

$$
= \frac{0.5 \times n!\times \frac{1}{n!} \times \log_2(n!)}{2\times \log_2(n!)} =  \frac{0.5}{2} = 0.25
$$

include number of games (might be a range)

Single Elimination - third place and without third place; series = 3,5,7

Double Elimination

-   Repechage

-   True Double Elimination

-   Consolation Bracket (loser's bracket best place is third)

~~Triple Elimination~~

Deterministic Outcomes

Round Robin - with m rounds (1,2,4)

Stepladder- 8 vs 7, winner plays 6, etc.

Waterfall - stepladder round robin

Group Stage - winners of group advance, top 2 advance

Staged Round Robin - top 2 in each group form new group, bottom 2 new group (each new group is a new round robin)

High School Tennis Ladder - 1 vs 2, 3 vs 4, etc., Re-rank 2 vs 3, 4 vs 5, 6 vs 7, number of teams - 1 iterations

## Combinations

-   Manually setting the top-K teams with r=1 to look at if the top K teams were in the top K spots

```{r}
compute_mut_plots_combinations <- function(dataset, num_teams) {

  results <- numeric(num_teams)

  for (k in 1:num_teams) {
    df <- dataset %>%
      mutate(
        true_rank = ifelse(true_rank <= k, 1, true_rank),
        rank_hat = ifelse(rank_hat <= k, 1, rank_hat)
      )
    binary_vector <- c(rep(1, k), rep(0, num_teams - k))
    results[k] <- mut_info(df, num_teams, binary_vector)
  }
  return(results)
}

combinations_se <- compute_mut_plots_combinations(all_results_norm, num_teams=8)
combinations_rr <- compute_mut_plots_combinations(test_rr_noTies, num_teams=8)

# Define x values
x <- 1:8

# Combine into one data frame for plotting
plot_data <- data.frame(
  x = rep(x, 2),
  y = c(1-mi_se_norm, 1-combinations_se, 1-mi_test_rr, 1-combinations_rr),
  type = rep(c("mi_se_norm", "combinations_se", "mi_test_rr", "combinatinos_rr"), each = 8)
)

# Create plot
ggplot(plot_data, aes(x = x, y = y, color = type, group = type)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_x_continuous(breaks = 1:8) +
  labs(
    x = "Salient Ranks",
    y = "Value",
    color = "Series",
    title = "Comparison of MI vs Top K-teams"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "top",
    plot.title = element_text(hjust = 0.5)
  )
```