### 2025-11-13 simulations.qmd

---
title: "Simulations"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(combinat)
library(ggplot2)
library(future)
library(future.apply)
library(progressr)

# Use all cores except one
cl <- parallel::makeCluster(parallel::detectCores() - 2)
plan(cluster, workers = cl)

teams <- 16
```

Note: The naming of each data frame has the following structure: tournament_distribution_ties_r_info

-   tournament is the tournament structure
-   distribution is the name of the distribution for $\theta$ (Normal, Uniform, Same, Manual)
    -   Same is equal strengths or coin flip

    -   Manual is an user created $\theta$ vector passed by adding theta_hat = *user input* at the end of the function
-   ties: ties or noTies
-   r is the number of games in a series or the number of iterations of round robin or group stage
-   info is any additional information like reseeded, third place game, or bad seeding structure

# Getting Simulations

## Single Elimination

```{r}
single_elimination_param_grid <- expand.grid( 
  distribution = c("Normal","Uniform", "Same"), 
  ties = c(FALSE, TRUE), 
  series = c(1, 3, 7), 
  third_place = c(FALSE, TRUE), 
  reseeded = c(FALSE, TRUE), 
  bad_seeding_structure = c(FALSE, TRUE) 
  )

single_elimination_run_sim_block <- function(
  teams, distribution, ties, series, third_place, reseeded,
  bad_seeding_structure, p_block) {

  # Apply seeding structure rule
  seeding_structure <- if (bad_seeding_structure) 1:16 else NULL

  # Build the key
  parts <- c(
  "singleElimination",
  tolower(distribution),
  ifelse(ties, "ties", "noTies"),
  paste0(series),
  if (third_place) "thirdplace",
  if (reseeded) "reseeded",
  if (bad_seeding_structure) "badseed"
  )

  key <- paste(parts, collapse = "_")

  # Run the 10,000 tournament sims in parallel
  sim_list <- future_lapply(
    X = 1:10000,
    FUN = function(rep) {
      simulate_single_elimination_tournament(
        num_teams = teams,
        distribution = distribution,
        ties = ties,
        series = series,
        third_place = third_place,
        reseeded = reseeded,
        seeding_structure = seeding_structure
      )
    },
    future.seed = TRUE   # parallel-safe RNG
  )

  # Bind the simulation output
  final_df <- do.call(rbind, sim_list)
  final_df$teams <- teams

  assign(key, final_df, envir = .GlobalEnv)

  # Update outer progress bar
  p_block()

  message("Finished: ", key)
}

with_progress({
  p_block <- progressor(steps = nrow(single_elimination_param_grid))
  for (i in seq_len(nrow(single_elimination_param_grid))) {
    row <- single_elimination_param_grid[i, ]

    single_elimination_run_sim_block(
      teams        = teams,
      distribution = row$distribution,
      ties         = row$ties,
      series       = row$series,
      third_place  = row$third_place,
      reseeded     = row$reseeded,
      bad_seeding_structure = row$bad_seeding_structure,
      p_block      = p_block
    )
  }
})

```

## Double Elimination

```{r}
double_elimination_param_grid <- expand.grid( 
  distribution = c("Normal","Uniform", "Same"), 
  ties = c(FALSE, TRUE), 
  series = c(1, 3, 7), 
  final = c(FALSE, TRUE), 
  bad_seeding_structure = c(FALSE, TRUE) 
  )

double_elimination_run_sim_block <- function(
  teams, distribution, ties, series, final, bad_seeding_structure, p_block) {

  # Apply bad seeding structure override if required
  seeding_structure <- if (bad_seeding_structure) 1:16 else NULL

  # Build key
    # Build the key
  parts <- c(
    "doubleElimination",
    tolower(distribution),
    ifelse(ties, "ties", "noTies"),
    paste0(series),
    if (!final) "consolation",
    if (bad_seeding_structure) "badseed"
  )

  key <- paste(parts, collapse = "_")

  # Run 10,000 simulations in parallel
  sim_list <- future_lapply(
    X = 1:10000,
    FUN = function(rep) {
      simulate_double_elimination_tournament(
        num_teams = teams,
        distribution = distribution,
        ties = ties,
        series = series,
        final = final,
        seeding_structure = seeding_structure
      )
    },
    future.seed = TRUE    # ensures parallel RNG is valid
  )

  final_df <- do.call(rbind, sim_list)
  final_df$teams <- teams

  assign(key, final_df, envir = .GlobalEnv)

  # UPDATE BLOCK-LEVEL PROGRESS BAR
  p_block()

  message("Finished: ", key)
}

with_progress({
  p_block <- progressor(steps = nrow(double_elimination_param_grid))

  for (i in seq_len(nrow(double_elimination_param_grid))) {
    row <- double_elimination_param_grid[i, ]

    double_elimination_run_sim_block(
      teams        = teams,
      distribution = row$distribution,
      ties         = row$ties,
      series       = row$series,
      final        = row$final,
      bad_seeding_structure = row$bad_seeding_structure,
      p_block = p_block
    )
  }
})

```

## Round Robin

```{r}
round_robin_param_grid <- expand.grid( 
  distribution = c("Normal","Uniform", "Same"), 
  ties = c(FALSE, TRUE), 
  r = c(1,2,4)
  )

round_robin_run_sim_block <- function(
  teams, distribution, ties, r, p_block) {

  # Build key
    # Build the key
  parts <- c(
    "roundRobin",
    tolower(distribution),
    ifelse(ties, "ties", "noTies"),
    paste0(r)
  )

  key <- paste(parts, collapse = "_")

  # Run 10,000 simulations in parallel
  sim_list <- future_lapply(
    X = 1:10000,
    FUN = function(rep) {
      simulate_roundrobin_r(
        num_teams = teams,
        distribution = distribution,
        ties = ties,
        r = r
      )
    },
    future.seed = TRUE    # ensures parallel RNG is valid
  )

  final_df <- do.call(rbind, sim_list)
  final_df$teams <- teams

  assign(key, final_df, envir = .GlobalEnv)

  # UPDATE BLOCK-LEVEL PROGRESS BAR
  p_block()

  message("Finished: ", key)
}

with_progress({
  p_block <- progressor(steps = nrow(round_robin_param_grid))

  for (i in seq_len(nrow(round_robin_param_grid))) {
    row <- round_robin_param_grid[i, ]

    round_robin_run_sim_block(
      teams        = teams,
      distribution = row$distribution,
      ties         = row$ties,
      r            = row$r,
      p_block = p_block
    )
  }
})
```

## Group Stage

```{r}
# Group stage is currently set so only the top 2 in each group advance
group_stage_param_grid <- expand.grid( 
  distribution = c("Normal","Uniform", "Same"), 
  ties = c(FALSE, TRUE), 
  rounds = c(1,2,4),
  group_length = c(4,8)
  )

group_stage_run_sim_block <- function(
  teams, distribution, ties, rounds, group_length, p_block) {

  # Build key
    # Build the key
  parts <- c(
    "groupStage",
    tolower(distribution),
    ifelse(ties, "ties", "noTies"),
    paste0(rounds),
    paste0(group_length, "groups")
  )

  key <- paste(parts, collapse = "_")

  # Run 10,000 simulations in parallel
  sim_list <- future_lapply(
    X = 1:10000,
    FUN = function(rep) {
      simulate_group_stage_tournament(
        num_teams = teams,
        distribution = distribution,
        ties = ties,
        rounds = rounds,
        group_length = group_length
      )
    },
    future.seed = TRUE    # ensures parallel RNG is valid
  )

  final_df <- do.call(rbind, sim_list)
  final_df$teams <- teams

  assign(key, final_df, envir = .GlobalEnv)

  # UPDATE BLOCK-LEVEL PROGRESS BAR
  p_block()

  message("Finished: ", key)
}

with_progress({
  p_block <- progressor(steps = nrow(group_stage_param_grid))

  for (i in seq_len(nrow(group_stage_param_grid))) {
    row <- group_stage_param_grid[i, ]

    group_stage_run_sim_block(
      teams        = teams,
      distribution = row$distribution,
      ties         = row$ties,
      rounds       = row$rounds,
      group_length = row$group_length,
      p_block = p_block
    )
  }
})

```

## Waterfall

## Step Ladder

```{r}
stepLadder_param_grid <- expand.grid( 
  distribution = c("Normal","Uniform", "Same"), 
  series = c(1, 3, 7), 
  bad_seeding_structure = c(FALSE, TRUE) 
  )

stepLadder_run_sim_block <- function(
  teams, distribution, bad_seeding_structure, series, p_block) {

  # Apply bad seeding structure override if required
  seeding_structure <- if (bad_seeding_structure) 1:16 else NULL

  # Build key
    # Build the key
  parts <- c(
    "stepLadder",
    tolower(distribution),
    paste0(series),
    if (bad_seeding_structure) "badseed"

  )

  key <- paste(parts, collapse = "_")

  # Run 10,000 simulations in parallel
  sim_list <- future_lapply(
    X = 1:10000,
    FUN = function(rep) {
      simulate_step_ladder(
        num_teams = teams,
        distribution = distribution,
        series = series,
        seeding_structure = seeding_structure
      )
    },
    future.seed = TRUE    # ensures parallel RNG is valid
  )

  final_df <- do.call(rbind, sim_list)
  final_df$teams <- teams

  assign(key, final_df, envir = .GlobalEnv)

  # UPDATE BLOCK-LEVEL PROGRESS BAR
  p_block()

  message("Finished: ", key)
}

with_progress({
  p_block <- progressor(steps = nrow(stepLadder_param_grid))

  for (i in seq_len(nrow(stepLadder_param_grid))) {
    row <- stepLadder_param_grid[i, ]

    stepLadder_run_sim_block(
      teams        = teams,
      distribution = row$distribution,
      series       = row$series,
      bad_seeding_structure = row$bad_seeding_structure,
      p_block = p_block
    )
  }
})
```

## Staged Round Robin

## High School Tennis Ladder

# Getting Mutual Information

```{r}
for (df in single_elimination_df){
  MI_singleElimination <- compute_mi_for_datasets(df)
}

```