### 2025-12-05 double_elimination.qmd
updates

---
title: "Double Elimination Tournament"
author: "Zach Culp"
format: html
editor: visual
---

# Double Elimination Tournament

```{r}
library(tidyverse)
library(dplyr)
library(shiny)
```

## Simulate a Single Match

```{r}
# Function to simulate a match using Bradley-Terry model
simulate_match <- function(team1, team2, strength1, strength2) {
  if(is.na(team1) && is.na(team2)){
    return(NA)
  }
  if (is.na(team1)) {
    return(team2)
  } else if (is.na(team2)) {
    return(team1)
  } else {
    p <- exp(strength1 - strength2) / (1 + exp(strength1 - strength2))
    return(ifelse(runif(1) < p, team1, team2))  # Randomize outcome
  }
}
```

## Creating the traditional tournament structure for n teams

```{r}
generate_elimination_tournament_structure <- function(n) {

  is_power_of_two <- function(x) {
    x > 0 && (x & (x - 1)) == 0
  }

  # If n is not a power of 2, round up to the next power of 2 and add placeholders
  if (!is_power_of_two(n)) {
    next_power_of_two <- 2^ceiling(log2(n))
    extra_teams <- next_power_of_two - n
    #message(sprintf("Rounding up to %d teams. Adding %d NA placeholders.", next_power_of_two, extra_teams))
    n <- next_power_of_two
  } else {
    extra_teams <- 0
  }

  rounds <- log(n, base = 2) - 1
  teams <- c(1, 2)  # Initial list of teams

  for (i in 1:rounds) {
    teams <- nextLayer(teams)
  }

  left_half <- teams[1:(length(teams)/2)]
  right_half <- teams[(length(teams)/2+1):length(teams)]
  right_half <- rev(right_half)
  teams <- c(left_half,right_half)

  # Ensure that the lower number team is on the left side of each matchup
  teams <- unlist(lapply(seq(1, length(teams), by = 2), function(i) {
    if (teams[i] > teams[i + 1]) {
      c(teams[i + 1], teams[i])  # Swap if needed
    } else {
      c(teams[i], teams[i + 1])
    }
  })
  )

  while (extra_teams > 0){
      teams[which.max(teams)] <- NA
      extra_teams <- extra_teams - 1
  }

  return(teams)
}

nextLayer <- function(teams) {
  out <- c()  # Initialize an empty vector
  length <- length(teams) * 2 + 1  # Calculate the length for pairing

  # Generate the next layer by pairing teams
  for (i in teams) {
    out <- c(out, i, length - i)  # Push the current team and its pair
  }

  return(out)
}

```

## Creating a Single Simulation of a Double Elimination Tournament

```{r}
simulate_double_elimination_tournament <- function(num_teams, distribution, seeding_structure = NULL, ties = T, series = 1, final = T, theta_hat = NULL) {
  if (num_teams <= 3) {
    stop("Number of Teams must be greater than 3.")
  }

  teams <- paste(num_teams:1)

  if (log2(num_teams) %% 1 != 0) {
    power_of_2 <- 2^ceiling(log2(num_teams))
    teams <- c(teams, rep(NA, power_of_2 - num_teams))
  }

  if (distribution == "Normal") {
    strengths <- sapply(num_teams, function(n) {
      qnorm(1:n/(n+1))
    })
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  } else if (distribution == "Uniform") {
    strengths <- sapply(num_teams, function(n) {
      qunif(1:n/(n+1), 0, sqrt(12))
    })
    unif_strength <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams)),
      rank_hat = rep(NA, length(teams))
    )
    df <- arrange(unif_strength, true_rank)
  } else if (distribution == "Same"){
    strengths <- sapply(num_teams, function(n) {rep(0,n)})
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA,length(teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  } else if (distribution == "Exponential"){
      strengths <- sapply(num_teams, function(n) { qexp(1:n/(n+1)) })
      unif_strength <- data.frame(
        true_rank = as.numeric(teams),
        true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
        game_wins = rep(0,length(teams)),
        game_losses = rep(0,length(teams)),
        rank_hat = rep(NA,length(teams))
      )
      df <- arrange(unif_strength, true_rank)
  } else if (distribution == "Manual") {
    strengths <- theta_hat
    manual_strengths <- data.frame(
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- manual_strengths %>%
      arrange(true_strength) %>%
      mutate(true_rank = as.numeric(teams)) %>%
      arrange(true_strength, true_rank)
  } else {
    stop("Distribution not found: Enter Manual to input your own strengths")
  }

  # Sets seed to that of common best vs worst structure
  if (is.null(seeding_structure)) {
    seeding_structure <- matrix(generate_elimination_tournament_structure(num_teams), nrow = 1)
  } else {
    seeding_structure <- matrix(seeding_structure, nrow = 1)
  }

  if (series %% 2 == 0) {
    stop("Series must be an odd number to avoid potential ties.")
  }

  results <- list() # Initialize the results list

  for (i in 1:nrow(seeding_structure)) {
    # Create a new data frame based on the current permutation in 'test'
    permutation <- seeding_structure[i, ]
    permuted_df <- df
    permuted_df$true_rank <- df$true_rank[permutation] # Update team order
    teams <- permuted_df$true_rank
    permuted_df$true_strength <- df$true_strength[permutation]
    permuted_df$rank_hat <- NA # Initialize ranks column
    permuted_df$game_wins <- 0 # Initialize wins column
    permuted_df$game_losses <- 0

    team_rank <- num_teams
    winners_bracket <- teams
    strengths <- permuted_df$true_strength
    losers_bracket <- c()

    # gets results of the winners bracket
    while (length(winners_bracket) > 1 || length(losers_bracket > 0)) {
      eliminated <- c()

      if (length(winners_bracket) > 1) {
        #cat("\n--- Winner's Bracket --- \n")
        next_winners <- c()
        round_losers <- c()

        for (j in seq(1, length(winners_bracket), by = 2)) {
          # Gets 2 teams for each match
          if (j + 1 > length(winners_bracket)) break

          team1 <- winners_bracket[j]
          team2 <- winners_bracket[j+1]
          strength1 <- strengths[which(permuted_df$true_rank == team1)]
          strength2 <- strengths[which(permuted_df$true_rank == team2)]

          wins_team1 <- 0
          wins_team2 <- 0
          losses_team1 <- 0
          losses_team2 <- 0

          for (game in 1:series) {
            match_winner <- simulate_match(team1, team2, strength1, strength2)
            #cat(sprintf("Matchup: %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", team1, team2, strength1, strength2, match_winner))

            if (!is.na(team1)) {
              if (match_winner == team1) {
                wins_team1 <- wins_team1 + 1
                losses_team2 <- losses_team2 + 1
              } else {
                wins_team2 <- wins_team2 + 1
                losses_team1 <- losses_team1 + 1
              }
            } else {
              wins_team2 <- (series+1) / 2
            }

            # Determine the overall winner if a majority is reached
            if (wins_team1 > series / 2 || wins_team2 > series / 2) {
              break
            }
          }

          # Update the game_wins and game_losses outside of the loop
          if (!any(is.na(c(team1, team2)))) {
            #cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
            permuted_df <- permuted_df %>%
              mutate(game_wins = case_when(
                true_rank == team1 ~ game_wins + wins_team1,
                true_rank == team2 ~ game_wins + wins_team2,
                TRUE ~ game_wins
              )) %>%
              mutate(game_losses = case_when(
                true_rank == team1 ~ game_losses + losses_team1,
                true_rank == team2 ~ game_losses + losses_team2,
                TRUE ~ game_losses
              ))
          }

          # Determine the match winner based on series results
          series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
          series_loser <- ifelse(series_winner == team1, team2, team1)
          #cat(sprintf("Matchup: %s vs %s | Winner: %s\n: | Loser: %s\n", team1, team2, series_winner, series_loser))

          next_winners <- c(next_winners, series_winner)
          round_losers <- c(round_losers, series_loser)
        }

        winners_bracket <- next_winners # Winners move to the next round
        losers_bracket <- c(losers_bracket, round_losers)
        if (length(losers_bracket) %% 2 != 0) {
          losers_bracket <- c(losers_bracket, NA)
        }
      }

      # Losers' bracket
      if (length(losers_bracket) > 0) {
        if (length(losers_bracket) > 1) {
          #cat("\n--- Loser's Bracket --- \n")
        }
        next_losers <- c()

        # alternates games for losers bracket
        if (length(losers_bracket) > 2) {
          if (length(losers_bracket) == 4 && is.na(losers_bracket[4])) {
            losers_bracket <- losers_bracket
          } else {
            top_bracket <- losers_bracket[seq(1, length(losers_bracket), by = 2)]
            bottom_bracket <- losers_bracket[seq(2, length(losers_bracket), by = 2)]
            losers_bracket <- c(top_bracket, bottom_bracket)
          }
        }

        for (j in seq(1, length(losers_bracket), by = 2)) {
          if (j + 1 > length(losers_bracket)) break

          team1 <- losers_bracket[j]
          team2 <- losers_bracket[j + 1]
          strength1 <- strengths[which(permuted_df$true_rank == team1)]
          strength2 <- strengths[which(permuted_df$true_rank == team2)]

          wins_team1 <- 0
          wins_team2 <- 0
          losses_team1 <- 0
          losses_team2 <- 0

          # Play series
          for (game in 1:series) {
            match_winner <- simulate_match(team1, team2, strength1, strength2)
            if (is.na(match_winner)) {
              wins_team1 == wins_team1 + 1
            }

            if (!is.na(team1)) {
              if (match_winner == team1) {
                wins_team1 <- wins_team1 + 1
                losses_team2 <- losses_team2 + 1
              } else {
                wins_team2 <- wins_team2 + 1
                losses_team1 <- losses_team1 + 1
              }
            } else {
              wins_team2 <- (series+1) / 2
            }

            # Determine overall series winner
            if (wins_team1 > series / 2 || wins_team2 > series / 2) {
              break
            }
          }

          if (!any(is.na(c(team1, team2)))) {
            #cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
            permuted_df <- permuted_df %>%
              mutate(game_wins = case_when(
                true_rank == team1 ~ game_wins + wins_team1,
                true_rank == team2 ~ game_wins + wins_team2,
                TRUE ~ game_wins
              )) %>%
              mutate(game_losses = case_when(
                true_rank == team1 ~ game_losses + losses_team1,
                true_rank == team2 ~ game_losses + losses_team2,
                TRUE ~ game_losses
              ))
          }

          if (is.na(team1) && is.na(team2)) {
            series_winner <- NA
            next_losers <- c(next_losers, NA)
          } else {
            series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
            series_loser <- ifelse(wins_team1 > wins_team2, team2, team1)
            next_losers <- c(next_losers, series_winner)
            eliminated <- c(eliminated, series_loser)
            #cat(sprintf("Matchup: %s vs %s | Winner: %s\n:", team1, team2, series_winner))
          }
        }

        team_rank <- team_rank - 1
        permuted_df$rank_hat[permuted_df$true_rank %in% eliminated] <- team_rank
        losers_bracket <- next_losers
      }

      # Final match for loser's bracket (aka third place game)
      if (length(winners_bracket) == 1 && length(losers_bracket) == 1) {
        if (final == F){
          # No final game played between winner of winner bracket and winner of loser bracket
          team1 <- winners_bracket[1]
          team2 <- losers_bracket[1]
          permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
          permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
          break
        }
        # Final matchup between winners and losers
        else if (final == T) {
          #cat("\n--- Final Match ---\n")
          final_match <- c(winners_bracket, losers_bracket)

          if (length(final_match) == 2) {
            team1 <- final_match[1]
            team2 <- final_match[2]
            strength1 <- strengths[which(permuted_df$true_rank == team1)]
            strength2 <- strengths[which(permuted_df$true_rank == team2)]

            wins_team1 <- 0
            wins_team2 <- 0
            losses_team1 <- 0
            losses_team2 <- 0

            for (game in 1:series) {
              match_winner <- simulate_match(team1, team2, strength1, strength2)
              if (match_winner == team1) {
                wins_team1 <- wins_team1 + 1
                losses_team2 <- losses_team2 + 1
              } else {
                wins_team2 <- wins_team2 + 1
                losses_team1 <- losses_team1 + 1
              }
              if (wins_team1 > series / 2 || wins_team2 > series / 2) {
                break
              }
            }

            if (!any(is.na(c(team1, team2)))) {
              #cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
              permuted_df <- permuted_df %>%
                mutate(game_wins = case_when(
                  true_rank == team1 ~ game_wins + wins_team1,
                  true_rank == team2 ~ game_wins + wins_team2,
                  TRUE ~ game_wins
                )) %>%
                mutate(game_losses = case_when(
                  true_rank == team1 ~ game_losses + losses_team1,
                  true_rank == team2 ~ game_losses + losses_team2,
                  TRUE ~ game_losses
                ))
            }

            if (wins_team1 > wins_team2) {
              #cat(sprintf("Matchup: %s vs %s | Winner: %s\n:", team1, team2, team1))
              permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
              permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
            } else {
              #cat(sprintf("Matchup: %s vs %s | Winner: %s\n:", team1, team2, team2))
              # second game if the loser's bracket team wins
              # RESET the counters for the second series
              wins_team1 <- 0
              wins_team2 <- 0
              losses_team1 <- 0
              losses_team2 <- 0

              for (game in 1:series) {
                match_winner <- simulate_match(team1, team2, strength1, strength2)
                if (match_winner == team1) {
                  wins_team1 <- wins_team1 + 1
                  losses_team2 <- losses_team2 + 1
                } else {
                  wins_team2 <- wins_team2 + 1
                  losses_team1 <- losses_team1 + 1
                }
                if (wins_team1 > series / 2 || wins_team2 > series / 2) {
                  break
                }
              }

              if (!any(is.na(c(team1, team2)))) {
                #cat(sprintf("Wins - %s: %d | %s: %d\n", team1, wins_team1, team2, wins_team2))
                permuted_df <- permuted_df %>%
                  mutate(game_wins = case_when(
                    true_rank == team1 ~ game_wins + wins_team1,
                    true_rank == team2 ~ game_wins + wins_team2,
                    TRUE ~ game_wins
                  )) %>%
                  mutate(game_losses = case_when(
                    true_rank == team1 ~ game_losses + losses_team1,
                    true_rank == team2 ~ game_losses + losses_team2,
                    TRUE ~ game_losses
                  ))
              }

              if (wins_team1 > wins_team2) {
                #cat(sprintf("Matchup: %s vs %s | Winner: %s\n:", team1, team2, team1))
                permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
                permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
              } else {
                #cat(sprintf("Matchup: %s vs %s | Winner: %s\n:", team1, team2, team2))
                permuted_df$rank_hat[permuted_df$true_rank == team2] <- 1
                permuted_df$rank_hat[permuted_df$true_rank == team1] <- 2
              }
            }
          }
        }
      }
    }

    permuted_df$seed <- rep(paste(seeding_structure[i, ], collapse = ", "), nrow(permuted_df))
    results[[i]] <- permuted_df
  }

  final_results <- do.call(rbind, results)

  # Remove ties if ties = FALSE
  if (!ties) {
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(true_rank) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "random"))
  } else {
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(rank_hat) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "average"))
  }

  final_results <- final_results %>%
    mutate(distribution = distribution) %>%
    mutate(true_rank = as.numeric(true_rank)) %>%
    arrange(true_rank)

  return(final_results)
}

```